### 2017-11-02 paper/abstract.tex
backu

\begin{abstract}

In recent years, The end of Dennard Scaling \cite{Dark-Silicon} and the need for agile development methodology in data centers is pushing the computer architecture community towards designing more specialized and energy efficient systems \cite{Catapult}. This has lead to a new application domain for FPGAs and the emergence of heterogeneous CPU-FPGA computing platforms \cite{Intel-FPGA} \cite{IBM-FPGA}, enabling the design of new energy efficient FPGA accelerators for domain specific applications. Graph Analytics, one the largest applications in production data centers today, faces scaling challenges with the ever increasing size of its workloads and its inherent low locality and memory-bound characteristic. The GraphMat \cite{GraphMat} programming framework was proposed that efficiently translates graph algorithms into high performance computation in hardware. We present FlexGraph, a reconfigurable domain-specific accelerator for Graph Analytics targeting commodity heterogeneous CPU-FPGA platforms. FlexGraph is designed using Cocoh \cite{Cocoh}, a C++ domain specific library for hardware design and simulation, providing a high level programming abstraction with the efficiency of native RTL. FlexGraph's novel architecture allows an easy implementation of GraphMat kernels using standard C++. To the best of our knowledge it is a first implementation of a graph accelerator on a commodity heterogeneous CPU-FPGA platform. FlexGraphs's preliminary performance is comparable to HLS, we later discuss several optimization strategies that could further improve its speed.

\end{abstract}

### 2017-11-02 paper/evaluation.tex
backu

\section{Experimental Setup}

\subsection{Cocoh's Simulation}

\subsection{Intel AAL Interface}

\subsection{Intel QPI Emulation}

\subsection{Graph Datasets}

\section{Results Analysis}

\subsection{Overall Performance}

\subsection{The Impact of Parallelism}

\subsection{The Impact of Memory Optimizations}

\subsection{FlexGraph Performance Hogs}

\subsection{Hardware Cost Evaluation}

### 2017-11-02 paper/introduction.tex
backu

\section{Introduction}

The end of Dennard Scaling \cite{Dark-Silicon} has moved the focus of computer architecture designers towards power and energy efficient architectures. However, energy efficient solutions such as ASIC designs present a serious limitation in flexibility and production cycle. These constraints have pushed production data centers towards using FPGA-based accelerators \cite{Catapult} for their reconfigurability and energy savings when compared to general-purpose graphics processing units (GPUs).
This trend has led to emergence of heterogeneous CPU-FPGA computing platforms \cite{Intel-FPGA} \cite{IBM-FPGA}, enabling the design of new energy efficient FPGA accelerators for domain specific applications. Graph Analytics is one the largest applications in production data centers today \cite{BigData}, its performance surfers from workload imbalance, frequent updates, limited data locality and low compute communication ratio making it memory-bound and energy inefficient. This problem is further exacerbated with the ever increasing size of its dataset, presenting a scalability challenge for the industry. Several solution have been proposed to address this problem both at the software level with better algorithms and programming abstraction \cite{GraphX} \cite{Galois} \cite{GraphMat} \cite{Pregel} \cite{GraphLab}, and at the hardware level with custom accelerators \cite {Graphicionado} \cite{Tesseract} \cite{GraphOps}. GraphMat \cite{GraphMat} and many other graph frameworks \cite{GraphMat} \cite{Pregel} \cite{GraphLab} define a vertex programming model to represent graph computation, but GraphMat is unique in that only the edges transformation of the kernel is accelerated, allowing the programmer to implement its algorithm is C++, instead of a proprietary language, increasing both accessibility and productivity. Flexgraph's design goals are aiming into achieving the same balance between efficiency and productivity via hardware acceleration. We developed FlexGraph using the Cocoh's \cite{Cocoh} framework, a C++ domain specific library for hardware design and simulation, providing a high level programming abstraction with the efficiency of native RTL. Several hardware description languages and libraries \cite{CHDL} \cite{Chisel} \cite{SystemC} \cite{BlueSpec} \cite{MyHDL} \cite{JHDL} \cite{Esterel} have been proposed as alternative to native verilog or VHDL to improve developer productivity by providing a higher level programming abstraction \cite{Chisel} \cite{SystemC} and execution model \cite{BlueSpec} \cite{Esterel}. Cocoh's unique strength is that it provides a uniform development and simulation environment based on the same source. Cocoh's modules are written directly in standard C++, debugged, tested and simulated using the same source, and later exported to verilog for FPGA deployment. Another strength of Cocoh's API is its integration with CHDL \cite{CHDL}, allowing gate-level simulation and analysis of the design. Using the Cocoh's API, developpers can write their custom GraphMat kernel for the FlexGraph accelerator and test their design before production. FlexGraph architecture uniquely targets commodity heteregenous CPU-FPGA computing platforms \cite{Intel-FPGA} for data centers. To the best of our knowledge it is a first implementation of a graph accelerator targeting this ecosystem. FlexGraphs's current performance is comparable to HLS, we have identified several performance optimizations that could further improve its performance which we will discuss later in the paper.
The remainder of this paper is organized as follows:
Section 2 describes GraphMat computing model and its SPMV kernel, section 3 describes FlexGraph's architecture, section 4 describes the experimental setup, section 5 describes our results and analysis, section 6 describes the related work, section 7 discusses future work and finally Section 8 summarizes our main contribution and results.

### 2017-11-02 paper/mechanism.tex
backu

\section{FlexGraph Architecture}

In this section, we describes the overall architecture of the FlexGraph Accelerator.

\subsection{FlexGraph Computing Model}

\begin{figure}[htbp]
\includegraphics[width=0.4\textwidth]{figures/graph_vertex_programming}
\caption{Graph Vertex Programming}
\label{fig:graph_vertex_programming}
\end{figure}

FlexGraph's architecture is based on GraphMat \cite{GraphMat}'s vertex computing paradigm where a graph algorithm is described as a linear algebra set of operations operating on sparse matrix with input vertices. Figure \ref{fig:graph_vertex_programming} shows an example where graph (a) is converted into matrix (b), allowing us to execute operations on the graph using simple matrix vertex multiplication (c). This above example will to generate the number of incoming edges for each vertex. All GraphMat algorithms, Breadth First Search (BFS) \cite{BFS}, PageRank \cite{PageRank}, Single Source Shortest-Path (SSSP) \cite{SSSP}, etc.. are expressed in this fashion such that computation can be efficiently accelerated in hardware. In FlexGraph, we accelerate the matrix vertex computation on the FPGA, instead of executing it on the multi-core host processor.

\subsection{The Doubly Compressed Sparse Matrix}

\begin{figure}[htbp]
\includegraphics[width=0.4\textwidth]{figures/DCSC_matrix_format}
\caption{DCSC hyperspace matrix format}
\label{fig:DCSC_matrix_format}
\end{figure}

Most Data Analytics graphs are represented in a sparse matrix format to save memory space due to the large amount of empty edges that exist. FlexGraph's matrices are stored using a Doubly Compressed Sparse Column (DCSC) \cite{DCSC} format. The format allows minimal traversal into the sparse matrix structure, saving necessary memory bandwidth when fetching empty columns. Figure \ref{fig:DCSC_matrix_format} shows a sample matrix A = \{(0,5,a), (0,7,b), (6,3,c), (7,1,d)\} with four non-zero edges (src, dst, weight) encoded using DCSC versus the conventional CSC \cite{CSC} format. It is encoded as follows: The CSC encoding for this matrix is as follows: The \textit{Cols} buffer holds the column ranges stored as incremental distances from first value, which allows efficient packing of consecutive non-zero columns. The limitation of this scheme is that it entries are replicated to skip empty columns. \textit{Rows} and \textit{Values} hold the corresponding rows and values referenced by a given non-empty column.
The traversal over CSC requires accessing all consecutive column ranges \textit{(cols[i+2] - cols[i])} even though most of the distances are empty. DCSC saves on bandwidth by using an additional indirection  buffer inside its structure to only encode non-zero columns. It is important to note that the saving this format introduced additional space and compute overhead that only pays off when the matrix is very large and sparse, a prevalent characteristic of Graph Analytics datasets.

\subsection{Sparse Matrix Multiplication Kernel}

\begin{lstlisting}[language=Python, caption=Pseudocode for SPMV kernel]
def SPMV_kernel(x_values, x_activemask) :
  y_values = {0}, y_activemask = {0}
    for col in cols :                        # +	
      (a_x, rows) = coldata[col]             # *
      x_value  = x_values[a_x]               # *
      x_active = x_activemask[a_x]           # *
      if (x_active):                         # ?
        for row in rows:                     # +
          (a_y, a_value) = rowdata[row]      # *
          y_values[a_y] += a_value * x_value # *
          x_activemask[a_y] = true           # *
  return (y_values, y_activemask)
\end{lstlisting}

\subsection{FlexGraph Micro-architecture}

\section{FlexGraph Parallelization}

\subsection{DCSC Matrix partitioning}

\subsection{Precessing Tasks Scheduling}

\subsection{Synchronising Memory Writes for Dirty Masks}

\section{FlexGraph Optimizations}

\subsection{Optimizing Memory Accesses via Stream Buffers}

\subsection{Caching Vertex Values and Masks}

### 2017-11-02 paper/related_work.tex
backu

\section{Related Work}

\subsection{Graphicionado}

\subsection{GraphOps}

\subsection{Tesseract}

### 2017-11-02 aal.h

@@ -1,0 +1,82 @@
+#pragma once
+
+#include <cash.h>
+
+namespace aal {
+
+//
+// Description:
+// - memory data are 64 byte cache line width
+// - there are two write response ports
+// - memory responses are returned in order
+//
+
+using namespace ch::core;
+
+template <unsigned ADDR, unsigned MDATA, unsigned CDATA>
+struct aal_qpi {
+  static constexpr unsigned addr_wdith  = ADDR;
+  static constexpr unsigned mdata_width = MDATA;
+  static constexpr unsigned cdata_width = CDATA;
+
+  __inout (rd_req_t, (
+    (ch_in<ch_bit1>)        almostfull,
+    (ch_out<ch_bit<ADDR>>)  addr,
+    (ch_out<ch_bit<MDATA>>) mdata,
+    (ch_out<ch_bit1>)       valid
+  ));
+
+  __inout (rd_rsp_t, (
+    (ch_in<ch_bit<MDATA>>) mdata,
+    (ch_in<ch_bit<CDATA>>) data,
+    (ch_in<ch_bit1>)       valid
+  ));
+
+  __inout (wr_req_t, (
+    (ch_in<ch_bit1>)        almostfull,
+    (ch_out<ch_bit<ADDR>>)  addr,
+    (ch_out<ch_bit<MDATA>>) mdata,
+    (ch_out<ch_bit<CDATA>>) data,
+    (ch_out<ch_bit1>)       valid
+  ));
+
+  __inout (wr_rsp_t, (
+    (ch_in<ch_bit<MDATA>>) mdata,
+    (ch_in<ch_bit1>)       valid
+  ));
+
+  __inout (io_t, (
+    (rd_req_t) rd_req,
+    (rd_rsp_t) rd_rsp,
+    (wr_req_t) wr_req,
+    (wr_rsp_t) wr_rsp0,
+    (wr_rsp_t) wr_rsp1
+  ));
+};
+
+using aal_qpi0 = aal_qpi<20, 14, 512>;
+
+template <typename Context, typename Qpi = aal_qpi0>
+class aal_device {
+public:
+  using qpi_io_t = typename Qpi::io_t;
+  static_assert(ch_bitwidth_v<Context> == Qpi::cdata_width, "invalid context size");
+  static_assert(ch_direction_v<Context> == ch_direction::in, "invalid context direction");
+
+  __io (
+    (qpi_io_t)         qpi,
+    (Context)          ctx,
[…]

### 2017-11-02 common.h

@@ -1,0 +1,129 @@
+#pragma once
+
+#include <string.h>
+#include <cash.h>
+#include <../stl/queue.h>
+#include "aal.h"
+
+using namespace ch::core;
+using namespace ch::literals;
+using namespace ch::stl;
+using namespace ch::sim;
+
+#define LOG2_BLOCK_SIZE   6
+#define BLOCK_SIZE        (1 << LOG2_BLOCK_SIZE)
+#define BLOCK_SIZE_MASK   (BLOCK_SIZE-1)
+
+#define __countof(a) (sizeof(a) / sizeof(a[0])) 
+
+#define __safediv(a, b) ((b) != 0) ? ((a) / (b)) : 0 
+
+#define __align(x, alignment) \
+  (((x) + ((alignment)-1)) & ~((alignment)-1))
+
+#define __div_ceil(x, y) (((x) + (y) - 1) / (y))
+#define __div_rnd(x, y)  (((x) + ((y)/2)) / (y))
+
+#define CHECK(x) if (!(x)) { assert(false); exit(1); }
+
+namespace spmv {
+
+using ch_block_t = ch_bit<aal::aal_qpi0::cdata_width>;
+
+enum {
+  PE_COUNT       = 2,
+  LOG2_PE_COUNT  = log2ceil(PE_COUNT),
+  RQ_OWNER_BITS  = log2ceil(PE_COUNT+1),
+
+  PTR_MAX_VALUE  = (1 << aal::aal_qpi0::addr_wdith)-1,
+
+  PARTITION_SIZE = 32,
+
+  PARTITION_VALUE_BITS = 32,
+  PARTITIONS_PER_BLOCK = ch_bitwidth_v<ch_block_t> / PARTITION_VALUE_BITS,
+  LOG2_PARTITIONS_PER_BLOCK = log2ceil(PARTITIONS_PER_BLOCK+1),
+};
+
+extern bool g_singlecore; 
+
+using byte_t = uint8_t;
+
+void DbgPrint(int level, const char *format, ...);
+
+inline void set_bitvector(uint32_t* bitvec, uint32_t idx) {
+  uint32_t block = idx >> 5;
+  uint32_t bit   = idx & 31;
+  uint32_t value = bitvec[block];
+  if ((value & (1 << bit)) == 0) {
+    bitvec[block] = value | (1 << bit);
+  }
+}
+
[…]

### 2017-11-02 cpu.cpp

@@ -1,0 +1,367 @@
+#include "math.h"
+#include "cpu.h"
+
+using namespace spmv;
+using namespace spmv::host;
+
+static int random_select(int vmin, int vmax, float skew) {
+  float alpha((rand() * 1.0f) / RAND_MAX);
+  return (int)std::ceil(vmin + (vmax - vmin) * alpha * skew);  
+}
+
+// poluate random values into vertex
+static void init_vertex(vertex_t* vertex) {
+  float sample[] = {0.0, 0.1, 0.2, 0.0, 0.4, 0.5, 0.0, 0.7, 0.8, 0.0, 1.0};  
+  uint32_t size = vertex->size;
+  float* const values = reinterpret_cast<float*>(vertex->values);
+  uint32_t* const masks = vertex->masks;
+  for (uint32_t i = 0; i < size; ++i) {
+    float value = sample[ random_select(0, __countof(sample)-1, 1.0) ];
+    values[i] = value;
+    uint32_t ivalue = *(uint32_t*)&value;
+    if (value != 0.0) {
+      set_bitvector(masks, i);
+    }
+  }           
+}
+
+cpu_device::cpu_device(const char* mtx_file) 
+  : m_shared_mem(nullptr) 
+  , m_matrix(nullptr)
+  , m_vertex(nullptr) {
+  // initialiaze matrix file
+  this->init(mtx_file);
+
+  // dump stats
+  ch_dumpStats(std::cout, m_accelerator);
+
+  // dump verilog
+  ch_toVerilog("spmv.v", m_accelerator);
+}
+
+cpu_device::~cpu_device() {
+  // release allocated resources
+  if (m_matrix)
+    delete m_matrix;
+  if (m_vertex)
+    delete m_vertex;
+  if (m_shared_mem)
+    delete[] m_shared_mem;
+  if (!m_qpi_read_requests.empty()) {
+    std::cout << "CPU had " << m_qpi_read_requests.size() << " pending read requests" << std::endl;
+    for (auto rq : m_qpi_read_requests) {
+      delete rq;
+    }
+  }
+  if (!m_qpi_read_requests.empty()) {
+    std::cout << "CPU had " << m_qpi_write_requests.size() << " pending write requests" << std::endl;
+    for (auto rq : m_qpi_write_requests) {
+      delete rq;
+    }
+  }
+}
+
[…]

### 2017-11-02 cpu.h

@@ -1,0 +1,106 @@
+#pragma once
+
+#include "common.h"
+#include "graph.h"
+#include "spmv.h"
+
+namespace spmv {
+namespace host {
+
+struct rq_data_t {
+  rq_data_t() {
+    for (int i = 0; i < sizeof(__m); ++i) {
+      __m[i] = 0;
+    }
+  }
+  byte_t __m[__div_ceil(aal::aal_qpi0::cdata_width, 8)];
+};
+
+struct rq_mdata_t {
+  rq_mdata_t() {
+    for (int i = 0; i < sizeof(__m); ++i) {
+      __m[i] = 0;
+    }
+  }
+  byte_t __m[__div_ceil(aal::aal_qpi0::mdata_width, 8)];
+};
+
+struct request_t {
+  uint32_t   addr;
+  ch_tick    rq_time;
+  ch_tick    rx_time;
+  ch_tick    tx_time;
+  rq_data_t  data;
+  rq_mdata_t mdata;
+
+  request_t(uint32_t _addr, ch_tick _rq_time, ch_tick _rx_time)
+    : addr(_addr)
+    , rq_time(_rq_time)
+    , rx_time(_rx_time)
+    , tx_time(0)
+  {}
+};
+
+class cpu_device {
+public:  
+  enum {
+    SHARE_MEM_SIZE      = 0x80000000, // 2GB
+    
+    PARTITION_SIZE      = 32,
+    
+    CACHE_HIT_RATE      = 80, // 80%
+    
+    MIN_RX_RD_CYCLES    = 2,  // cache hit
+    MIN_TX_RD_CYCLES    = 2,  // cache hit
+    MAX_RX_RD_CYCLES    = 2,
+    MAX_TX_RD_CYCLES    = 32,    
+    MIN_RX_RW_CYCLES    = 2,  // cache hit
+    MIN_TX_RW_CYCLES    = 2,  // cache hit
+    MAX_RX_RW_CYCLES    = 2,
+    MAX_TX_RW_CYCLES    = 32,
+    
+    RQS_BUF_SIZE        = 90,
+    RQS_BUF_FULL_DIST   = 4,    
+  };
+  
+  cpu_device(const char* mtx_file);  
+  ~cpu_device();
+  
+  const auto& get_accelerator() const {
+    return m_accelerator;
+  }
+  
+  bool tick(ch_tick t);
+  
+  void dump_stats(ch_tick t);
+    
+private:
+  
+  void init(const char* mtx_file);
+  
[…]

### 2017-11-02 graph.cpp

@@ -1,0 +1,507 @@
+#include "graph.h"
+
+using namespace spmv;
+
+struct __attribute__((aligned(16))) edge_t {
+  uint32_t src;
+  uint32_t dst;
+  uint32_t vid;
+  uint32_t pid;
+};
+
+// sort by partition (pid), then cols (dst), then rows (src)
+static bool compare_edges_dcsc(const edge_t& a, const edge_t& b) {
+  if (a.pid < b.pid)
+    return true;
+  else if (a.pid > b.pid)
+    return false;
+  if (a.dst < b.dst)
+    return true;
+  else if (a.dst > b.dst)
+    return false;
+  if (a.src < b.src)
+    return true;
+  else if (a.src > b.src)
+    return false;
+  return false;
+}
+
+static uint32_t static_partition(uint32_t *partitions, uint32_t num_elements, uint32_t num_partitions, uint32_t alignment) {
+  partitions[0] = 0;
+  uint32_t p = 1;
+  uint32_t nnz = 1;
+  uint32_t aligned_size = __div_ceil(num_elements, alignment) * alignment;
+  uint32_t partition_size = __div_ceil(aligned_size, num_partitions);
+  partition_size = __div_ceil(partition_size, alignment) * alignment;
+  for (uint32_t n = partition_size; n < aligned_size; n += partition_size) {
+    partitions[p++] = n;
+    ++nnz;
+  }
+  do {
+    partitions[p++] = aligned_size;
+  } while (p <= num_partitions);
+  return nnz;
+}
+
+static void set_edge_pointers(uint32_t *edge_pointers,
+                              const uint32_t *row_pointers,
+                              uint32_t num_partitions,
+                              const edge_t* edges,
+                              uint32_t nnz) {
+#ifdef SET_EDGE_POINTERS_BINARY_SEARCH
+#pragma omp parallel for num_threads(nthreads)
+  for (uint32_t p = 0; p < num_partitions; ++p) {
+    // binary search
+    uint32_t e1 = 0;
+    uint32_t e2 = nnz;
+    uint32_t eh;
+    while (e2 >= e1) {
+      eh = e2 - (e2 - e1) / 2;
+      if (eh == 0) {
+        break;
+      }
+      if ((edges[eh - 1].src < row_pointers[p]) &&
+          edges[eh].src >= row_pointers[p]) {
+        break;
[…]

### 2017-11-02 graph.h

@@ -1,0 +1,64 @@
+#pragma once
+
+#include "common.h"
+#include "mtx.h"
+
+namespace spmv {
+
+struct mdcsc_t { 
+public:
+  uint32_t  cols;
+  uint32_t  rows;
+  uint32_t  nnz;
+  uint32_t  nzx;
+  uint32_t  numparts;
+  uint32_t  element_size;
+  uint32_t* partitions;
+  uint32_t* xindices;
+  uint32_t* ystarts;
+  uint32_t* yindices;  
+  byte_t*   values;    
+  
+  mdcsc_t();
+  mdcsc_t(
+    uint32_t cols,
+    uint32_t rows,
+    uint32_t nnz,
+    uint32_t nzx,
+    uint32_t numparts,
+    uint32_t element_size) {
+    this->init(cols, rows, nnz, nnz, numparts, element_size);  
+  }
+  
+  mdcsc_t(const char* mtx_file, mtx_format edge_type, uint32_t partition_size);
+  ~mdcsc_t();
+  
+  
+  void init(uint32_t cols,
+            uint32_t rows,
+            uint32_t nnz,
+            uint32_t nzx,
+            uint32_t numparts,
+            uint32_t element_size);
+  
+  void allocate();
+  
+  size_t copy(byte_t* dest, size_t offset, size_t size, ch_port_t<ch_matrix_dcsc_t>& desc);
+};
+
+struct vertex_t {
+  byte_t*   values;
+  uint32_t* masks;
+  uint32_t  size;
+  uint32_t  element_size;
+  
+  vertex_t();
+  vertex_t(uint32_t size, uint32_t element_size);
+  ~vertex_t();
+  
+  size_t copy(byte_t* dest, size_t offset, size_t size, ch_port_t<ch_vertex_t>& desc);
+  
+  void dump();
+};
+
+}

### 2017-11-02 lsu.cpp

@@ -1,0 +1,773 @@
+#include "lsu.h"
+#include "spmv.h"
+
+using namespace spmv;
+using namespace spmv::accelerator;
+
+spmv_lsu::spmv_lsu()
+  : m_writemask0_addr(PTR_MAX_VALUE)
+  , m_writemask1_addr(PTR_MAX_VALUE) {
+  // bind nested modules
+  ctrl_.pbuf_module.io(ctrl_.pbuf_io);
+  for (int i = 0; i < PE_COUNT; ++i) {
+    pe_[i].axbuf_module.io(pe_[i].axbuf_io);
+    pe_[i].asbuf_module.io(pe_[i].asbuf_io);
+    pe_[i].aybuf_module.io(pe_[i].aybuf_io);
+    pe_[i].avbuf_module.io(pe_[i].avbuf_io);
+    pe_[i].xvbuf_module.io(pe_[i].xvbuf_io);
+    pe_[i].xmbuf_module.io(pe_[i].xmbuf_io);
+  }
+}
+
+spmv_lsu::~spmv_lsu() {
+  //--
+}
+
+void spmv_lsu::describe() {
+  this->read_req_thread();
+  this->write_req_thread();
+  this->read_rsp_thread();
+  this->write_rsp0_thread();
+  this->write_rsp1_thread();
+  this->writemask_thread();
+
+  //
+  // return outputs
+  //
+
+}
+
+ch_bit1 spmv_lsu::check_pe_buf_ready(uint32_t pe_id, const ch_rd_request& rq_type) {
+  ch_bit1 ret;
+  __switch (rq_type) (
+  __case (ch_rd_request::a_xindices) (
+    ret = (pe_[pe_id].axbuf_pending_size != AXBUF_SIZE);
+  )
+  __case (ch_rd_request::a_ystarts) (
+    ret = (pe_[pe_id].asbuf_pending_size != ASBUF_SIZE);
+  )
+  __case (ch_rd_request::a_yindices) (
+    ret = (pe_[pe_id].aybuf_pending_size != AYBUF_SIZE);
+  )
+  __case (ch_rd_request::a_values) (
+    ret = (pe_[pe_id].avbuf_pending_size != AVBUF_SIZE);
+  )
+  __case (ch_rd_request::x_values) (
+    ret = (pe_[pe_id].xvbuf_pending_size != XVBUF_SIZE);
+  )
+  __case (ch_rd_request::x_masks) (
[…]

### 2017-11-02 lsu.h

@@ -1,0 +1,303 @@
+#pragma once
+
+#include "common.h"
+
+namespace spmv {
+namespace accelerator {
+
+#define INT32_TO_BLOCK_ADDR(x) \
+   (((x) << 2) >> LOG2_BLOCK_SIZE)
+
+#define CEIL_INT32_TO_BLOCK_ADDR(x) \
+   ((((x) << 2) + BLOCK_SIZE_MASK) >> LOG2_BLOCK_SIZE)
+
+#define INT32_TO_BLOCK_BITSHIFT(x) \
+  (((x) & 0xf) << 5)
+
+__enum (ch_rd_request, 3, (
+  a_partition,
+  a_xindices,
+  a_ystarts,
+  a_yindices,
+  a_values,
+  x_values,
+  x_masks
+));
+
+__enum (ch_wr_request, 2, (
+  y_values,
+  y_masks,
+  hwcntrs
+));
+
+__enum (ch_rq_owner, 2, (
+  PE0,
+  PE1,             
+  ctrl
+));
+
+enum {
+  LOG2_PBUF_SIZE = 2,
+  PBUF_SIZE = 1 << LOG2_PBUF_SIZE,
+
+  LOG2_AXBUF_SIZE = 1,
+  AXBUF_SIZE = 1 << LOG2_AXBUF_SIZE,
+
+  LOG2_ASBUF_SIZE = 1,
+  ASBUF_SIZE = 1 << LOG2_ASBUF_SIZE,
+
+  LOG2_AYBUF_SIZE = 5,
+  AYBUF_SIZE = 1 << LOG2_AYBUF_SIZE,
+
+  LOG2_AVBUF_SIZE = 5,
+  AVBUF_SIZE = 1 << LOG2_AVBUF_SIZE,
+
+  LOG2_XVBUF_SIZE = 1,
+  XVBUF_SIZE = 1 << LOG2_XVBUF_SIZE,
+
+  LOG2_XMBUF_SIZE = 1,
+  XMBUF_SIZE = 1 << LOG2_XMBUF_SIZE,
+
+  LOG2_ABUF_SIZE = 3,
+  ABUF_SIZE = 1 << LOG2_ABUF_SIZE,
+};
+
+struct ch_rd_mdata_t {
+  using value_type = ch_bit<aal::aal_qpi0::mdata_width>;
+  ch_rd_request type;
+  ch_rq_owner owner;
+
[…]

### 2017-11-02 main.cpp

@@ -1,0 +1,88 @@
+#include <iostream>
+#include <fstream>
+#include <stdarg.h>
+#include <unistd.h>
+#include "simulator.h"
+
+using namespace spmv;
+
+#define CPU_FREQ (2.3e9)
+
+int verbose = 0;
+int run_ticks = 10000000;
+const char* mtx_file = "sample.mtx";
+
+namespace spmv {
+bool g_singlecore = false; 
+};
+
+void spmv::DbgPrint(int level, const char *format, ...) {
+  char buf[256];
+  va_list args;
+  va_start(args, format);
+  if (level <= verbose) {
+    vsnprintf(buf, 256, format, args);
+    std::cout << buf;
+  }
+  va_end(args);
+}
+
+static inline uint64_t __rdtsc() {
+  uint32_t hi, lo;
+  __asm__ volatile("rdtsc" : "=a" (lo), "=d" (hi));
+  return ((uint64_t)hi << 32) | lo;
+}
+
+static double elapsed_time(uint64_t start, uint64_t end) {
+  return (end - start) / CPU_FREQ;
+}
+
+static void parse_args(int argc, char **argv) {
+  int c;
+  while ((c = getopt(argc, argv, "sm:v:t:i:h?")) != -1) {
+    switch (c) {
+    case 'v':
+      verbose = atoi(optarg);
+      break;
+    case 's':
+      spmv::g_singlecore = true;
+      break;
+    case 'm':
+      mtx_file = optarg;
+      break;
+    case 't':
+      run_ticks = atoi(optarg);
+      break;
+    case '?':
+    case 'h':
+      printf("SPMV RTL Simulator v1.0 - by Blaise Tine.\n");
+      printf("Usage: [-m mtx matrix] [-v verbose_level] [-t run_ticks] [-h help]\n");
+    default:
+      exit(1);
+    }
+  }
+  
+  std::ifstream ifile(mtx_file);
+  if (!ifile) {
+    printf("** couldn't open matrix file: %s", mtx_file);
+    exit(1);
+  }    
+}
+
+int main(int argc, char **argv) {  
+  //--  
+  parse_args(argc, argv);
+  
+  //--
+  simulator* sim = new simulator(mtx_file);
+  uint64_t start = __rdtsc();
+  
+  sim->run(run_ticks);
+  
+  uint64_t end = __rdtsc();
+  printf("Total elapsed time = %.3f s\n", elapsed_time(start, end));
+
+  delete sim;
+
+  return 0;
+}

### 2017-11-02 mtx.cpp

@@ -1,0 +1,280 @@
+#include <iostream>
+#include <sstream>
+#include <assert.h>
+#include "mtx.h"
+
+static bool is_space_or_comment(const std::string& line) {
+  const char * str = line.c_str();
+  if (*str == '#')
+    return true;
+  while (std::isspace(*str))
+    ++str;
+  return (*str == 0);
+}
+
+template <typename T>
+int readline(std::ifstream& file, mtx_edge_t* edges, T* weights) {
+  std::string line;
+  while (getline(file, line)) {
+    if (is_space_or_comment(line))
+      continue;
+    std::istringstream ln_ss(line);
+    ln_ss >> edges->src >> edges->dst >> *weights++;
+    if (ln_ss.fail())
+      return -1;
+    ++edges;
+  }
+  return 0;
+}
+
+template <typename T>
+void writeline(std::ofstream& file, const mtx_edge_t* edges, const T* weights, uint32_t nnzs) {
+  for (uint32_t i = 0; i < nnzs; ++i) {
+    file << edges[i].src << edges[i].dst << weights[i] << std::endl;
+  }
+}
+
+size_t mxt_format_size(mtx_format fmt) {
+  switch (fmt) {
+  case mtx_format::None:
+    return 0;
+  case mtx_format::Int32:
+    return sizeof(int);
+  case mtx_format::Float:
+    return sizeof(float);
+  case mtx_format::Int64:
+    return sizeof(int64_t);
+  case mtx_format::Double:
+    return sizeof(double);
+  }
+}
+
+void mtx_format_cast(void* dst, mtx_format fmt, double value) {
+  switch (fmt) {
+  case mtx_format::Int32:
+    *(int*)dst = value;
+    break;
+  case mtx_format::Float:
+    *(float*)dst = value;
+    break;
+  case mtx_format::Int64:
+    *(int64_t*)dst = value;
+    break;
+  case mtx_format::Double:
+    *(double*)dst = value;
+    break;
+  }
+}
+
+bool mtx_is_binary(std::ifstream& file) {
+  char magic[4];
+
+  file.read((char*)magic, sizeof(magic));
+  if (!file)
+    return -1;
+
+  file.seekg(0, file.beg);
+  return magic[0] == 'M'
+      && magic[1] == 'T'
+      && magic[2] == 'X'
+      && magic[3] == 0;
+}
+
[…]

### 2017-11-02 mtx.h

@@ -1,0 +1,52 @@
+#pragma once
+
+#include <fstream>
+#include <cstdlib>
+
+enum class mtx_format : uint32_t {
+  None,
+  Int32,
+  Float,
+  Int64,
+  Double,
+};
+
+struct  mtx_header_t {
+  mtx_format fmt;
+  uint32_t   rows;
+  uint32_t   cols;
+  uint32_t   nnz;
+};
+
+struct  mtx_edge_t {
+  uint32_t src;
+  uint32_t dst;
+};
+
+size_t mxt_format_size(mtx_format fmt);
+
+void mtx_format_cast(void* dst, mtx_format fmt, double value);
+
+bool mtx_is_binary(std::ifstream& file);
+
+int mtx_read_header_from_binary(std::ifstream& file, mtx_header_t* header);
+
+int mtx_read_data_from_binary(std::ifstream& file, const mtx_header_t& header,
+                              mtx_edge_t* edges, int edges_size,
+                              void* weights, int weights_size);
+
+int mtx_read_header_from_text(std::ifstream& file, mtx_header_t* header);
+
+int mtx_read_data_from_text(std::ifstream& file, const mtx_header_t& header,
+                            mtx_edge_t* edges, int edges_size,
+                            void* weights, int weights_size);
+
+int mtx_detect_header_from_text(std::ifstream& file, mtx_header_t* header);
+
+int mtx_write_binary(std::ofstream& file, const mtx_header_t& header,
+                     const mtx_edge_t* edges, int edges_size,
+                     const void* weights, int weights_size);
+
+int mtx_write_text(std::ofstream& file, const mtx_header_t& header,
+                   const mtx_edge_t* edges, int edges_size,
+                   const void* weights, int weights_size);

### 2017-11-02 pe.cpp

@@ -1,0 +1,473 @@
+#include "pe.h"
+#include "spmv.h"
+
+using namespace spmv;
+using namespace spmv::accelerator;
+
+__enum (ch_exec_state, 4, (
+  get_partition,
+  clear_output,
+  get_a_xindex,
+  check_x_mask,             
+  get_x_mask,
+  get_a_ystart,
+  get_a_yend,
+  get_x_value,
+  get_a_yindex,
+  get_a_value,
+  wait_for_data,
+  execute,
+  next_column,
+  write_y_value0,
+  write_y_value1,
+  write_y_mask
+));
+
+spmv_pe::spmv_pe(uint32_t id)
+  : id_(id)
+  , xmblock_addr_(PTR_MAX_VALUE) // default initialize to max
+  , xvblock_addr_(PTR_MAX_VALUE) // default initialize to max
+{}
+
+spmv_pe::~spmv_pe() {
+  //--
+}
+
+void spmv_pe::describe() {
+  //--
+  ch_seq<ch_exec_state> state;
+  
+  //--
+  ch_ram<ch_bit32, PARTITION_SIZE> y_values;
+  __if (y_wenable_) (
+    y_values[y_addr_] = y_value_;
+  );
+  
+  ch_block_t y_value0;
+  for (int i = 0; i < 16; ++i) {
+    y_value0.slice<32>(i*32) = y_values[i];
+  }
+  
+  ch_block_t y_value1;
+  for (int i = 0; i < 16; ++i) {
+    y_value1.slice<32>(i*32) = y_values[i+16];
+  }
+  
+  //--
+  /*if (id_ == 0) {
+    ch_print(fstring("{0}: PE[%d]: state={1}, rq_ack={2}, rq_typ={3}, rq_adr={4}, col={5}, coln={6}, ax={7}, xv={8}, row={9}, rown={10}, ya={11}, yv={12}, ym={13}, wq_ack={14}, wq_typ={15}, wq_adr={16}", id_).c_str(),
+           ch_getTick(), state, lsu_.rd_req_syn, lsu_.rd_req_type, lsu_.rd_req_addr,
+           col_curr_, col_end_, a_xindex_, x_value_, row_curr_, row_end_, y_addr_, y_value_, y_mask_,
+           lsu_.wr_req_syn, lsu_.wr_req_type, lsu_.wr_req_addr);
+  }*/
+
+  // ch_exec_state FSM
+  __switch (state) (
+  __case (ch_exec_state::get_partition) (
[…]

### 2017-11-02 pe.h

@@ -1,0 +1,104 @@
+#pragma once
+
+#include "common.h"
+#include "lsu.h"
+
+namespace spmv {
+namespace accelerator {
+
+__inout (ch_pe_start_req_io_t, (
+  __in(ch_bool)        syn,
+  __in(ch_dcsc_part_t) part,
+  __out(ch_bool)       ack
+));
+
+__inout (ch_ctrl_pe_io_t, (
+  (ch_pe_start_req_io_t) start_req,
+  __in(ch_bit64) timer,
+  __out(ch_pe_hwcntrs_t) hwcntrs
+));
+
+class spmv_pe {
+public:
+
+  __io(
+    (ch_ctrl_pe_io_t) ctrl,
+    (ch_flip_t<ch_pe_lsu_io_t>) lsu
+  );
+  
+  spmv_pe(uint32_t id);
+  ~spmv_pe();
+  
+  void describe();
+
+private:
+
+  using ch_bit480 = ch_bit<480>;
+
+  struct ctrl_t {
+    ch_seq<ch_bit1> start_req_ack;
+
+    ctrl_t() : start_req_ack(1) // default initialize to '1'
+    {}
+  };
+
+  struct lsu_t {
+    ch_seq<ch_bit1>       rd_req_syn;
+    ch_seq<ch_rd_request> rd_req_type;
+    ch_seq<ch_blk_addr>   rd_req_addr;
+
+    ch_seq<ch_bit1>       wr_req_syn;
+    ch_seq<ch_wr_request> wr_req_type;
+    ch_seq<ch_blk_addr>   wr_req_addr;
+    ch_seq<ch_block_t>    wr_req_data;
+
+    ch_bit1 axbuf_dequeue;
+    ch_bit1 asbuf_dequeue;
+    ch_bit1 aybuf_dequeue;
+    ch_bit1 avbuf_dequeue;
+    ch_bit1 xvbuf_dequeue;
+    ch_bit1 xmbuf_dequeue;
+  };
+  
+  uint32_t  id_;
+
+  ctrl_t ctrl_;
+
+  lsu_t lsu_;
+  
+  ch_seq<ch_ptr>      col_curr_;
+  ch_seq<ch_ptr>      col_end_;
+
+  ch_seq<ch_block_t>  axblock_;
+  ch_seq<ch_block_t>  asblock_;
+  ch_seq<ch_block_t>  ayblock_;
[…]

### 2017-11-02 report.py

@@ -1,0 +1,66 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import sys
+import os
+import subprocess
+import multiprocessing
+import re
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+
+def GetStats(logdir, injection_rate):
+	logfile = logdir + '/result_' + str(injection_rate) + '.log'
+	
+	with open(logfile, 'r') as f:
+		lines = f.readlines()		
+	
+	for (i, line) in enumerate(lines):
+		if 'average packet latency' in line:
+			match = re.search('average packet latency *= *([+-]?(\d+(\.\d*)?|\.\d+)([eE][+-]?\d+)?)', line)
+			latency = float(match.group(1))
+			break
+		
+	print('injection_rate=%f, latency=%f' % (injection_rate, latency))
+	return latency
+
+def Plot(logdir, injection_rates, outputfile) :
+	_colors = ['r', 'b', 'g', 'c', 'm', 'y', 'k', 'w']
+	
+	latencies = []	
+	for injection_rate in injection_rates:
+		latencies.append(GetStats(logdir, injection_rate))
+		
+	fig = plt.figure()
+	plt.title('latency vs throughput')
+	plt.xlabel('injection-rate (flit/node/cycle)')
+    plt.ylabel('average flit latency (ticks)')
+	plt.plot(injection_rates, latencies)   
+	plt.grid()
+	fig.savefig(outputfile)
+
+def Plot2(logdirs, labels, injection_rates, outputfile) :
+	_colors = ['r', 'b', 'g', 'c', 'm', 'y', 'k', 'w']
+	
+	fig = plt.figure()
+	
+	for idx in xrange(len(logdirs)):
+		latencies = []	
+		for injection_rate in injection_rates:
+			latencies.append(GetStats(logdirs[idx], injection_rate))	
+		plt.plot(injection_rates, latencies, color=_colors[idx], label=labels[idx])        
+    	
+	plt.title('latency vs throughput')
+	plt.xlabel('injection-rate (flit/node/cycle)')
+    plt.ylabel('average flit latency (ticks)')
+	plt.plot(injection_rates, latencies)   
+	plt.grid()
+	plt.legend()	
+	fig.savefig(outputfile)
+	
+Plot('results_x', (np.arange(16)+1)/16.0, 'resuls_x.png')
[…]

### 2017-11-02 runtest.py

@@ -1,0 +1,54 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import sys
+import os
+import subprocess
+import multiprocessing
+import re
+import math
+import numpy as np
+
+MAX_THREADS = 2
+
+def RunTestOnce(params):
+	(mtxfile, logfile) = params
+	# build the command line
+	cmd = './spmv.out -m %s > %s' % (mtxfile, logfile)
+	print(cmd)
+	# spawn the process in pipe mode
+	process = subprocess.Popen(cmd, shell=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+	output = process.communicate()[0];
+	exitcode = process.returncode   # get return value
+	if exitcode != 0:
+		print('usimm failed with exitcode %d' % exitcode)
+		print(output)		
+	return exitcode
+
+def RunTest(matrices, log_dir):	
+	num_samples = len(matrices)
+	print('number of samples = %d' % num_samples)
+	
+	# create processors pool 
+	num_procs = min(multiprocessing.cpu_count(), MAX_THREADS)
+	print('number of processors = %d' % num_procs)
+	procs_pool = multiprocessing.Pool(processes=num_procs)
+	
+	wIdx = 0
+	done = False
+	while not done:			
+		# fill up the input list
+		input_args = []
+		for p in range(0, num_procs):
+			matrix = matrices[wIdx]
+			logfile = log_dir + '/' + matrix + '.log'
+			input_args.append((matrix, logfile))
+			wIdx += 1
+			if wIdx == num_samples:
+				done = True
+				break
+		
+		# execute the job batch
+		procs_pool.map(func=RunTestOnce, iterable=input_args)
+
+RunTest(['graph500_6.mtx','graph500_8.mtx','graph500_10.mtx','graph500_12.mtx', 'graph500_14.mtx', 'graph500_16.mtx', 'graph500_18.mtx', 'graph500_20.mtx'], 'results')

### 2017-11-02 simulator.cpp

@@ -1,0 +1,65 @@
+#include "simulator.h"
+
+using namespace spmv;
+
+#define CPU_FREQ (2.3e9)
+
+static uint64_t __rdtsc() {
+  unsigned int hi, lo;
+  __asm__ volatile("rdtsc" : "=a"(lo), "=d"(hi));
+  return ((uint64_t)hi << 32) | lo;
+}
+
+static double elapsed_time(uint64_t start, uint64_t end) {
+  return (end - start) / CPU_FREQ;
+}
+
+simulator::simulator(const char* mtx_file)
+  : m_vcd_file("spmv.vcd") {
+  // initialize random seed
+  srand(0);
+    
+  //--
+  m_cpu = new host::cpu_device(mtx_file);  
+  m_tracer = new ch_vcdtracer(m_vcd_file, m_cpu->get_accelerator());
+}
+
+simulator::~simulator() {   
+  // release resources
+  delete m_tracer;
+  delete m_cpu;
+}
+
+void simulator::run(ch_tick run_ticks) {
+  // reset native simulator
+  ch_tick t = m_tracer->reset(0);
+
+  uint64_t start = __rdtsc();
+
+  // run simulation  
+  for (;t < run_ticks;) {
+    
+    // tick the cpu device
+    if (!m_cpu->tick(t))
+      break;
+    
+    // advance native simulation
+    t = m_tracer->step(t);
+  }
+
+  uint64_t end = __rdtsc();
+  printf("Total elapsed time = %.3f s\n", elapsed_time(start, end));
+  
+  // dump stats
+  this->dump_stats(t);
+}
+
+void simulator::trace(ch_tick t) {
+  //--
+  DbgPrint(2, "%06ld:\n", t);
+}
+
+void simulator::dump_stats(ch_tick t) {
+  //--
+  m_cpu->dump_stats(t);
+}

### 2017-11-02 simulator.h

@@ -1,0 +1,26 @@
+#pragma once
+
+#include "cpu.h"
+#include "spmv.h"
+
+namespace spmv {
+
+class simulator {
+public:
+  simulator(const char* mtx_file);
+  ~simulator();    
+  
+  void run(ch_tick run_ticks);
+   
+protected:
+  
+  void trace(ch_tick t);
+  
+  void dump_stats(ch_tick t);
+  
+  host::cpu_device* m_cpu;
+  ch_vcdtracer* m_tracer;
+  std::ofstream m_vcd_file;
+};
+
+}

### 2017-11-02 spmv.cpp

@@ -1,0 +1,270 @@
+#include "spmv.h"
+
+using namespace spmv;
+using namespace spmv::accelerator;
+
+__enum (ch_ctrl_state, 3, (
+  ready,            
+  get_partition,
+  execute,
+  flush_buffers,
+  write_hwcntrs
+));
+
+__enum (ch_dispatch_state, 2, (
+  get_partition,
+  dispatch_PE0,
+  dispatch_PE1
+));
+
+spmv_device::spmv_device() {
+  // create the PEs
+  for (int i = 0; i < PE_COUNT; ++i) {
+    pe_[i].module = std::make_unique<ch_module<spmv_pe>>(i);
+  }
+  // create LSU
+  lsu_.module = std::make_unique<ch_module<spmv_lsu>>();
+
+  // bind PEs
+  for (int i = 0; i < PE_COUNT; ++i) {
+    pe_[i].module->io.ctrl(pe_[i].io);
+    lsu_.module->io.pe[i](pe_[i].module->io.lsu);
+  }
+
+  // bind LSU
+  lsu_.module->io.ctx(io.ctx);
+  lsu_.module->io.qpi(io.qpi);
+  lsu_.module->io.ctrl(lsu_.io);
+}
+
+spmv_device::~spmv_device() {
+  //--
+}
+
+void spmv_device::describe() {
+  // print the context block
+  ch_print("{0}: ctx.a.numparts={1}, ctx.a.partitions={2}",
+              ch_getTick(), io.ctx.a.numparts, io.ctx.a.partitions);
+
+  // emit timer
+  timer_.next = timer_ + 1;
+  
+  // emit main thread
+  this->main_thread();
+    
+  // emit dispatcher
+  this->dispatch_thread();
+
+  //
+  // return outputs
+  //
+
+  for (int i = 0; i < PE_COUNT; ++i) {
+    pe_[i].io.start_req.syn = pe_[i].start_req_syn;
+  }
+
+  lsu_.io.rd_req.syn  = lsu_.rd_req_syn;
+  lsu_.io.rd_req.type = lsu_.rd_req_type;
+  lsu_.io.rd_req.addr = lsu_.rd_req_addr;
+  lsu_.io.wr_req.syn  = lsu_.wr_req_syn;
+  lsu_.io.wr_req.type = lsu_.wr_req_type;
+  lsu_.io.wr_req.addr = lsu_.wr_req_addr;
[…]

### 2017-11-02 spmv.h

@@ -1,0 +1,67 @@
+#pragma once
+
+#include "common.h"
+#include "lsu.h"
+#include "pe.h"
+
+namespace spmv {
+namespace accelerator {
+
+using ch_part_buf_t = ch_bit<ch_bitwidth_v<ch_block_t> + PARTITION_VALUE_BITS>;
+
+class spmv_device : public spmv_aal_device {
+public:
+  spmv_device();
+  ~spmv_device();
+
+  void describe() override;
+
+protected:
+
+  void main_thread();
+  void dispatch_thread();
+
+  struct pe_t {
+    std::unique_ptr<ch_module<spmv_pe>> module;
+    ch_flip_t<ch_ctrl_pe_io_t> io;
+    ch_seq<ch_dcsc_part_t> parts;
+    ch_seq<ch_bit1> start_req_syn;
+  };
+
+  struct lsu_t {
+    std::unique_ptr<ch_module<spmv_lsu>> module;
+    ch_flip_t<ch_ctrl_lsu_io_t> io;
+
+    ch_seq<ch_bit1> rd_req_syn;
+    ch_seq<ch_rd_request> rd_req_type;
+    ch_blk_addr rd_req_addr;
+
+    ch_seq<ch_bit1> wr_req_syn;
+    ch_seq<ch_wr_request> wr_req_type;
+    ch_blk_addr wr_req_addr;
+    ch_block_t wr_req_data;
+
+    ch_bit1 pbuf_dequeue;
+  };
+
+  pe_t pe_[PE_COUNT];
+
+  lsu_t lsu_;
+
+  ch_seq<ch_part_buf_t> part_buf_;
+  ch_seq<ch_bit<LOG2_PARTITIONS_PER_BLOCK>> part_buf_size_;
+
+  ch_seq<ch_ptr> part_blk_curr_;
+  ch_seq<ch_ptr> part_blk_end_;
+  ch_seq<ch_ptr> part_curr_;
+  ch_ptr numparts_;
+
+  ch_seq<ch_bit64> timer_;
+
+  ch_seq<ch_hwcntrs_t> hwcntrs_;
+
+  ch_seq<ch_bit1> done_;
+};
+
+}
+}

### 2017-11-02 test.h

@@ -1,0 +1,4 @@
+#pragma once
+
+#include "common.h"
+#include "catch.h"

### 2017-11-02 test_lsu.cpp

@@ -1,0 +1,12 @@
+#include <iostream>
+#include <fstream>
+#include <unistd.h>
+#include "test.h"
+
+using namespace spmv;
+
+
+int main(int argc, char **argv) {
+  //--
+  return 0;
+}

### 2017-11-02 tmp.h

@@ -1,0 +1,24 @@
+/*{
+  // spmv_device FSM
+  ch_seq<ch_ctrl_state> state;
+     
+  __switch (state) (
+  __case (ch_ctrl_state::ready) (
+    __if (m_start) (
+      state.next = ch_ctrl_state::execute;
+    );
+  )
+  __case (ch_ctrl_state::execute) (
+    __if (m_PE.complete()) (
+      state.next = ch_ctrl_state::write_hwcntrs;
+    );
+  )
+  __case (ch_ctrl_state::write_hwcntrs) (
+    __if (m_LSU.done()) (
+      state.next = ch_ctrl_state::ready;
+    );
+  )
+  __default (
+    state.next = state;        
+  ));
+}*/

### 2017-11-02 vcache.h

@@ -1,0 +1,93 @@
+#pragma once
+
+#include "common.h"
+
+namespace spmv {
+namespace accelerator {
+
+//
+// Description:
+// this code implements a vertex cache, entries are insert and evicted in fifo order.  
+//
+
+template <unsigned SIZE, unsigned DATA_WIDTH, unsigned TAG_WIDTH>
+class vertex_cache {
+public:
+  
+  enum {
+    address_width = log2ceil(SIZE),
+  };  
+  
+  using ch_addr_t = ch_bit<address_width>;
+  using ch_tag_t  = ch_bit<TAG_WIDTH>;
+  using ch_data_t = ch_bit<DATA_WIDTH>;  
+  
+  void emit() {
+    // emit tag valid test
+    ch_bit1 tag_valid[SIZE];
+    ch_addr_t match_pos[SIZE];
+    tag_valid[0] = (m_query_tag == (ch_tag_t)m_tags[0]);
+    match_pos[0] = 0;
+    for (int i = 1; i < SIZE; ++i) {
+      tag_valid[i] = tag_valid[i-1] | (m_query_tag == (ch_tag_t)m_tags[i-1]);
+      match_pos[i] = ch_select<address_width>((m_query_tag == (ch_tag_t)m_tags[i]), i, match_pos[i-1]);
+    }
+    m_query_valid.next = tag_valid[SIZE-1];
+    m_query_pos.next = match_pos[SIZE-1];        
+    
+    // emit block insersion
+    __if (m_enq_enable) (
+      m_buf[m_pos] = m_enq_data;
+      m_tags[m_pos] = m_enq_tag;  
+      m_pos.next = m_pos + 1; // advance position
+    )__else (
+      m_pos.next = m_pos; 
+    );    
+      
+    // match data
+    m_query_data = m_buf[m_query_pos];
+  }
+  
+  void enq_tag(const ch_tag_t& tag) {
+    m_enq_tag = tag;
+  }
+  
+  void enq_valid(const ch_bit1& enable) {
+    m_enq_enable = enable;
+  }
+  
+  void enq_data(const ch_data_t& data) {
+    m_enq_data = data;
+  }
+  
+  void query_tag(const ch_tag_t& tag) {
+    m_query_tag = tag;
+  }
+  
+  const ch_bit1& get_query_valid(const ch_tag_t& tag) const {
[…]

### 2017-11-02 cpu.cpp
first-pass refactoring - buggy :-(

@@ -77,7 +77,7 @@ void cpu_device::init(const char* mtx_file) {
   offset = m_matrix->copy(m_shared_mem, offset, SHARE_MEM_SIZE, m_accelerator.io.ctx.a);
   offset = m_vertex->copy(m_shared_mem, offset, SHARE_MEM_SIZE, m_accelerator.io.ctx.x);
   offset = m_vertex->copy(nullptr, offset, SHARE_MEM_SIZE, m_accelerator.io.ctx.y);
-  ch_poke(m_accelerator.io.ctx.hwcntrs, (offset / BLOCK_SIZE));
+  m_accelerator.io.ctx.hwcntrs = (offset / BLOCK_SIZE);
   assert(__align(offset + __div_ceil(ch_bitwidth_v<ch_hwcntrs_t>, 8), BLOCK_SIZE) <= SHARE_MEM_SIZE);
   
   /*printf("*** partitions=0x%x, a_xindices=0x%x, a_ystarts=0x%x, a_yindices=0x%x, a_values=0x%x, x_values=x0%x, x_masks=0x%x\n",
@@ -172,8 +172,8 @@ void cpu_device::check_output(const float* values, const uint32_t* masks, bool d
 
 bool cpu_device::tick(ch_tick t) {
   // check done signal
-  if (ch_peek<bool>(m_accelerator.io.done)) {
-    ch_poke(m_accelerator.io.start, false); // stop execution
+  if (m_accelerator.io.done) {
+    m_accelerator.io.start = false; // stop execution
     this->check_result(t);
     return false;
   }  
@@ -190,49 +190,49 @@ bool cpu_device::tick(ch_tick t) {
     this->process_wr_rsp(t);
   
   // process read requests
-  if (ch_peek<bool>(m_accelerator.io.qpi.rd_req.valid))
+  if (m_accelerator.io.qpi.rd_req.valid)
     this->process_rd_req(t);
   
   // process write requests
-  if (ch_peek<bool>(m_accelerator.io.qpi.wr_req.valid))
+  if (m_accelerator.io.qpi.wr_req.valid)
     this->process_wr_req(t);
   
   // start execution
   if (4 == t) {
-    ch_poke(m_accelerator.io.start, true);
+    m_accelerator.io.start = true;
   }
   
[…]

### 2017-11-02 cpu.h

@@ -90,7 +90,7 @@ private:
   
   void check_output(const float* values, const uint32_t* masks, bool debug);
   
-  ch_module<accelerator::spmv_device> m_accelerator;
+  ch_device<accelerator::spmv_device> m_accelerator;
 
   std::list<request_t*> m_qpi_read_requests;
   std::list<request_t*> m_qpi_write_requests;

### 2017-11-02 graph.cpp

@@ -431,14 +431,14 @@ void mdcsc_t::init(
   values     = new byte_t[nnz_ * element_size_];  
 }
 
-size_t mdcsc_t::copy(byte_t* dest, size_t offset, size_t size, ch_port_t<ch_matrix_dcsc_t>& desc) {
-  ch_poke(desc.numparts, numparts);
+size_t mdcsc_t::copy(byte_t* dest, size_t offset, size_t size, ch_scalar_t<ch_matrix_dcsc_t>& desc) {
+  desc.numparts = numparts;
   
 #define  __copy_data(s, S) \
   assert((offset & BLOCK_SIZE_MASK) == 0); \
   assert(offset + __align(S, BLOCK_SIZE) <= size); \
   assert((offset >> LOG2_BLOCK_SIZE) < 0x100000); \
-  ch_poke(desc.s, (offset >> LOG2_BLOCK_SIZE)); \
+  desc.s = (offset >> LOG2_BLOCK_SIZE); \
   if (dest) memcpy(dest + offset, s, S); \
   offset += __align(S, BLOCK_SIZE);  
   
@@ -474,13 +474,13 @@ vertex_t::~vertex_t() {
   delete[] masks;
 }
 
-size_t vertex_t::copy(byte_t* dest, size_t offset, size_t size, ch_port_t<ch_vertex_t>& desc) {
+size_t vertex_t::copy(byte_t* dest, size_t offset, size_t size, ch_scalar_t<ch_vertex_t>& desc) {
 
 #define  __copy_data(s, S) \
   assert((offset & BLOCK_SIZE_MASK) == 0); \
   assert(offset + __align(S, BLOCK_SIZE) <= size); \
   assert((offset >> LOG2_BLOCK_SIZE) < 0x100000); \
-  ch_poke(desc.s, (offset >> LOG2_BLOCK_SIZE)); \
+  desc.s = (offset >> LOG2_BLOCK_SIZE); \
   if (dest) memcpy(dest + offset, s, S); \
   offset += __align(S, BLOCK_SIZE);  
   

### 2017-11-02 graph.h

@@ -43,7 +43,7 @@ public:
   
   void allocate();
   
-  size_t copy(byte_t* dest, size_t offset, size_t size, ch_port_t<ch_matrix_dcsc_t>& desc);
+  size_t copy(byte_t* dest, size_t offset, size_t size, ch_scalar_t<ch_matrix_dcsc_t>& desc);
 };
 
 struct vertex_t {
@@ -56,7 +56,7 @@ struct vertex_t {
   vertex_t(uint32_t size, uint32_t element_size);
   ~vertex_t();
   
-  size_t copy(byte_t* dest, size_t offset, size_t size, ch_port_t<ch_vertex_t>& desc);
+  size_t copy(byte_t* dest, size_t offset, size_t size, ch_scalar_t<ch_vertex_t>& desc);
   
   void dump();
 };

### 2017-11-02 lsu.cpp

@@ -7,34 +7,50 @@ using namespace spmv::accelerator;
 spmv_lsu::spmv_lsu()
   : m_writemask0_addr(PTR_MAX_VALUE)
   , m_writemask1_addr(PTR_MAX_VALUE) {
-  // bind nested modules
-  ctrl_.pbuf_module.io(ctrl_.pbuf_io);
-  for (int i = 0; i < PE_COUNT; ++i) {
-    pe_[i].axbuf_module.io(pe_[i].axbuf_io);
-    pe_[i].asbuf_module.io(pe_[i].asbuf_io);
-    pe_[i].aybuf_module.io(pe_[i].aybuf_io);
-    pe_[i].avbuf_module.io(pe_[i].avbuf_io);
-    pe_[i].xvbuf_module.io(pe_[i].xvbuf_io);
-    pe_[i].xmbuf_module.io(pe_[i].xmbuf_io);
+  //
+  // io binding
+  //
+  io.ctrl.rd_req.ack = ctrl_.rd_req_ack;
+  io.ctrl.wr_req.ack = ctrl_.wr_req_ack;
+  io.ctrl.pbuf(ctrl_.pbuf.io.deq);
+  ctrl_.pbuf.io.enq.data = io.qpi.rd_rsp.data;
+  io.ctrl.outstanding_writes = outstanding_writes_;
+
+  for (int i =0; i < PE_COUNT; ++i) {
+    io.pe[i].rd_req.ack  = pe_[i].rd_req_ack;
+    io.pe[i].rd_rsp.syn  = pe_[i].rd_rsp_valid;
+    io.pe[i].rd_rsp.type = pe_[i].rd_rsp_type;
+
+    io.pe[i].wr_req.ack  = pe_[i].wr_req_ack;
+
+    io.pe[i].axbuf(pe_[i].axbuf.io.deq);    
+    io.pe[i].asbuf(pe_[i].asbuf.io.deq);
+    io.pe[i].aybuf(pe_[i].aybuf.io.deq);
+    io.pe[i].avbuf(pe_[i].avbuf.io.deq);    
+    io.pe[i].xvbuf(pe_[i].xvbuf.io.deq);
+    io.pe[i].xmbuf(pe_[i].xmbuf.io.deq);
+
+    io.pe[i].avbuf_size(pe_[i].avbuf.io.size);
+
+    pe_[i].axbuf.io.enq.data = io.qpi.rd_rsp.data;
[…]

### 2017-11-02 lsu.h

@@ -36,6 +36,16 @@ __enum (ch_rq_owner, 2, (
   ctrl
 ));
 
+__struct(ch_rd_mdata_t, (
+  (ch_rd_request) type,
+  (ch_rq_owner) owner
+));
+
+__struct(ch_wr_mdata_t, (
+  (ch_wr_request) type,
+  (ch_rq_owner) owner
+));
+
 enum {
   LOG2_PBUF_SIZE = 2,
   PBUF_SIZE = 1 << LOG2_PBUF_SIZE,
@@ -60,44 +70,6 @@ enum {
 
   LOG2_ABUF_SIZE = 3,
   ABUF_SIZE = 1 << LOG2_ABUF_SIZE,
-};
-
-struct ch_rd_mdata_t {
-  using value_type = ch_bit<aal::aal_qpi0::mdata_width>;
-  ch_rd_request type;
-  ch_rq_owner owner;
-
-  ch_rd_mdata_t(const ch_rd_request& _type, const ch_rq_owner& _owner)
-    : type(_type)
-    , owner(_owner)
-  {}
-
-  ch_rd_mdata_t(const value_type& rhs) {
-    ch_tie(type, owner) = rhs.slice<ch_bitwidth_v<ch_rd_request> + ch_bitwidth_v<ch_rq_owner>>();
-  }
-
-  operator value_type() const {
-    return ch_zext<aal::aal_qpi0::mdata_width>(ch_cat(type, owner));
-  }
-};
-
-struct ch_wr_mdata_t {
-  using value_type = ch_bit<aal::aal_qpi0::mdata_width>;
-  ch_wr_request type;
-  ch_rq_owner owner;
-
-  ch_wr_mdata_t(const ch_wr_request& _type, const ch_rq_owner& _owner)
-    : type(_type)
-    , owner(_owner)
-  {}
-
-  ch_wr_mdata_t(const value_type& rhs) {
-    ch_tie(type, owner) = rhs.slice<ch_bitwidth_v<ch_wr_request> + ch_bitwidth_v<ch_rq_owner>>();
-  }           
-
-  operator value_type() const {
-    return ch_zext<aal::aal_qpi0::mdata_width>(ch_cat(type, owner));
-  }
 };
 
 using ch_blk_addr = ch_bit<aal::aal_qpi0::addr_wdith>;
@@ -203,14 +175,8 @@ private:
     done
   ));
 
[…]

### 2017-11-02 pe.cpp

@@ -27,7 +27,10 @@ spmv_pe::spmv_pe(uint32_t id)
   : id_(id)
   , xmblock_addr_(PTR_MAX_VALUE) // default initialize to max
   , xvblock_addr_(PTR_MAX_VALUE) // default initialize to max
-{}
+{
+  // io binding
+  io.ctrl.hwcntrs = hwcntrs_;
+}
 
 spmv_pe::~spmv_pe() {
   //--
@@ -52,22 +55,40 @@ void spmv_pe::describe() {
   for (int i = 0; i < 16; ++i) {
     y_value1.slice<32>(i*32) = y_values[i+16];
   }
+
+  auto ctrl_start_req_ack = io.ctrl.start.ack.asSeq();
   
+  auto lsu_rd_req_syn  = io.lsu.rd_req.syn.asSeq();
+  auto lsu_rd_req_type = io.lsu.rd_req.type.asSeq();
+  auto lsu_rd_req_addr = io.lsu.rd_req.addr.asSeq();
+
+  auto lsu_wr_req_syn  = io.lsu.wr_req.syn.asSeq();
+  auto lsu_wr_req_type = io.lsu.wr_req.type.asSeq();
+  auto lsu_wr_req_addr = io.lsu.wr_req.addr.asSeq();
+  auto lsu_wr_req_data = io.lsu.wr_req.data.asSeq();
+
+  auto& lsu_axbuf_dequeue = io.lsu.axbuf.ready;
+  auto& lsu_asbuf_dequeue = io.lsu.asbuf.ready;
+  auto& lsu_aybuf_dequeue = io.lsu.aybuf.ready;
+  auto& lsu_avbuf_dequeue = io.lsu.avbuf.ready;
+  auto& lsu_xvbuf_dequeue = io.lsu.xvbuf.ready;
+  auto& lsu_xmbuf_dequeue = io.lsu.xmbuf.ready;
+
   //--
   /*if (id_ == 0) {
     ch_print(fstring("{0}: PE[%d]: state={1}, rq_ack={2}, rq_typ={3}, rq_adr={4}, col={5}, coln={6}, ax={7}, xv={8}, row={9}, rown={10}, ya={11}, yv={12}, ym={13}, wq_ack={14}, wq_typ={15}, wq_adr={16}", id_).c_str(),
-           ch_getTick(), state, lsu_.rd_req_syn, lsu_.rd_req_type, lsu_.rd_req_addr,
[…]

### 2017-11-02 pe.h

@@ -13,7 +13,7 @@ __inout (ch_pe_start_req_io_t, (
 ));
 
 __inout (ch_ctrl_pe_io_t, (
-  (ch_pe_start_req_io_t) start_req,
+  (ch_pe_start_req_io_t) start,
   __in(ch_bit64) timer,
   __out(ch_pe_hwcntrs_t) hwcntrs
 ));
@@ -34,37 +34,8 @@ public:
 private:
 
   using ch_bit480 = ch_bit<480>;
-
-  struct ctrl_t {
-    ch_seq<ch_bit1> start_req_ack;
-
-    ctrl_t() : start_req_ack(1) // default initialize to '1'
-    {}
-  };
-
-  struct lsu_t {
-    ch_seq<ch_bit1>       rd_req_syn;
-    ch_seq<ch_rd_request> rd_req_type;
-    ch_seq<ch_blk_addr>   rd_req_addr;
-
-    ch_seq<ch_bit1>       wr_req_syn;
-    ch_seq<ch_wr_request> wr_req_type;
-    ch_seq<ch_blk_addr>   wr_req_addr;
-    ch_seq<ch_block_t>    wr_req_data;
-
-    ch_bit1 axbuf_dequeue;
-    ch_bit1 asbuf_dequeue;
-    ch_bit1 aybuf_dequeue;
-    ch_bit1 avbuf_dequeue;
-    ch_bit1 xvbuf_dequeue;
-    ch_bit1 xmbuf_dequeue;
-  };
   
   uint32_t  id_;
-
-  ctrl_t ctrl_;
-
-  lsu_t lsu_;
   
   ch_seq<ch_ptr>      col_curr_;
   ch_seq<ch_ptr>      col_end_;
@@ -89,7 +60,7 @@ private:
   ch_seq<ch_ptr>      row_blk_end_;
   ch_seq<ch_bit<6>>   row_blk_cnt_;
 
-  ch_seq<ch_pe_hwcntrs_t> hwcntr_;
+  ch_seq<ch_pe_hwcntrs_t> hwcntrs_;
   ch_seq<ch_bit64>    prof_start_;
   
   ch_seq<ch_bit<5>>   y_clr_cntr_;

### 2017-11-02 spmv.cpp

@@ -18,23 +18,20 @@ __enum (ch_dispatch_state, 2, (
 ));
 
 spmv_device::spmv_device() {
+  // create LSU
+  lsu_ = std::make_unique<ch_module<spmv_lsu>>();
+
   // create the PEs
   for (int i = 0; i < PE_COUNT; ++i) {
-    pe_[i].module = std::make_unique<ch_module<spmv_pe>>(i);
+    pe_[i] = std::make_unique<ch_module<spmv_pe>>(i);
   }
-  // create LSU
-  lsu_.module = std::make_unique<ch_module<spmv_lsu>>();
 
-  // bind PEs
+  // bind modules
   for (int i = 0; i < PE_COUNT; ++i) {
-    pe_[i].module->io.ctrl(pe_[i].io);
-    lsu_.module->io.pe[i](pe_[i].module->io.lsu);
+    lsu_->io.pe[i](pe_[i]->io.lsu);
   }
-
-  // bind LSU
-  lsu_.module->io.ctx(io.ctx);
-  lsu_.module->io.qpi(io.qpi);
-  lsu_.module->io.ctrl(lsu_.io);
+  lsu_->io.ctx(io.ctx);
+  lsu_->io.qpi(io.qpi);
 }
 
 spmv_device::~spmv_device() {
@@ -46,37 +43,42 @@ void spmv_device::describe() {
   ch_print("{0}: ctx.a.numparts={1}, ctx.a.partitions={2}",
               ch_getTick(), io.ctx.a.numparts, io.ctx.a.partitions);
 
-  // emit timer
-  timer_.next = timer_ + 1;
+  // timer
+  ch_seq<ch_bit64> timer;
+  timer.next = timer + 1;
+  for (auto& pe : pe_) {
+    pe->io.ctrl.timer = timer;
+  }
   
-  // emit main thread
+  // main thread
   this->main_thread();
     
-  // emit dispatcher
+  // dispatcher
   this->dispatch_thread();
-
-  //
-  // return outputs
-  //
-
-  for (int i = 0; i < PE_COUNT; ++i) {
-    pe_[i].io.start_req.syn = pe_[i].start_req_syn;
-  }
-
-  lsu_.io.rd_req.syn  = lsu_.rd_req_syn;
-  lsu_.io.rd_req.type = lsu_.rd_req_type;
-  lsu_.io.rd_req.addr = lsu_.rd_req_addr;
[…]

### 2017-11-02 spmv.h

@@ -21,32 +21,8 @@ protected:
   void main_thread();
   void dispatch_thread();
 
-  struct pe_t {
-    std::unique_ptr<ch_module<spmv_pe>> module;
-    ch_flip_t<ch_ctrl_pe_io_t> io;
-    ch_seq<ch_dcsc_part_t> parts;
-    ch_seq<ch_bit1> start_req_syn;
-  };
-
-  struct lsu_t {
-    std::unique_ptr<ch_module<spmv_lsu>> module;
-    ch_flip_t<ch_ctrl_lsu_io_t> io;
-
-    ch_seq<ch_bit1> rd_req_syn;
-    ch_seq<ch_rd_request> rd_req_type;
-    ch_blk_addr rd_req_addr;
-
-    ch_seq<ch_bit1> wr_req_syn;
-    ch_seq<ch_wr_request> wr_req_type;
-    ch_blk_addr wr_req_addr;
-    ch_block_t wr_req_data;
-
-    ch_bit1 pbuf_dequeue;
-  };
-
-  pe_t pe_[PE_COUNT];
-
-  lsu_t lsu_;
+  std::unique_ptr<ch_module<spmv_pe>> pe_[PE_COUNT];
+  std::unique_ptr<ch_module<spmv_lsu>> lsu_;
 
   ch_seq<ch_part_buf_t> part_buf_;
   ch_seq<ch_bit<LOG2_PARTITIONS_PER_BLOCK>> part_buf_size_;
@@ -56,11 +32,7 @@ protected:
   ch_seq<ch_ptr> part_curr_;
   ch_ptr numparts_;
 
-  ch_seq<ch_bit64> timer_;
-
   ch_seq<ch_hwcntrs_t> hwcntrs_;
-
-  ch_seq<ch_bit1> done_;
 };
 
 }

### 2017-11-03 cpu.cpp
LKG build - baseline implementation

@@ -360,8 +360,5 @@ void cpu_device::dump_stats(ch_tick t) {
   DbgPrint(0,"PE1 matrix values stalls = %d\n", (uint32_t)hwcntrs.PE1.a_values_stalls);
   DbgPrint(0,"PE1 vertex masks stalls = %d\n", (uint32_t)hwcntrs.PE1.x_masks_stalls);
 
-  CHECK(712 == t);
-  CHECK(580 == hwcntrs.PE0.min_latency);
-  CHECK(580 == hwcntrs.PE0.max_latency);
-  CHECK(36  == hwcntrs.PE0.execute_latency);
+  CHECK(19658 == t);
 }

### 2017-11-03 pe.cpp

@@ -56,7 +56,7 @@ void spmv_pe::describe() {
     y_value1.slice<32>(i*32) = y_values[i+16];
   }
 
-  auto ctrl_start_req_ack = io.ctrl.start.ack.asSeq();
+  auto ctrl_start_ack = io.ctrl.start.ack.asSeq();
   
   auto lsu_rd_req_syn  = io.lsu.rd_req.syn.asSeq();
   auto lsu_rd_req_type = io.lsu.rd_req.type.asSeq();
@@ -74,9 +74,9 @@ void spmv_pe::describe() {
   auto& lsu_xvbuf_dequeue = io.lsu.xvbuf.ready;
   auto& lsu_xmbuf_dequeue = io.lsu.xmbuf.ready;
 
-  //--
-  /*if (id_ == 0) {
-    ch_print(fstring("{0}: PE[%d]: state={1}, rq_ack={2}, rq_typ={3}, rq_adr={4}, col={5}, coln={6}, ax={7}, xv={8}, row={9}, rown={10}, ya={11}, yv={12}, ym={13}, wq_ack={14}, wq_typ={15}, wq_adr={16}", id_).c_str(),
+  /*//--
+  if (id_ == 0) {
+    ch_print(fstring("{0}: PE%d: state={1}, rq_ack={2}, rq_typ={3}, rq_adr={4}, col={5}, coln={6}, ax={7}, xv={8}, row={9}, rown={10}, ya={11}, yv={12}, ym={13}, wq_ack={14}, wq_typ={15}, wq_adr={16}", id_).c_str(),
            ch_getTick(), state, lsu_rd_req_syn, lsu_rd_req_type, lsu_rd_req_addr,
            col_curr_, col_end_, a_xindex_, x_value_, row_curr_, row_end_, y_addr_, y_value_, y_mask_,
            lsu_wr_req_syn, lsu_wr_req_type, lsu_wr_req_addr);
@@ -86,7 +86,7 @@ void spmv_pe::describe() {
   __switch (state) (
   __case (ch_exec_state::get_partition) (
     // wait for partition data
-    __if (io.ctrl.start.syn == ctrl_start_req_ack) (
+    __if (io.ctrl.start.syn != ctrl_start_ack) (
       // get partition columns data
       auto& part = io.ctrl.start.part;
       col_curr_.next = part.start.slice<ch_bitwidth_v<ch_ptr>>();
@@ -110,7 +110,6 @@ void spmv_pe::describe() {
     );
   )
   __case (ch_exec_state::clear_output) (
[…]

### 2017-11-09 lsu.cpp
backup

@@ -3,9 +3,15 @@
 using namespace spmv;
 using namespace spmv::accelerator;
 
-spmv_lsu::spmv_lsu()
-  : m_writemask0_addr(PTR_MAX_VALUE)
-  , m_writemask1_addr(PTR_MAX_VALUE) {
+__enum (ch_wr_req_state, 3, (
+  get_request,
+  submit,
+  writemask,
+  flush,
+  wait_for_flush
+));
+
+spmv_lsu::spmv_lsu() {
   //
   // bind modules
   //
@@ -26,7 +32,6 @@ void spmv_lsu::describe() {
   this->read_rsp_thread();
   this->write_rsp0_thread();
   this->write_rsp1_thread();
-  this->writemask_thread();
 }
 
 void spmv_lsu::read_req_thread() {
@@ -60,7 +65,11 @@ void spmv_lsu::write_req_thread() {
   ch_seq<ch_wr_req_state> state;
 
   //--
-  wr_req_arb_.io.out.ready = (state == ch_wr_req_state::get_request);
+  wr_cache_.io.evict.ready = (state == ch_wr_req_state::get_request);
+
+  //--
+  wr_req_arb_.io.out.ready = (state == ch_wr_req_state::get_request
+                           && !wr_cache_.io.evict.valid);
 
   //--
   auto qpi_wr_req_valid = io.qpi.wr_req.valid.asSeq();
@@ -77,17 +86,34 @@ void spmv_lsu::write_req_thread() {
 
   //---
   auto outstanding_writes = io.ctrl.outstanding_writes.asSeq();
+
+  // QPI write thread
+  {
+
+    __switch (qstate) (
+    __case (
+
+    )
+
+    );
+  }
   
-  //--
+  // control thread
   __switch (state) (
   __case (ch_wr_req_state::get_request) (
-    __if (wr_req_arb_.io.out.valid) (
+    __if (wr_cache_.io.evict.valid) (
+      qpi_wr_req_data.next  = wr_cache_.io.evict.data.data;
+      qpi_wr_req_mdata.next = ch_wr_mdata_t(ch_zext<PE_COUNT+1>(wr_cache_.io.evict.data.owner), ch_wr_request::y_masks);
+      qpi_wr_req_addr.next  = get_baseaddr(ch_wr_request::y_masks) + wr_cache_.io.evict.data.tag;
+      state.next = ch_wr_req_state::submit;
+    )
[…]

### 2017-11-09 lsu.h

@@ -1,6 +1,7 @@
 #pragma once
 
 #include "common.h"
+#include "wrcache.h"
 
 namespace spmv {
 namespace accelerator {
@@ -102,21 +103,6 @@ public:
 
 private:
 
-  __enum (ch_wr_req_state, 3, (
-    get_request,
-    submit,
-    writemask,
-    flush_writemask1,
-    flush_writemask2
-  ));
-
-  __enum (ch_writemask_state, 2, (
-    ready,
-    check_mask0,
-    check_mask1,
-    done
-  ));
-
   void read_req_thread();
   void write_req_thread();
   
@@ -124,27 +110,12 @@ private:
   void write_rsp0_thread();
   void write_rsp1_thread();
   
-  void writemask_thread();
-  
   ch_blk_addr get_baseaddr(const ch_rd_request& rq_type);
   ch_blk_addr get_baseaddr(const ch_wr_request& rq_type);
 
   ch_module<ch_rd_req_arb_t> rd_req_arb_;
   ch_module<ch_wr_req_arb_t> wr_req_arb_;
-  
-  ch_seq<ch_block>    m_writemask0;
-  ch_seq<ch_block>    m_writemask1;
-  ch_seq<ch_block>    m_writemask_flush;
-  ch_seq<ch_blk_addr> m_writemask_flush_addr;
-  ch_seq<ch_bit<2>>   m_writemask0_owners;
-  ch_seq<ch_bit<2>>   m_writemask1_owners;
-  ch_seq<ch_blk_addr> m_writemask0_addr;
-  ch_seq<ch_blk_addr> m_writemask1_addr;
-  ch_seq<ch_bit1>     m_writemask_flush_enable;
-  ch_seq<ch_bit1>     m_pe_writemask_out;
-  
-  ch_wr_req_state     m_wr_req_state;
-  ch_writemask_state  m_writemask_state;
+  ch_module<spmv_write_cache<ch_block, PE_COUNT, ch_bitwidth_v<ch_blk_addr>>> wr_cache_;
 };
 
 }

### 2017-11-09 wrcache.h

@@ -1,0 +1,205 @@
+#pragma once
+
+#include "common.h"
+
+namespace spmv {
+namespace accelerator {
+
+template <typename T, unsigned N, unsigned TagBits, unsigned LookupCycles = 1 + N/8>
+class spmv_write_cache {
+private:
+
+  using ch_idx = ch_bit<log2ceil(N)>;
+  using ch_tag = ch_bit<TagBits>;
+
+  enum {
+    MAX_TAG_VALUE = (1 << TagBits) - 1,
+  };
+
+  __enum (ch_state, 2, (
+    ready,
+    lookup,
+    evict,
+    flush
+  ));
+
+  __struct (tag_t, (
+    (ch_bit<N>) owners,
+    (ch_tag)    tag,
+    (ch_bool)   valid
+  ));
+
+  ch_ram<tag_t, N>    tags_;
+  ch_ram<ch_block, N> data_;
+  ch_seq<ch_idx>      last_used_idx_;
+  ch_seq<ch_tag>      last_used_tag_;
+
+public:
+
+  __struct (entry_t, (
+    (ch_bit<N>) owner,  // one-hot representation
+    (ch_tag)    tag,
+    (T)         data
+  ));
+
+  __io (
+    (ch_deq_io<entry_t>) enq,
+    (ch_enq_io<entry_t>) evict,
+    __in(ch_bool)        flush,
+  );
+
+  spmv_write_cache() : last_used_tag_(MAX_TAG_VALUE)
+  {}
+
+  void describe() {
+    //--
+    ch_idx match_block_idx(0);
+    ch_idx owned_block_idx(0);
+    ch_idx free_block_idx(0);
+    ch_bool lookup_valid = io.enq.valid;
+
+    // lookup thread
+    {
+      int step = CH_CEILDIV(N, LookupCycles);
+      for (int i = 1; i < N; ++i) {
+        match_block_idx = ch_select((tags_[i].tag == io.enq.data.tag), i, ch_clone(match_block_idx));
+        owned_block_idx = ch_select((tags_[i].owners & io.enq.data.owner) != 0, i, ch_clone(owned_block_idx));
+        free_block_idx  = ch_select(tags_[i].owners == 0, i, ch_clone(free_block_idx));
+        if (0 == (i % step)) {
+          match_block_idx = ch_reg(match_block_idx);
+          owned_block_idx = ch_reg(owned_block_idx);
+          free_block_idx  = ch_reg(free_block_idx);
[…]

### 2017-11-10 common.h
write cache refactoring

@@ -50,7 +50,11 @@ enum {
   PARTITION_VALUE_BITS = 32,
   PARTITIONS_PER_BLOCK = ch_bitwidth_v<ch_block> / PARTITION_VALUE_BITS,
   LOG2_PARTITIONS_PER_BLOCK = log2ceil(PARTITIONS_PER_BLOCK+1),
+
+  CTRL_ID = (1 << PE_COUNT),
 };
+
+#define PE_ID(x) (1 << x)
 
 extern bool g_singlecore; 
 

### 2017-11-10 lsu.cpp

@@ -11,6 +11,11 @@ __enum (ch_wr_req_state, 3, (
   wait_for_flush
 ));
 
+__enum (ch_qpi_write_state, 1, (
+  ready,
+  write
+));
+
 spmv_lsu::spmv_lsu() {
   //
   // bind modules
@@ -63,112 +68,132 @@ void spmv_lsu::read_req_thread() {
 void spmv_lsu::write_req_thread() {
   //--
   ch_seq<ch_wr_req_state> state;
+  ch_seq<ch_qpi_write_state> qw_state;
 
   //--
-  wr_cache_.io.evict.ready = (state == ch_wr_req_state::get_request);
+  ch_bool lsu_write_valid;
+  auto qpi_write_ready = (qw_state == ch_qpi_write_state::ready);
 
   //--
-  wr_req_arb_.io.out.ready = (state == ch_wr_req_state::get_request
-                           && !wr_cache_.io.evict.valid);
+  wr_req_arb_.io.out.ready = (state == ch_wr_req_state::get_request);
+  wr_cache_.io.evict.ready = (qpi_write_ready & !lsu_write_valid);
 
-  //--
-  auto qpi_wr_req_valid = io.qpi.wr_req.valid.asSeq();
-  auto qpi_wr_req_addr  = io.qpi.wr_req.addr.asSeq();
-  auto qpi_wr_req_data  = io.qpi.wr_req.data.asSeq();
-
-  //--
-  ch_seq<ch_wr_mdata_t> qpi_wr_req_mdata;
-  io.qpi.wr_req.mdata = ch_zext<aal::aal_qpi0::mdata_width>(qpi_wr_req_mdata.asBits());
-
-  //--
-  auto wr_rsp_cnt = ch_select(io.qpi.wr_rsp0.valid & io.qpi.wr_rsp1.valid, 0x2_h32)
-                              (io.qpi.wr_rsp0.valid | io.qpi.wr_rsp1.valid, 0x1_h32)(0);
-
-  //---
-  auto outstanding_writes = io.ctrl.outstanding_writes.asSeq();
+  ch_seq<ch_lsu_wr_req_t> req_data;
+  ch_seq<ch_bit<PE_COUNT+1>> req_owner;
 
   // QPI write thread
   {
+    auto wr_rsp_cnt = ch_select(io.qpi.wr_rsp0.valid & io.qpi.wr_rsp1.valid, 0x2_h32)
[…]

### 2017-11-10 lsu.h

@@ -18,11 +18,7 @@ namespace accelerator {
 enum {
   RD_REQ_INPUTS = PE_COUNT + 1,
   WR_REQ_INPUTS = PE_COUNT + 1,
-
-  CTRL_ID = (1 << PE_COUNT),
 };
-
-#define PE_ID(x) (1 << x)
 
 __enum (ch_rd_request, 3, (
   a_partition,

### 2017-11-10 test_wmcache.cpp

@@ -1,0 +1,74 @@
+#include <iostream>
+#include <fstream>
+#include <unistd.h>
+#include "test.h"
+#include "wrcache.h"
+
+using namespace spmv;
+using namespace spmv::accelerator;
+
+int main(int argc, char **argv) {
+
+  ch_device<spmv_write_cache<ch_bit4, 8, 8, 2>> device;
+  ch_simulator sim(device);
+  ch_tick t = sim.reset(0);
+  t = sim.step(t);
+  int ret = 1;
+  assert(!!device.io.enq.ready);
+
+  device.io.enq.data.tag   = 0;
+  device.io.enq.data.owner = PE_ID(0);
+  device.io.enq.data.data  = 0001_b;
+  device.io.enq.valid = true;
+  t = sim.step(t);
+  device.io.enq.valid = false;
+  t = sim.step(t,3);
+
+  device.io.enq.data.tag   = 0;
+  device.io.enq.data.owner = PE_ID(0);
+  device.io.enq.data.data  = 0010_b;
+  device.io.enq.valid = true;
+  t = sim.step(t);
+  device.io.enq.valid = false;
+  t = sim.step(t,3);
+
+  device.io.enq.data.tag   = 0;
+  device.io.enq.data.owner = PE_ID(1);
+  device.io.enq.data.data  = 0100_b;
+  device.io.enq.valid = true;
+  t = sim.step(t);
+  device.io.enq.valid = false;
+  t = sim.step(t,3);
+
+  device.io.enq.data.tag   = 1;
+  device.io.enq.data.owner = PE_ID(0);
+  device.io.enq.data.data  = 0001_b;
+  device.io.enq.valid = true;
+  t = sim.step(t);
+  device.io.enq.valid = false;
+  t = sim.step(t,3);
+
+  device.io.enq.data.tag   = 1;
+  device.io.enq.data.owner = PE_ID(1);
+  device.io.enq.data.data  = 0010_b;
+  device.io.enq.valid = true;
+  t = sim.step(t);
+  device.io.enq.valid = false;
+  t = sim.step(t,3);
+
+  assert(0x7 == device.io.evict.data.data);
+  device.io.evict.ready = true;
+  t = sim.step(t);
+  device.io.evict.ready = false;
+  t = sim.step(t);
+  device.io.flush = true;
+  t = sim.step(t);
+  device.io.flush = false;
+  t = sim.step(t, 4);
[…]

### 2017-11-10 wrcache.h

@@ -5,7 +5,7 @@
 namespace spmv {
 namespace accelerator {
 
-template <typename T, unsigned N, unsigned TagBits, unsigned LookupCycles = 1 + N/8>
+template <typename T, unsigned N, unsigned TagBits, unsigned LookupCycles = CH_CEILDIV(N, 4)>
 class spmv_write_cache {
 private:
 
@@ -16,23 +16,34 @@ private:
     MAX_TAG_VALUE = (1 << TagBits) - 1,
   };
 
-  __enum (ch_state, 2, (
+  __enum (ch_state, 3, (
     ready,
     lookup,
+    check_evict,
     evict,
     flush
   ));
 
   __struct (tag_t, (
     (ch_bit<N>) owners,
-    (ch_tag)    tag,
-    (ch_bool)   valid
+    (ch_tag)    tag
   ));
 
-  ch_ram<tag_t, N>    tags_;
-  ch_ram<ch_block, N> data_;
-  ch_seq<ch_idx>      last_used_idx_;
-  ch_seq<ch_tag>      last_used_tag_;
+  ch_ram<tag_t, N> tags_;
+  ch_ram<T, N>     data_;
+  ch_seq<ch_idx>   last_used_idx_;
+  ch_seq<ch_tag>   last_used_tag_;
+
+  ch_idx port0_;
+  ch_idx port1_;
+
+  T data_value_;
+  ch_bool data_wenable_;
+
+  tag_t tag_value_;
+  ch_bool tag_wenable_;
+
+  ch_seq<ch_idx> counter_;
 
 public:
 
@@ -45,45 +56,58 @@ public:
   __io (
     (ch_deq_io<entry_t>) enq,
     (ch_enq_io<entry_t>) evict,
-    __in(ch_bool)        flush,
+    __in(ch_bool)        flush
   );
 
-  spmv_write_cache() : last_used_tag_(MAX_TAG_VALUE)
-  {}
+  spmv_write_cache() {
+    //--
+    __if (data_wenable_) (
+      data_[port0_] = data_value_;
+    );
+
+    //--
+    __if (tag_wenable_) (
+      tags_[port0_] = tag_value_;
+    );
+  }
 
   void describe() {
-    //--
-    ch_idx match_block_idx(0);
-    ch_idx owned_block_idx(0);
-    ch_idx free_block_idx(0);
-    ch_bool lookup_valid = io.enq.valid;
-
-    // lookup thread
-    {
-      int step = CH_CEILDIV(N, LookupCycles);
[…]

### 2017-11-10 common.h
backup

@@ -88,21 +88,23 @@ inline bool test_bitvector(const uint32_t* bitvec, uint32_t idx) {
 }
 
 __struct (ch_pe_hwcntrs_t, (
-  (ch_bit<32>) num_partitions,
-  (ch_bit<32>) min_latency,
-  (ch_bit<32>) max_latency,
-  (ch_bit<32>) total_latency,
-  (ch_bit<32>) execute_latency,
-  (ch_bit<32>) a_xindices_stalls,
-  (ch_bit<32>) x_masks_stalls,
-  (ch_bit<32>) a_startys_stalls,
-  (ch_bit<32>) a_values_stalls
+  (ch_bit32) num_partitions,
+  (ch_bit32) min_latency,
+  (ch_bit32) max_latency,
+  (ch_bit32) total_latency,
+  (ch_bit32) execute_latency,
+  (ch_bit32) a_xindices_stalls,
+  (ch_bit32) x_masks_stalls,
+  (ch_bit32) a_startys_stalls,
+  (ch_bit32) a_values_stalls,
+  (ch_bit32) write_value_stalls,
+  (ch_bit32) write_mask_stalls
 ));
 
 __struct (ch_ctrl_hwcntrs_t, (
-  (ch_bit<32>) a_partitions_stalls,
-  (ch_bit<32>) vcache_hits,
-  (ch_bit<32>) vcache_stalls
+  (ch_bit32) a_partitions_stalls,
+  (ch_bit32) vcache_hits,
+  (ch_bit32) vcache_stalls
 ));
 
 __inout (ch_matrix_dcsc_t, (

### 2017-11-10 cpu.cpp

@@ -362,5 +362,7 @@ void cpu_device::dump_stats(ch_tick t) {
     DbgPrint(0, "PE%d matrix ystarts stalls = %d\n", i, (uint32_t)hwcntrs.a_startys_stalls);
     DbgPrint(0, "PE%d matrix values stalls = %d\n", i, (uint32_t)hwcntrs.a_values_stalls);
     DbgPrint(0, "PE%d vertex masks stalls = %d\n", i, (uint32_t)hwcntrs.x_masks_stalls);
+    DbgPrint(0, "PE%d matrix write values stalls = %d\n", i, (uint32_t)hwcntrs.write_value_stalls);
+    DbgPrint(0, "PE%d matrix write masks stalls = %d\n", i, (uint32_t)hwcntrs.write_mask_stalls);
   }
 }

### 2017-11-10 pe.cpp

@@ -467,6 +467,10 @@ void spmv_pe::describe() {
     __if (io.lsu.wr_req.ready) (
       // write second y_value block
       state.next = ch_exec_state::write_y_value1;
+    )
+    __else (
+      // profiling
+      hwcntrs_.next.write_value_stalls = hwcntrs_.write_value_stalls + 1;
     );
   )
   __case (ch_exec_state::write_y_value1) (        
@@ -478,6 +482,10 @@ void spmv_pe::describe() {
     // wait for LSU ack
     __if (io.lsu.wr_req.ready) (
       state.next = ch_exec_state::write_y_mask;
+    )
+    __else (
+      // profiling
+      hwcntrs_.next.write_value_stalls = hwcntrs_.write_value_stalls + 1;
     );
   )
   __case (ch_exec_state::write_y_mask) (
@@ -496,7 +504,11 @@ void spmv_pe::describe() {
       hwcntrs_.next.num_partitions = hwcntrs_.num_partitions + 1;
       // advance to the next partition
       state.next = ch_exec_state::get_partition;      
-    );
+    )
+      __else (
+        // profiling
+        hwcntrs_.next.write_mask_stalls = hwcntrs_.write_mask_stalls + 1;
+      );
   )
   __default (
     //--

### 2017-11-10 spmv.h

@@ -43,7 +43,7 @@ protected:
 
   ch_seq<ch_bit32> pbuf_pending_size_;
 
-  ch_seq<ch_bit<32>> a_partitions_stalls_;
+  ch_seq<ch_bit32> a_partitions_stalls_;
 
   ch_seq<ch_ptr> part_blk_curr_;
   ch_seq<ch_ptr> part_blk_end_;

### 2017-11-15 paper/abstract.tex
backup

@@
 
-In recent years, The end of Dennard Scaling \cite{Dark-Silicon} and the need for agile development methodology in data centers is pushing the computer architecture community towards designing more specialized and energy efficient systems \cite{Catapult}.
-This has lead to a new application domain for FPGAs and the emergence of heterogeneous CPU-FPGA computing platforms \cite{Intel-FPGA} \cite{IBM-FPGA}, enabling the design of new energy efficient FPGA accelerators for domain specific applications.
-Graph Analytics, one the largest applications in production data centers today, faces scaling challenges with the ever increasing size of its workloads and its inherent low locality and memory-bound characteristic.
-The GraphMat \cite{GraphMat} programming framework was proposed that efficiently translates graph algorithms into high performance computation in hardware.
-We present FlexGraph, a reconfigurable domain-specific accelerator for Graph Analytics targeting commodity heterogeneous CPU-FPGA platforms.
-FlexGraph is designed using Cocoh \cite{Cocoh}, a C++ domain specific library for hardware design and simulation, providing a high level programming abstraction with the efficiency of native RTL.
-FlexGraph's novel architecture allows an easy implementation of GraphMat kernels using standard C++.
-To the best of our knowledge it is a first implementation of a graph accelerator on a commodity heterogeneous CPU-FPGA platform.
-FlexGraphs's preliminary performance is comparable to HLS, we later discuss several optimization strategies that could further improve its speed.
+In recent years, The end of Dennard Scaling is pushing the computer architecture community towards designing more specialized and energy efficient systems.
+This move has led to a new application domain for FPGAs and the emergence of heterogeneous CPU-FPGA computing platforms, enabling the design of new energy efficient FPGA accelerators for domain specific applications.
+Graph Analytics, one the largest applications in production data centers today, faces scaling challenges with its ever increasing workload size, inherent sparsity and memory-bound characteristic.
+We present FlexGraph, a flexible and energy-efficient Graph Analytics Accelerator for heterogeneous CPU-FPGA architectures.
+FlexGraph uses a Doubly Compressed Sparse Matrix format to eliminate unnecessary data transfers and reduce storage requirements.
+Our architecture allows support for other compressed format by decoupling the matrix data structure traversal from the compute and memory units.
+We introduce memory access hardware primitives for unstructured fine-grained accesses on cache-coherent shared memory.
+Flexgraph uses a C++ Hardware Generation Language to extend its vertex programming model with domain centric reconfigurability in an integrated single source development environment, providing a design efficient alternative to High Level Synthesis.
+Flexgraph achieves X GFlop/s, performing Nx faster than HLS implementation.
 

### 2017-11-15 paper/evaluation.tex
backup

@@
 
-\subsection{Cocoh's Simulation}
+In this work, we simulated our evaluation environment entirely in software using the Cocoh \cite{Cocoh} C++ framework.
+Figure \ref{fig:simulation} shows an overview of what the simulation environment contains.
+At the top we have the accelerator implementation with its AAL interface, both modelled using Cocoh's API.
+At the bottom, we have the Cocoh's simulation runtime that bind the accelerator and the back-end host CPU simulator.
+Inside our host CPU simulator we modelled a QPI memory simulator which managed the shared memory space and served all the memory requests coming from the accelerator.
+In the current design we used a 1 GB shared memory space for holding the matrix and vertex data, including the output result from the accelerator.
+On the actual Xeon platform \cite{Intel-FPGA}, the shared memory space limit is 2 GB.
 
-\subsection{Intel AAL Interface}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.3\textwidth]{figures/simulation}
+\caption{FlexGraph Simulation Environment}
+\label{fig:simulation}
+\end{figure}
 
-\subsection{Intel QPI Emulation}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.4\textwidth]{figures/simulation_parameters}
+\caption{FlexGraph Simulation Parameters}
+\label{fig:simulation_parameters}
+\end{figure}
+
+\subsection{Intel QPI Simulation}
+We simulated the Intel QPI memory transaction using an analytic model capturing the latency of memory transfers when the request hits local CCI \cite{CCI} cache or the host processor LLC.
+All outgoing memory requests from the accelerator to CCI take two cycles and the responses from CCI hits take another two cyles, giving a minimum round trip latency of four cycles for read requests.
+Figure \ref{fig:simulation_parameters} shows some configuration parameters we use for both the simulation and accelerator modelling.
 
@@
 
+We used the Graph500 \cite{Graph500} synthetic datasets for our evaluation.
+We configured graph generator similar to GraphMat \cite{GraphMat} setting the default RMAT paramters (A=0.57, B=C=0.15 and D=1-A-B-C).
+We used five scaling factors for our graphs (6, 8, 10, 12, 14) with corresponding number of non-zero ranging from 1024 to 262,144 accordingly.
+These datasets are relatively small mainly due to simulation time constraints, but the setup could have supported up to a scaling factor of 20 (about 16,777,216 non-zeros).
+Due to time constraints we did not test any real world dataset, that will be discussed in our future work section.
+Figure \ref{fig:graph500_dataset} lists down the number of vertices and non-zeros for our dataset.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.4\textwidth]{figures/graph500_dataset}
+\caption{Graph500 dataset}
+\label{fig:graph500_dataset}
+\end{figure}
+
 \section{Results Analysis}
 
-\subsection{Overall Performance}
+In this section, we discuss our evaluation results.
+
+\subsection{Throughput Performance}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/dual_core_perf}
+\caption{FlexGraph Throughput Performance}
+\label{fig:dual_core_perf}
+\end{figure}
+
+Figure \ref{fig:dual_core_perf} shows the throughput performance of the accelerator in GFlops. We estimated the performance using the following equation:
+\[Perf = \dfrac{2 * nonzeros}{latency}\]
+The performance varies from 0.67 Gflops to 0.1 Gflops as the size of the dataset increases.
+This result is comparable to our implementation of a similar kernel using the Altera OpenCL HLS, also showing similar performance degradation as the workload size increases.
+In prior work, GraphOps \cite{GraphOps} peak performance for their FPGA implementation was 0.18 GFlops with a graph size of 512K vertices, about x3 our throughput.
+The target platform has peak memory bandwidth is 6.0 GB/s, this shows that there is room for improvement.
+than We identified several optimization opportunities that we will discuss later in this section.
 
@@
 
+Figure \ref{fig:single_core_perf} shows the throughput performance of the accelerator in GFlops for a single-core system versus the dual-core implementation.
+We can see a considerable drop in performance (by almost half) across all datasets when a single processing element is executing.
+This is happening because there is much less memory level parallelism that can be exploited when a single core is running.
+In a dual-core system, the LSU receives a lot more memory requests from the processing element and has much less idle time when QPI is taking more cycles to return the blocks.
+When a processing element is stalled on a request, another may be able to make progress.
+We believe that increase the number of processing element beyond two will have a possible impact in performance for FlexGraph, possibly doubling the current throughput.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/single_core_perf}
+\caption{FlexGraph Single Core Performance}
+\label{fig:single_core_perf}
+\end{figure}
+
 \subsection{The Impact of Memory Optimizations}
 
-\subsection{FlexGraph Performance Hogs}
+Figure \ref{fig:no_opts_perf} shows the accelerator performance with all memory optimizations disabled in GFlops.
+Once again the graph shows a similar performance degradation as the dataset size decreases.
+The memory optimizations give the accelerator an average performance speed-up of 45\%, which is significant considering the fact the the system is memory bound.
 
-\subsection{Hardware Cost Evaluation}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/no_opts_perf}
+\caption{FlexGraph Performance with memory optimizations}
+\label{fig:no_opts_perf}
+\end{figure}
+
+\subsection{Summary Discussion}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/memory_stalls}
+\caption{FlexGraph Memory Stalls}
+\label{fig:memory_stalls}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/hardware_cost}
+\caption{Cocoh's hardware cost evaluation}
+\label{fig:hardware_cost}
+\end{figure}
+
+In summary, FlexGraph current average performance is 0.1 Gflops (using the lower bound of the largest dataset in our experiment).
+This is in part with our HLS implementation of the same kernel.
+We have identified several optimization opportunities to boost the accelerator performance.\\
+We added some performance counters into FlexGraph accelerator to monitor the memory stalls in each processing elements and identify bottlenecks.
+Figure \ref{fig:memory_stalls} shows the memory stalls distribution over three datasets \textit{m8}, \textit{m10} and \textit{m12}.
+The blue region represents the execution cycles, the orange region represents the stall cycles occurring when we access the matrix column indices, this is \textit{coldata} buffer in the SPMV speudo-code (line 4 in listing 1).
+The gray region represents the stall cycles when accessing the vertices active masks (line 6 in listing 1).
+We can significantly reduce the stall cycles by prefetching the blocks before they are needed to hide the memory latency.
+Using a simple next-line prefetcher will suffice since we know the access pattern on \textit{coldata}.\\
+Another performance optimization is in reducing the impact of the branch operation when checking if a vertex is active of not (line 7 in listing 1).
+We will investigate adding a preliminary test on the entire 32-bit mask to see if it is zero and skip the whole block altogether.\\
+Lastly, another performance opportunity is increasing the number of precessing elements.
+Figure \ref{fig:single_core_perf} showed the performance gain of using two cores versus a single one.
+Having multiple processing elements will improve the system memory level parallelism by overlapping the execution of some processing element while others are stalled on memory accesses.
+We ran some estimate of the current design hardware cost using Cocoh's framework (see Figure \ref{fig:hardware_cost}) and it showed that our memory usage is very small (about 1\% of the target Stratix V FPGA capacity).
+Also, FlexGraph accelerator only currently consumes 15\% of the registers capacity.
+With this reserve, we anticipate extending the number of processing elements to 16.
+A challenge that we will have to resolve when adding additional processing elements is the synchronization of write accesses for the active masks.
+Another challenge will be the arbitration latency in the LSU when communicating with all nodes,    since the single LSU will now become a bottleneck in the system.

### 2017-11-15 paper/introduction.tex
backup

@@
 These constraints have pushed production data centers towards using FPGA-based accelerators \cite{Catapult} for their reconfigurability and energy savings when compared to general-purpose graphics processing units (GPUs).
-This trend has led to emergence of heterogeneous CPU-FPGA computing platforms \cite{Intel-FPGA} \cite{IBM-FPGA}, enabling the design of new energy efficient FPGA accelerators for domain specific applications.
-Graph Analytics is one the largest applications in production data centers today \cite{BigData}, its performance surfers from workload imbalance, frequent updates, limited data locality and low compute communication ratio making it memory-bound and energy inefficient.
+This trend has led to emergence of heterogeneous CPU-FPGA computing platforms \cite{Intel-FPGA} \cite{IBM-FPGA}, enabling fast development of new energy efficient accelerators \cite{CPU-FPGA} for domain specific applications.
+Graph Analytics is one the largest applications in production data centers today \cite{BigData},
+spanning domains such as bio-informatics, social network, mining, cyber-security, etc.
+Graph Analytics performance suffers from workload imbalance, frequent updates, limited data locality and low compute communication ratio making it memory-bound and energy inefficient.
 This problem is further exacerbated with the ever increasing size of its dataset, presenting a scalability challenge for the industry.
-Several solution have been proposed to address this problem both at the software level with better algorithms and programming abstraction \cite{GraphX} \cite{Galois} \cite{GraphMat} \cite{Pregel} \cite{GraphLab}, and at the hardware level with custom accelerators \cite {Graphicionado} \cite{Tesseract} \cite{GraphOps}.
-GraphMat \cite{GraphMat} and many other graph frameworks \cite{GraphMat} \cite{Pregel} \cite{GraphLab} define a vertex programming model to represent graph computation, but GraphMat is unique in that only the edges transformation of the kernel is accelerated, allowing the programmer to implement its algorithm is C++, instead of a proprietary language, increasing both accessibility and productivity.
-Flexgraph's design goals are aiming into achieving the same balance between efficiency and productivity via hardware acceleration.
-We developed FlexGraph using the Cocoh's \cite{Cocoh} framework, a C++ domain specific library for hardware design and simulation, providing a high level programming abstraction with the efficiency of native RTL.
-Several hardware description languages and libraries \cite{CHDL} \cite{Chisel} \cite{SystemC} \cite{BlueSpec} \cite{MyHDL} \cite{JHDL} \cite{Esterel} have been proposed as alternative to native verilog or VHDL to improve developer productivity by providing a higher level programming abstraction \cite{Chisel} \cite{SystemC} and execution model \cite{BlueSpec} \cite{Esterel}.
-Cocoh's unique strength is that it provides a uniform development and simulation environment based on the same source.
-Cocoh's modules are written directly in standard C++, debugged, tested and simulated using the same source, and later exported to verilog for FPGA deployment.
-Another strength of Cocoh's API is its integration with CHDL \cite{CHDL}, allowing gate-level simulation and analysis of the design.
-Using the Cocoh's API, developpers can write their custom GraphMat kernel for the FlexGraph accelerator and test their design before production.
-FlexGraph architecture uniquely targets commodity heteregenous CPU-FPGA computing platforms \cite{Intel-FPGA} for data centers.
-To the best of our knowledge it is a first implementation of a graph accelerator targeting this ecosystem.
-FlexGraphs's current performance is comparable to HLS, we have identified several performance optimizations that could further improve its performance which we will discuss later in the paper.
+Several solutions have been proposed to address this problem both at the software level, with better algorithms and programming abstraction \cite{GraphX} \cite{Galois} \cite{GraphMat} \cite{Pregel} \cite{GraphLab}, and at the hardware level with custom accelerators \cite {Graphicionado} \cite{Tesseract} \cite{GraphOps}.
+Most Graph Analytics accelerators \cite {Graphicionado} \cite{Tesseract} \cite{GraphOps} define a proprietary graph data structure to exploit efficient computation on their hardware, limiting flexibility on the host processor for efficient software processing.
+FlexGraph uses sparse matrices as underlying data structure to enable efficient computation in a shared CPU-FPGA environment where both the host processor and the accelerator are modifying the same graph.
+However, accessing unstructured fine-grained data on a cache coherent FPGA fabric pauses challenges because of the restricted coarse-grained cache line access granularity and longer miss penalty.
+Additionally, generalized sparse matrix storage encoding COO, ELL, CSR, CSC, lack the same level of compaction as directly using non-zero edges list \cite {Graphicionado}, pausing a problem for energy efficiency and memory bandwidth for hyper-sparse Graph Analytics application.
+FlexGraph uses a Doubly Compressed Column Based Sparse matrix encoding similar to GraphMath \cite{GraphMat}, however introducing additional indirections to the matrix traversal path.
+We decoupled FlexGraph's matrix traversal unit from the rest of the compute and memory fabric to scaling by increasing memory bandwidth and provide support for other formats without altering the rest of the system.\\
+Our contributions are as follows:\\
+1.	A specialized Graph Analytics Accelerator for heterogeneous CPU-FPGA fabric that provide efficient collaborative computation while maximizing energy efficiency.\\
+2. 	A decoupled data structure traversal and compute architecture for sparse matrices enabling maximum bandwidth utilization and compute scaling.\\
+3.	We introduce hardware primitives for unstructured fine-grained accesses on coherent shared memory architectures.\\
+4.	A domain specific accelerator with single source programming and reconfiguration environment providing a design-efficient alternative to High-Level Synthesis.
+
 The remainder of this paper is organized as follows:
-Section 2 describes GraphMat computing model and its SPMV kernel, section 3 describes FlexGraph's architecture, section 4 describes the experimental setup, section 5 describes our results and analysis, section 6 describes the related work, section 7 discusses future work and finally Section 8 summarizes our main contribution and results.
+Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 describe FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 5 describes the experimental setup, section 6 describes our results analysis, section 7 describes the related work, section 8 summarizes our main contribution and results.

### 2017-11-15 paper/mechanism.tex
backup

@@
+\section{Background and Motivation}
+
+\subsection{Graph Vertex Processing}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.4\textwidth]{figures/sample_graph}
+\caption{Sample Graph Representation}
+\label{fig:sample_graph}
+\end{figure}
+
+\begin{figure}[ htbp ] 
+\begin{equation}
+\#Edges = A^{T} \times Identity
+\end{equation}
+\begin{equation}
+\#Edges = 
+\left(\begin{array}{ccc} 0 & 0 & 0 \\ 1 & 0 & 0 \\ 1 & 1 & 0 \end{array}\right) \times
+\left(\begin{array}{c} 1 \\ 1 \\ 1 \end{array}\right) 
+= \left(\begin{array}{c} 0 \\ 1 \\ 2 \end{array}\right) 
+\end{equation}
+\caption{Incoming Edges Algorithm}
+\label{fig:sample_algorithm}
+\end{figure}
+
+\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=Graph Processing Kernel]
+GraphKernel (A, P, iterations) :
+  for_each iterations
+    x := Assign(P)
+  	y := Process(A,x)
+  	P := Apply(y)	  
+  end_for
+
+Incoming Edges := GraphKernel(A, identity, 1):
+	with Assign(P)    := P
+	     Process(A,x) := A * x
+	     Apply(y)     := y
+\end{lstlisting}
+
+FlexGraph's architecture is based on GraphMat \cite{GraphMat}'s vertex computing paradigm where a graph algorithm is described as a set of linear algebra operations on vectors and matrices.
+Figure \ref{fig:graph_vertex_programming} shows an example where graph (a) is represented as a matrix (b), allowing us to execute vertex multiplication on the matrix transpose (c).
+The above linear algebra operation (c) generates the number of incoming edges for each vertex in the graph.
+All GraphMat algorithms, Breadth First Search (BFS) \cite{BFS}, PageRank \cite{PageRank}, Single Source Shortest-Path (SSSP) \cite{SSSP}, etc..
+are expressed in this fashion such that computation can be efficiently accelerated in hardware.
+In FlexGraph, we accelerate the matrix vertex computation on the FPGA, instead of executing it on the multi-core host processor.
+
+\subsection{Intel Heterogeneous CPU-FPGA Platform}
+
+\subsection{Collaborative CPU-FPGA Computation}
+
 \section{FlexGraph Architecture}
@@
 
-\subsection{FlexGraph Computing Model}
+\subsection{The DCSC Matrix Format}
 
 \begin{figure}[htbp]
-\includegraphics[width=0.4\textwidth]{figures/graph_vertex_programming}
-\caption{Graph Vertex Programming}
-\label{fig:graph_vertex_programming}
-\end{figure}
-
-FlexGraph's architecture is based on GraphMat \cite{GraphMat}'s vertex computing paradigm where a graph algorithm is described as a linear algebra set of operations operating on sparse matrix with input vertices.
-Figure \ref{fig:graph_vertex_programming} shows an example where graph (a) is converted into matrix (b), allowing us to execute operations on the graph using simple matrix vertex multiplication (c).
-This above example will to generate the number of incoming edges for each vertex.
-All GraphMat algorithms, Breadth First Search (BFS) \cite{BFS}, PageRank \cite{PageRank}, Single Source Shortest-Path (SSSP) \cite{SSSP}, etc..
-are expressed in this fashion such that computation can be efficiently accelerated in hardware.
-In FlexGraph, we accelerate the matrix vertex computation on the FPGA, instead of executing it on the multi-core host processor.
-
-\subsection{The Doubly Compressed Sparse Matrix}
-
-\begin{figure}[htbp]
+\centering
 \includegraphics[width=0.4\textwidth]{figures/DCSC_matrix_format}
@@
 Figure \ref{fig:DCSC_matrix_format} shows a sample matrix A = \{(0,5,a), (0,7,b), (6,3,c), (7,1,d)\} with four non-zero edges (src, dst, weight) encoded using DCSC versus the conventional CSC \cite{CSC} format.
-It is encoded as follows: The CSC encoding for this matrix is as follows: The \textit{Cols} buffer holds the column ranges stored as incremental distances from first value, which allows efficient packing of consecutive non-zero columns.
-The limitation of this scheme is that it entries are replicated to skip empty columns.
-\textit{Rows} and \textit{Values} hold the corresponding rows and values referenced by a given non-empty column.
 The traversal over CSC requires accessing all consecutive column ranges \textit{(cols[i+2] - cols[i])} even though most of the distances are empty.
@@
 
-\subsection{Sparse Matrix Multiplication Kernel}
+\subsection{The Sparse Matrix Multiplication Kernel}
 
-\begin{lstlisting}[language=Python, caption=Pseudocode for SPMV kernel]
+FlexGraph architecture implements a custom Sparse Matrix Vertex Multiplication (SPMV) kernel.
+Its architecture diverges from conventional implementations in two ways: First, it uses the dirty masks from  the input vertices to select the edges to process and return a new dirty masks capturing the intersection of the active input vertices and the non-zero edges actually visited.
+Second, it matrix data structure (DCSC) contains an additional indirection buffer for accessing the matrix columns.
+These two properties present unique performance challenges when designing the accelerator.
+Listing 1 shows the pseudo-code of the SPMV kernel.
+The program iterates through each column and fetch the corresponding input vertex active mask to check if the column should be process, then access all the non-zero rows of the matrix for that column to evaluate each edge.
+In the code's comments on the right are highlighted the different types performance hogs present; semi-random memory accesses (*), fully-random memory accesses (!) and control branches (?).
+The memory accesses for the columns and rows data (lines 4 and 9) are semi-random because only the access cannot be predicted and after that the address is simply incremented.
+FlexGraph's architecture attempts to address some of those performance hogs using several optimizations detailed in section 4.
+
+\begin{lstlisting}[language=Python, caption=Pseudo-code for SPMV kernel]
 def SPMV_kernel(x_values, x_activemask) :
   y_values = {0}, y_activemask = {0}
-    for col in cols :                        # +	
-      (a_x, rows) = coldata[col]             # *
-      x_value  = x_values[a_x]               # *
-      x_active = x_activemask[a_x]           # *
+    for col in (c_start, c_end) :            	
+      (a_x, r_start, r_end) = coldata[col]   # *
+      x_value  = x_values[a_x]               # !
+      x_active = x_activemask[a_x]           # !
       if (x_active):                         # ?
-        for row in rows:                     # +
+        for row in (r_start, r_end):         
           (a_y, a_value) = rowdata[row]      # *
-          y_values[a_y] += a_value * x_value # *
-          x_activemask[a_y] = true           # *
+          y_values[a_y] += a_value * x_value # !
+          x_activemask[a_y] = true           # !
   return (y_values, y_activemask)
@@
 
-\subsection{FlexGraph Micro-architecture}
+\subsection{FlexGraph Microarchitecture}
+
+Externally, FlexGraph input and output signals implement an Accelerator Functional Unit (AFU) interface defined by Intel's Accelerator Abstraction Layer (AAL) \cite{Intel-FPGA}.
+AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL software runnign on the host processor.
+Listing 2 shows AAL device interface implementation using Cocoh's API.
+Our FlexGraph Accelerator's class in Cocoh simply derives from this interface and override the \textit{initialize()} function to provide its implementation and Cocoh takes care of generating the corresponding verilog module.
+It implements three input signals \textit{(start, qpi\_in, ctx)} and two output signals \textit{(qpi\_out, done)}.
+The AFU starts execution when the \textit{start} signal is asserted and communicate completion by asserting the \textit{done} signal.
+The \textit{ctx} signal provides application's specific data like constants and buffers address in the case of FlexGraph.
+The \textit{qpi} in/out signals implement Intel Quick Path Interconnect (QPI) interface \cite{QPI}, providing single channel read/write ports for accessing external shared memory.
+The FPGA socket hosts a 64-byte cache coherent interface (CCI) \cite{CCI} that connects to the host processor's last level cache (LLC) via QPI.
+
+\begin{lstlisting}[language=C++, caption=AAL Device Interface in Cocoh C++]
+class aal_device {
+public: 
+  virtual out_t initialize()(
+    const ch_logic& start, 
+    const qpi::in_t& qpi_in, 
+    const afu_ctx_t& ctx, 
+    qpi::out_t& qpi_out, 
+    ch_logic& done
+  ) const = 0;  
+};
+\end{lstlisting}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.4\textwidth]{figures/microarchitecture}
+\caption{FlexGraph Microarchitecture}
+\label{fig:microarchitecture}
+\end{figure}
+
+Figure \ref{fig:microarchitecture} illustrates FlexGraph accelerator microarchitecture.
+It is comprised of four main module types, a controller, processing elements (PEs), a load store unit (LSU) and vertex caches.
+The main controller is responsible for starting the accelerator, scheduling tasks for execution, reporting hardware counters and terminating the execution.
+The processing elements execute partition tasks assigned to them by the controller and communicate with the LSU to access the matrix and vertex data for their partition.
+They are also responsible for sending their final output result back to memory.
+The LSU is the module responsible for managing external communication between the accelerator and memory via the QPI interface.
+It directly binds to the QPI ports defined in listing 2.
+The vertex caches store intermediate vertex data and active masks for sharing between the processing elements.
+FlexGraph processing pipeline sightly resembles the SPMV execution steps illustrated in listing 1.
+We made several important modifications to it to improve performance.
 
@@
 
+The first optimization we performed in FlexGraph early during the design phase was the support of multiple processing units.
+The reasoning behind it was to extract the maximum bandwidth out of the LSU such that the QPi is always busy processing a request when some processing element are stalled waiting for their data to return.
+the other advantage of parallelization for FlexGraph is to increase the overall accelerator throughput by processing multiple a larger chunk of the workload per unit of time.
+The scheduling of the partitions for execution on the processing elements is controlled by a dispatch unit inside the main controller module.
+The dispatch unit fetches partitions data from the LSU and pass down each partition in a first come first serve fashion to the processing elements.
+
 \subsection{DCSC Matrix partitioning}
 
-\subsection{Precessing Tasks Scheduling}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.4\textwidth]{figures/matrix_partitioning}
+\caption{DCSC Matrix Partitioning}
+\label{fig:matrix_partitioning}
+\end{figure}
 
-\subsection{Synchronising Memory Writes for Dirty Masks}
+To enable the parallelization of FlexGraph tasks, the DCSC sparse matrix is first partitioned into aligned partitions of 32 consecutive rows containing non-zero edges.
+we choose a partition size of 32 mainly to match the 32-bit size of the bitmasks encoding the vertices that are active.
+These masks are used by the software on the host processor to determine the regions of the acceleration's output buffer that have been updated.
+Figure \ref{fig:matrix_partitioning} illustrates sample matrix partitioning in which the non-zero horizontal regions have been broken into two partitions P0 and P1.
+The non-zero column ranges \textit{(c0, c1, c2, c3)} represent the selected chunks covered by each partition.
+Partition P0 will only contain ranges \textit{c0} and \textit{c2} while P2 has non-zeros in all four ranges.
+The partitioning scheme is not ideal because of the workload in-balance that might exist when the number of non-zero varies disproportionally between partitions.
+
+\subsection{Synchronising Memory Accesses}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.35\textwidth]{figures/write_masks_synchronization}
+\caption{Active Masks Write Synchronization}
+\label{fig:write_masks_synchronization}
+\end{figure}
+
+An implementation challenge we faced when supporting multiple processing elements in FlexGraph was the synchronization of write accesses for the output active masks.
+Because the LSU data transfer granularity is 64 byte (matching CCI \cite{CCI}), FlexGraph need to manage shared write accesses to the same block to avoid having a processing element override the content of another one.
+Luckily this doesn't pose any problem for writing out partition output values because the 32 rows within a partition occupy 32 * 4 bytes = 2 x 64 bytes blocks.
+Because the partitioning reinforces 32 rows alignment, each processing element can safely write into their assigned blocks.
+However, writing out the active masks poses a challenge because a single 4-byte mask encodes the active states of all 32 rows in a partition.
+This causes the processing elements to potentially share a single 64-byte block when writing the masks out to memory.
+To alleviate the contention, FlexGraph keep N active 64-byte blocks locally assigned to either one of the N processing elements.
+It assigns an ownership bit mask to each block to track their reference count and the processing element assigned to them.
+It also tracks the current address value assigned to each block.
+When a processing element is ready to write its 4-byte active mask, it passes the mask address and value to the LSU.
+The LSU goes to last active block the processing element previous wrote to and check if the block address matches.
+If so, it simply adds the new content to the existing block.
+If there is no match, it clears its ownership bit and flushes the block to memory if the ownership mask goes to zero.
+Then it looks up the other active block if anyone already has the address to use it, otherwise if it acquires the last unused block.
+Figure \ref{fig:write_masks_synchronization} shows a simplified illustration of the scheme with two processing elements.
+We don't need to keep more than N blocks in local storage for this scheme to work because the block address references are always incremental and never regress, meaning that there is always going to be free block available to use.
+It is important to also point out that for this scheme to work, the operation has to be atomic.
+FlexGraph has a single communication channel between all processing elements and the LSU which does some round robin arbitration and blocks the write mask request until it is committed.
+We also investigated an alternative solution to avoid this synchronization, which is simply increasing the partition size to 16 * 32 rows, allowing the aggregated active masks to occupy a full 64-byte block.
+They were two major issues with this approach.
+Firstly, the total size of all resident partitions will be too large if we support multiple processing elements.
+Secondly, it will worsen the workload imbalance that is already present with 32 rows.
 
@@
 
+This section describes some optimizations we added to FlexGraph to improve its performance.
+
 \subsection{Optimizing Memory Accesses via Stream Buffers}
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.4\textwidth]{figures/stream_buffers}
+\caption{Stream Buffers}
+\label{fig:stream_buffers}
+\end{figure}
+
+As stated before, FlexGraph's LSU transfer granularity is 64 bytes, which is much larger than the 4 bytes of elements accessed by the processing elements.
+To save on bandwidth, the LSU uses stream buffers that fetch a 64-byte blocks of data but extracts 4-byte elements at the time.
+It is implemented using a fifo structure in the back-end where 64-byte responses from QPI go
+ into.
+In the front-end, there is a temporary 64-byte block that is extracted from the fifo on demand to deliver 4-byte element at the time.
+We employ a shift register to extract the 4-byte element to pass it to the processing element.
+Figure \ref{fig:stream_buffers} shows an illustration of the stream buffer concept.
+
 \subsection{Caching Vertex Values and Masks}
+
+Another optimization we implemented in FlexGraph is the caching of vertex values and masks.
+Looking at the SPMV speudo-code in listing 1, we can observe in lines 5 and 6 that there is a random memory access to buffers \textit{x\_values} and \textit{x\_activemask}.
+FlexGraph employs a stream buffer like concept to consume 64-byte vertex data once they arrive inside the processing element to each iteration loop.
+However, the same block can be referenced by another processing element and sharing them inside the cache structure can save unnecessary memory traffic.
+We implemented two small fully associative caches with first-in-First-out replacement to hold active vertex data during execution.
+The caches are implemented inside the LSU which is responsible for managing them.
+When a vertex request arrives from a processing element, the LSU first looks up the cache if the block is already present and return it.
+If the block is not there, it sends the request to QPI.
+QPI output interface support a 14-bit metadata field that is used to identify the block once it is returned.
+We use that field to store the index of the block such that we can compute the cache tag when it arrives.
+Because the cache is small, we implemented a one cycle tag lookup for the LSU to known if the block is present and accessible.
+
+\subsection{Executing Non-blocking Memory writes}
+
+FlexGraph memory writes are all non-blocking, this applies to both the output values and active masks.
+This allows the processing elements to push their write request and resume execution while the LSU is processing them.
+Upon write responses from the QPI channel, the LSU internally keeps track of the count of outstanding requested to ensure that all writes have completed.
+The QPI interface exposes two write response ports by which the memory replies could be sent.
+This allows servicing two responses simultaneously.
+The LSU implements a mechanism to process both channels simultaneously when they are active.
+At completion time when all processing elements are done executing the current run, the controller waits for all outstanding write requests to complete before asserting the \textit{done} signal.
+
+\section{FlexGraph Software-Hardware Codesign}

### 2017-11-15 paper/related_work.tex
backup

@@
 \section{Related Work}
+
+In this section, we discuss prior works related to this project.
+
+\subsection{GraphMat}
+
+GraphMat \cite{GraphMat} is a high-performance Graph Analytics Framework for multi-core CPU platforms.
+It defines a C++ template-based vertex programming model for describing various graph algorithms in software and use Sparse Matrix Vector Multiplication (SPMV) as underlying compute kernel for efficient parallel processing.
+GraphMat also uses a doubly-compressed DCSC \cite{DCSC} matrix format for storage and compute efficiency.
+FlexGraph's programming model is an adaptation of GraphMat's model for collaborative computation  with an FPGA accelerator.
 
@@
 
+Graphicionado \cite{Graphicionado} is a much recent project that is similar to FlexGraph.
+In this work, the authors implemented a graph accelerator targeting Intel Xeon platform.
+Graphicionado also uses a vertex programming model of graphs computation similar to GraphMat \cite{GraphMat}.
+It is a more specialized accelerator compared to FlexGraph in that it implements the entire graph algorithm in hardware, sacrificing flexibility for efficiency.
+Although the design allows some reconfigurability using FPGA, it still add many inconveniences compared to software.
+FlexGraph on the other hand only accelerate the graph edges computation in hardware, most of the algorithm is still written expressed in C++ and runs on the host processor like in GraphMat.
+FlexGraph reconfigurability is minimal and restricted to the SPMV module (changing the matrix datatype or the reduce operation).
+Graphicionado main architecture strength is their efficient optimizations of the graph access patterns to reduce stalls inside the pipeline.
+
 \subsection{GraphOps}
 
-\subsection{Tesseract}
+GraphOps \cite{GraphOps} is another graph accelerator also targeting FPGAs like FlexGraph.
+In this work, the authors propose an accelerator architecture that efficiently process a graph in memory encoded using a proposed locality-optimized scheme.
+Their main contribution is the graph storage representation which provides an efficient memory access pattern.
+A drawback of their proposal is that fact that the host processor has to pre-process the graph before the accelerator can consume it, which can add some considerable latency, considering that Graph Analytics algorithms tend to update the graph frequently.

### 2017-11-15 common.h

@@ -2,10 +2,11 @@
 
 #include <string.h>
 #include <cash.h>
-#include <../htl/queue.h>
-#include <../htl/xbar.h>
-#include <../htl/arbiter.h>
-#include <../htl/counter.h>
+#include <htl/queue.h>
+#include <htl/xbar.h>
+#include <htl/arbiter.h>
+#include <htl/counter.h>
+#include <htl/pipe.h>
 #include "aal.h"
 
 using namespace ch::core;
@@ -92,7 +93,7 @@ __struct (ch_pe_hwcntrs_t, (
   (ch_bit32) min_latency,
   (ch_bit32) max_latency,
   (ch_bit32) total_latency,
-  (ch_bit32) execute_latency,
+  (ch_bit32) execute_stalls,
   (ch_bit32) a_xindices_stalls,
   (ch_bit32) x_masks_stalls,
   (ch_bit32) a_startys_stalls,

### 2017-11-15 cpu.cpp

@@ -357,7 +357,7 @@ void cpu_device::dump_stats(ch_tick t) {
     DbgPrint(0, "PE%d average latency = %d ticks\n", i, __safediv((uint32_t)hwcntrs.total_latency, (uint32_t)hwcntrs.num_partitions));
     DbgPrint(0, "PE%d minimum latency = %d ticks\n", i, (uint32_t)hwcntrs.min_latency);
     DbgPrint(0, "PE%d maximum latency = %d ticks\n", i, (uint32_t)hwcntrs.max_latency);
-    DbgPrint(0, "PE%d execution latency = %d ticks\n", i, (uint32_t)hwcntrs.execute_latency);
+    DbgPrint(0, "PE%d execution latency = %d ticks\n", i, (uint32_t)hwcntrs.execute_stalls);
     DbgPrint(0, "PE%d matrix xindices stalls = %d\n", i, (uint32_t)hwcntrs.a_xindices_stalls);
     DbgPrint(0, "PE%d matrix ystarts stalls = %d\n", i, (uint32_t)hwcntrs.a_startys_stalls);
     DbgPrint(0, "PE%d matrix values stalls = %d\n", i, (uint32_t)hwcntrs.a_values_stalls);

### 2017-11-15 dcsc_walk.cpp

@@ -1,0 +1,485 @@
+#include "dcsc_walk.h"
+
+using namespace spmv;
+using namespace spmv::accelerator;
+
+__enum (ch_walk_state, 4, (
+  (get_partition, 0),
+  (get_a_xindex, 1),
+  (get_a_starty, 2),
+  (wait_for_a_starty, 3),
+  (get_a_endy, 4),
+  (wait_for_a_endy, 5),
+  (check_x_mask, 6),
+  (get_x_mask, 7),
+  (wait_for_x_mask, 8),
+  (get_x_value, 9),
+  (get_a_yindex, 0xA),
+  (get_a_value, 0xB),
+  (wait_for_data, 0xC),
+  (execute, 0xD),
+  (next_column, 0xE),
+  (end_partition, 0xF)
+));
+
+spmv_dcsc_walk::spmv_dcsc_walk(uint32_t id)
+  : id_(id)
+  , xmblock_addr_(PTR_MAX_VALUE) // default initialize to max
+  , xvblock_addr_(PTR_MAX_VALUE) // default initialize to max{
+{
+  //--
+  io.ctrl.hwcntrs = hwcntrs_;
+
+  //--
+  axbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
+  asbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
+  aybuf_.io.enq.data = io.lsu.rd_rsp.data.data;
+  avbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
+  xvbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
+  xmbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
+
+  //--
+  axbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_xindices);
+  asbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_startys);
+  aybuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_yindices);
+  avbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_values);
[…]

### 2017-11-15 dcsc_walk.h

@@ -1,0 +1,98 @@
+#pragma once
+
+#include "common.h"
+#include "lsu.h"
+#include "pe.h"
+
+namespace spmv {
+namespace accelerator {
+
+enum {
+  AXBUF_SIZE = 2,
+  ASBUF_SIZE = 2,
+  AYBUF_SIZE = 32,
+  AVBUF_SIZE = 32,
+  XVBUF_SIZE = 2,
+  XMBUF_SIZE = 2,
+};
+
+using ch_axbuf_t = ch_queue<ch_block, AXBUF_SIZE>;
+using ch_asbuf_t = ch_queue<ch_block, ASBUF_SIZE>;
+using ch_aybuf_t = ch_queue<ch_block, AYBUF_SIZE>;
+using ch_avbuf_t = ch_queue<ch_block, AVBUF_SIZE>;
+
+using ch_xvbuf_t = ch_queue<ch_block, XVBUF_SIZE>;
+using ch_xmbuf_t = ch_queue<ch_block, XMBUF_SIZE>;
+
+__struct (ch_pe_start_req_t, (
+  __in(ch_dcsc_part_t) part
+));
+
+__inout (ch_ctrl_pe_io, (
+  (ch_deq_io<ch_pe_start_req_t>) start,
+  __in(ch_bit64) timer,
+  __out(ch_pe_hwcntrs_t) hwcntrs
+));
+
+class spmv_dcsc_walk {
+public:
+
+  __io(
+    (ch_ctrl_pe_io) ctrl,
+    (ch_flip_t<ch_walker_lsu_io>) lsu,
+    (ch_enq_io<ch_pe_req_t>) pe
+  );
+
+  spmv_dcsc_walk(uint32_t id);
+  ~spmv_dcsc_walk();
+
+  void describe();
+
+private:
+
+  uint32_t id_;
+
+  ch_seq<ch_ptr> col_curr_;
+  ch_seq<ch_ptr> col_end_;
+
+  ch_seq<ch_ptr> a_xindex_;
+
+  ch_seq<ch_ptr> row_curr_;
+  ch_seq<ch_ptr> row_end_;
+
+  ch_seq<ch_ptr> xmblock_addr_;
+  ch_seq<ch_ptr> xvblock_addr_;
+
+  ch_seq<ch_ptr>    row_blk_curr_;
+  ch_seq<ch_ptr>    row_blk_end_;
+  ch_seq<ch_bit<6>> row_blk_cnt_;
+
[…]

### 2017-11-15 lsu.cpp

@@ -21,8 +21,8 @@ spmv_lsu::spmv_lsu() {
   // bind modules
   //
   for (int i = 0; i < PE_COUNT; ++i) {
-    rd_req_arb_.io.in[i](io.pe[i].rd_req);
-    wr_req_arb_.io.in[i](io.pe[i].wr_req);
+    rd_req_arb_.io.in[i](io.walkers[i].rd_req);
+    wr_req_arb_.io.in[i](io.PEs[i].wr_req);
   }
   rd_req_arb_.io.in[PE_COUNT](io.ctrl.rd_req);
   wr_req_arb_.io.in[PE_COUNT](io.ctrl.wr_req);
@@ -217,9 +217,9 @@ void spmv_lsu::read_rsp_thread() {
 
   //--
   for (int i = 0; i < PE_COUNT; ++i) {
-    io.pe[i].rd_rsp.valid     = io.qpi.rd_rsp.valid & (mdata.owner == PE_ID(i));
-    io.pe[i].rd_rsp.data.data = io.qpi.rd_rsp.data;
-    io.pe[i].rd_rsp.data.type = mdata.type;
+    io.walkers[i].rd_rsp.valid     = io.qpi.rd_rsp.valid & (mdata.owner == PE_ID(i));
+    io.walkers[i].rd_rsp.data.data = io.qpi.rd_rsp.data;
+    io.walkers[i].rd_rsp.data.type = mdata.type;
   }
 
   //--

### 2017-11-15 lsu.h

@@ -73,14 +73,17 @@ __inout (ch_ctrl_lsu_io, (
   __out(ch_bit32) outstanding_writes
 ));
 
-__inout (ch_pe_lsu_io, (
+__inout (ch_walker_lsu_io, (
   (ch_deq_io<ch_lsu_rd_req_t>) rd_req,
-  (ch_deq_io<ch_lsu_wr_req_t>) wr_req,
   (ch_valid_io<ch_lsu_rd_rsp_t>) rd_rsp
 ));
 
-using ch_rd_req_arb_t = ch_arbiter<ch_lsu_rd_req_t, RD_REQ_INPUTS>;
-using ch_wr_req_arb_t = ch_arbiter<ch_lsu_wr_req_t, WR_REQ_INPUTS>;
+__inout (ch_pe_lsu_io, (
+  (ch_deq_io<ch_lsu_wr_req_t>) wr_req
+));
+
+using ch_rd_req_arb_t = ch_xbar_switch<ch_lsu_rd_req_t, RD_REQ_INPUTS>;
+using ch_wr_req_arb_t = ch_xbar_switch<ch_lsu_wr_req_t, WR_REQ_INPUTS>;
 
 class spmv_lsu {
 public:
@@ -89,7 +92,8 @@ public:
     (spmv_aal_device::qpi_io) qpi,
     (ch_ctx_io) ctx,
     (ch_ctrl_lsu_io) ctrl,
-    (ch_vec<ch_pe_lsu_io, PE_COUNT>) pe
+    (ch_vec<ch_walker_lsu_io, PE_COUNT>) walkers,
+    (ch_vec<ch_pe_lsu_io, PE_COUNT>) PEs
   );
 
   spmv_lsu();

### 2017-11-15 pe.cpp

@@ -1,494 +1,137 @@
 #include "pe.h"
+#include "pipew.h"
 
 using namespace spmv;
 using namespace spmv::accelerator;
 
-__enum (ch_exec_state, 5, (
-  get_partition,
-  clear_output,
-  get_a_xindex,
-  get_a_starty,
-  wait_for_a_starty,
-  get_a_endy,
-  wait_for_a_endy,
-  check_x_mask,             
-  get_x_mask,
-  wait_for_x_mask,
-  get_x_value,
-  get_a_yindex,
-  get_a_value,
-  wait_for_data,
-  execute,
-  next_column,
-  write_y_value0,
-  write_y_value1,
-  write_y_mask
+__enum (ch_pe_state, 3, (
+  ready,
+  write_value0,
+  write_value1,
+  write_mask,
+  clear
 ));
 
-spmv_pe::spmv_pe(uint32_t id)
-  : id_(id)
-  , xmblock_addr_(PTR_MAX_VALUE) // default initialize to max
-  , xvblock_addr_(PTR_MAX_VALUE) // default initialize to max
-{
+spmv_pe::spmv_pe(uint32_t id) : id_(id) {
   //--
-  io.ctrl.hwcntrs = hwcntrs_;
-
-  //--
-  axbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
-  asbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
-  aybuf_.io.enq.data = io.lsu.rd_rsp.data.data;
-  avbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
-  xvbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
-  xmbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
-
-  //--
-  axbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_xindices);
-  asbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_startys);
-  aybuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_yindices);
-  avbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_values);
[…]

### 2017-11-15 pe.h

@@ -6,39 +6,25 @@
 namespace spmv {
 namespace accelerator {
 
-enum {
-  AXBUF_SIZE = 2,
-  ASBUF_SIZE = 2,
-  AYBUF_SIZE = 32,
-  AVBUF_SIZE = 32,
-  XVBUF_SIZE = 2,
-  XMBUF_SIZE = 2,
-};
-
-using ch_axbuf_t = ch_queue<ch_block, AXBUF_SIZE>;
-using ch_asbuf_t = ch_queue<ch_block, ASBUF_SIZE>;
-using ch_aybuf_t = ch_queue<ch_block, AYBUF_SIZE>;
-using ch_avbuf_t = ch_queue<ch_block, AVBUF_SIZE>;
-
-using ch_xvbuf_t = ch_queue<ch_block, XVBUF_SIZE>;
-using ch_xmbuf_t = ch_queue<ch_block, XMBUF_SIZE>;
-
-__struct (ch_pe_start_req_t, (
-  __in(ch_dcsc_part_t) part
+__struct (ch_pe_req_t, (
+  (ch_ptr)     a_yindex,
+  (ch_float32) a_value,
+  (ch_float32) x_value,
+  (ch_bool)    is_end
 ));
 
-__inout (ch_ctrl_pe_io, (
-  (ch_deq_io<ch_pe_start_req_t>) start,
-  __in(ch_bit64) timer,
-  __out(ch_pe_hwcntrs_t) hwcntrs
+__struct (ch_pipe_data_t, (
+  (ch_ptr)  a_yindex,
+  (ch_bool) is_end
 ));
 
 class spmv_pe {
 public:
 
   __io(
-    (ch_ctrl_pe_io) ctrl,
-    (ch_flip_t<ch_pe_lsu_io>) lsu
+    (ch_deq_io<ch_pe_req_t>) req,
+    (ch_flip_t<ch_pe_lsu_io>) lsu,
+    __out(ch_bool) is_idle
   );
   
   spmv_pe(uint32_t id);
@@ -50,54 +36,27 @@ private:
 
   using ch_bit480 = ch_bit<480>;
   
-  uint32_t  id_;  
+  uint32_t  id_;
 
-  ch_module<ch_axbuf_t> axbuf_;
-  ch_module<ch_asbuf_t> asbuf_;
-  ch_module<ch_aybuf_t> aybuf_;
-  ch_module<ch_avbuf_t> avbuf_;
-  ch_module<ch_xvbuf_t> xvbuf_;
[…]

### 2017-11-15 pipew.h

@@ -1,0 +1,47 @@
+#pragma once
+
+#include "common.h"
+
+namespace spmv {
+namespace accelerator {
+
+template <typename T, unsigned Delay>
+class pipe_window {
+public:
+  pipe_window(const T& value, const ch_bool& valid) {
+    //--
+    value_[0] = value;
+    for (unsigned i = 1; i < Delay; ++i) {
+      value_[i] = ch_reg(value_[i-1]);
+    }
+
+    //--
+    valid_[0] = valid;
+    for (unsigned i = 1; i < Delay; ++i) {
+      valid_[i] = ch_reg(valid_[i-1]);
+    }
+  }
+
+  ch_bool in_flight(const T& value) const {
+    ch_bool ret = (value_[0] == value) && valid_[0];
+    for (unsigned i = 1; i < Delay; ++i) {
+      ret = (value_[i] == value && valid_[i]) || ch_clone(ret);
+    }
+    return ret;
+  }
+
+  T value() const {
+    return value_[Delay-1];
+  }
+
+  ch_bool valid() const {
+    return valid_[Delay-1];
+  }
+
+private:
+  T       value_[Delay];
+  ch_bool valid_[Delay];
+};
+
+}
+}

### 2017-11-15 spmv.cpp

@@ -14,14 +14,18 @@ __enum (ch_ctrl_state, 3, (
 
 spmv_device::spmv_device() {
   // create the PEs
-  pe_.reserve(PE_COUNT);
+  PEs_.reserve(PE_COUNT);
+  walkers_.reserve(PE_COUNT);
   for (int i = 0; i < PE_COUNT; ++i) {
-    pe_.emplace_back(i);
+    PEs_.emplace_back(i);
+    walkers_.emplace_back(i);
   }
 
   // bind modules
   for (int i = 0; i < PE_COUNT; ++i) {
-    lsu_.io.pe[i](pe_[i].io.lsu);
+    PEs_[i].io.req(walkers_[i].io.pe);
+    lsu_.io.walkers[i](walkers_[i].io.lsu);
+    lsu_.io.PEs[i](PEs_[i].io.lsu);
   }
   lsu_.io.ctx(io.ctx);
   lsu_.io.qpi(io.qpi);
@@ -38,8 +42,8 @@ void spmv_device::describe() {
   // timer
   ch_seq<ch_bit64> timer;
   timer.next = timer + 1;
-  for (auto& pe : pe_) {
-    pe.io.ctrl.timer = timer;
+  for (auto& w : walkers_) {
+    w.io.ctrl.timer = timer;
   }
   
   // main thread
@@ -65,9 +69,9 @@ void spmv_device::main_thread() {
       ch_zext<32>(pbuf_.io.deq.ready);
 
   //--
-  ch_bool all_PEs_ready = pe_[0].io.ctrl.start.ready;
+  ch_bool all_PEs_ready = walkers_[0].io.ctrl.start.ready && PEs_[0].io.is_idle;
   for (int i = 1; i < PE_COUNT; ++i) {
-    all_PEs_ready = ch_clone(all_PEs_ready) & pe_[i].io.ctrl.start.ready;
+    all_PEs_ready = ch_clone(all_PEs_ready) && walkers_[i].io.ctrl.start.ready && PEs_[i].io.is_idle;
   }
       
   //--
@@ -84,7 +88,6 @@ void spmv_device::main_thread() {
       __else (
         // no work to do!
         done.next = true; // set done signal
-        state.next = ch_ctrl_state::ready;
       );
     );
   )
@@ -198,10 +201,10 @@ void spmv_device::dispatch_thread() {
   for (int i = 0; i < PE_COUNT; ++i) {
     sw __case (1+i) (
       // check if PE is ready
[…]

### 2017-11-15 spmv.h

@@ -2,6 +2,7 @@
 
 #include "common.h"
 #include "lsu.h"
+#include "dcsc_walk.h"
 #include "pe.h"
 
 namespace spmv {
@@ -34,7 +35,8 @@ protected:
 
   ch_block get_hwnctrs(const ch_hwcntr_addr& addr);
 
-  std::vector<ch_module<spmv_pe>> pe_;
+  std::vector<ch_module<spmv_dcsc_walk>> walkers_;
+  std::vector<ch_module<spmv_pe>> PEs_;
   ch_module<spmv_lsu> lsu_;
   ch_module<ch_pbuf_t> pbuf_;
 

### 2017-11-15 test_vcache.cpp

@@ -2,11 +2,19 @@
 #include <fstream>
 #include <unistd.h>
 #include "test.h"
+//#include "vcache.h"
 
 using namespace spmv;
-
+//using namespace spmv::accelerator;
 
 int main(int argc, char **argv) {
-  //--
+
+  /*ch_device<spmv_vertex_cache<ch_bit4, 8, 8, 2>> device;
+  ch_simulator sim(device);
+  ch_tick t = sim.reset(0);
+  t = sim.step(t);
+  int ret = 1;
+  assert(!!device.io.enq.ready);*/
+
   return 0;
 }

### 2017-11-15 tmp.h

@@ -1,24 +1,0 @@
-/*{
-  // spmv_device FSM
-  ch_seq<ch_ctrl_state> state;
-     
-  __switch (state) (
-  __case (ch_ctrl_state::ready) (
-    __if (m_start) (
-      state.next = ch_ctrl_state::execute;
-    );
-  )
-  __case (ch_ctrl_state::execute) (
-    __if (m_PE.complete()) (
-      state.next = ch_ctrl_state::write_hwcntrs;
-    );
-  )
-  __case (ch_ctrl_state::write_hwcntrs) (
-    __if (m_LSU.done()) (
-      state.next = ch_ctrl_state::ready;
-    );
-  )
-  __default (
-    state.next = state;        
-  ));
-}*/

### 2017-11-15 vcache.h

@@ -5,88 +5,224 @@
 namespace spmv {
 namespace accelerator {
 
-//
-// Description:
-// this code implements a vertex cache, entries are insert and evicted in fifo order.  
-//
+template <typename T, unsigned N, unsigned TagBits, unsigned LookupCycles = CH_CEILDIV(N, 4)>
+class spmv_vertex_cache {
+private:
 
-template <unsigned SIZE, unsigned DATA_WIDTH, unsigned TAG_WIDTH>
-class vertex_cache {
+  using ch_idx = ch_bit<log2ceil(N)>;
+  using ch_tag = ch_bit<TagBits>;
+
+  __enum (ch_state, 3, (
+    ready,
+    lookup,
+    check_evict,
+    evict,
+    flush
+  ));
+
+  __struct (entry_t, (
+    (ch_tag)    tag,
+    (T)         data
+  ));
+
+  __struct (tag_t, (
+    (ch_bool) valid,
+    (ch_tag)  tag
+  ));
+
+  ch_ram<tag_t, N> tags_;
+  ch_ram<T, N>     data_;
+  ch_seq<ch_idx>   last_used_idx_;
+  ch_seq<ch_tag>   last_used_tag_;
+
+  ch_idx port0_;
+  ch_idx port1_;
+
+  T data_value_;
+  ch_bool data_wenable_;
+
+  ch_tag tag_value_;
+  ch_bool tag_wenable_;
+
 public:
-  
-  enum {
-    address_width = log2ceil(SIZE),
-  };  
-  
-  using ch_addr_t = ch_bit<address_width>;
-  using ch_tag_t  = ch_bit<TAG_WIDTH>;
-  using ch_data_t = ch_bit<DATA_WIDTH>;  
-  
-  void emit() {
-    // emit tag valid test
-    ch_bit1 tag_valid[SIZE];
-    ch_addr_t match_pos[SIZE];
-    tag_valid[0] = (m_query_tag == (ch_tag_t)m_tags[0]);
-    match_pos[0] = 0;
-    for (int i = 1; i < SIZE; ++i) {
-      tag_valid[i] = tag_valid[i-1] | (m_query_tag == (ch_tag_t)m_tags[i-1]);
-      match_pos[i] = ch_select<address_width>((m_query_tag == (ch_tag_t)m_tags[i]), i, match_pos[i-1]);
-    }
-    m_query_valid.next = tag_valid[SIZE-1];
-    m_query_pos.next = match_pos[SIZE-1];        
-    
[…]

### 2017-11-15 wrcache.h

@@ -12,16 +12,18 @@ private:
   using ch_idx = ch_bit<log2ceil(N)>;
   using ch_tag = ch_bit<TagBits>;
 
-  enum {
-    MAX_TAG_VALUE = (1 << TagBits) - 1,
-  };
-
   __enum (ch_state, 3, (
     ready,
     lookup,
     check_evict,
     evict,
     flush
+  ));
+
+  __struct (entry_t, (
+    (ch_bit<N>) owner,  // one-hot representation
+    (ch_tag)    tag,
+    (T)         data
   ));
 
   __struct (tag_t, (
@@ -46,12 +48,6 @@ private:
   ch_seq<ch_idx> counter_;
 
 public:
-
-  __struct (entry_t, (
-    (ch_bit<N>) owner,  // one-hot representation
-    (ch_tag)    tag,
-    (T)         data
-  ));
 
   __io (
     (ch_deq_io<entry_t>) enq,
@@ -91,7 +87,7 @@ public:
         owned_block_idx = ch_select((tags_[i].owners & lookup_data.owner) != 0, i, ch_clone(owned_block_idx));
         free_block_idx  = ch_select(tags_[N-1-i].owners == 0, N-1-i, ch_clone(free_block_idx));
         if (0 == (i % step)) {
-          lookup_data = enq_data; // io.enq.data is not avaiable beyond first cycle, use cbackup value
+          lookup_data = enq_data; // io.enq.data is not avaiable beyond first cycle, use backup value
           match_block_idx = ch_reg(ch_clone(match_block_idx));
           owned_block_idx = ch_reg(ch_clone(owned_block_idx));
           free_block_idx  = ch_reg(ch_clone(free_block_idx));
@@ -118,7 +114,7 @@ public:
           port0_ = last_used_idx_;
           __if (io.enq.data.tag == last_used_tag_
              && (tags_[port0_].owners & io.enq.data.owner) != 0) (
-            ch_print("@ fast hit!");
+            //ch_print("@ fast hit!");
             data_wenable_ = true;
             tag_wenable_  = true;
             data_value_ = data_[port0_] | io.enq.data.data;
@@ -139,7 +135,7 @@ public:
           port1_ = match_block_idx;
           __if (tags_[port1_].owners != 0
             &&  tags_[port1_].tag == enq_data.tag) (
-            ch_print("@ lookup match!");
+            //ch_print("@ lookup match!");
             // append data to matching block
[…]

### 2017-11-15 paper/mechanism.tex
backup

@@
 \begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=Graph Processing Kernel]
-GraphKernel (A, P, iterations) :
+GraphKernel (G, P, iterations) :
   for_each iterations
     x := Assign(P)
-  	y := Process(A,x)
+  	y := Process(G,x)
   	P := Apply(y)	  
@@
 
-Incoming Edges := GraphKernel(A, identity, 1):
+e.g.: 
+	FindIncomingEdges (A, identity, 1):
 	with Assign(P)    := P

### 2017-11-15 paper/mechanism.tex
backup

@@
 \section{Background and Motivation}
+
+In this section, we provide the background and motivation for this work.
 
@@
 
-\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=Graph Processing Kernel]
-GraphKernel (G, P, iterations) :
-  for_each iterations
-    x := Assign(P)
-  	y := Process(G,x)
-  	P := Apply(y)	  
-  end_for
+\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=GraphMath Processing Model]
+GraphProgram(%\textit{G}%, %\textit{P}%, %\textit{N}%) :
+  for %\textit{N}% iterations :
+    x := Assign(%\textit{P}%)
+  	y := Process(%\textit{G}%,x)
+  	%\textit{P}% := Apply(y)
 
-e.g.: 
-	FindIncomingEdges (A, identity, 1):
-	with Assign(P)    := P
-	     Process(A,x) := A * x
+e.g. FindIncomingEdges(%\textit{G}%=A, %\textit{P}%=%\textit{\textbf{I}}%, %\textit{N}%=1) :
+       Assign(%\textit{P}%)    := %\textit{P}%
+  with Process(%\textit{G}%,x) := %\textit{G}% * x
        Apply(y)     := y
@@
 
-FlexGraph's architecture is based on GraphMat \cite{GraphMat}'s vertex computing paradigm where a graph algorithm is described as a set of linear algebra operations on vectors and matrices.
+A large variety of programming models have been proposed for describing Graph algorithms, expressing computation using vertex operations on matrices \cite{GraphMat} \cite{GraphLab}, \cite{Pregel}, \cite{MapGraph} \cite{GraphX} task-based models \cite{Galois} or domain-specific languages \cite{GreenMarl}.
+The vertex programming model has show great adoption for its ease of abstraction for describing algorithms using linear algebra and its efficient computation on commodity multi-core processors.
+Figure \ref{fig:sample_graph} shows a simple 3 vertices graph example with its corresponding 3x3 matrix representation.
+For every edge between a source and destination vertex, the corresponding row and column entry in the matrix is activate.
+
+ deal  FlexGraph's architecture is based on GraphMat \cite{GraphMat}'s vertex computing paradigm where a graph algorithm is described as a set of linear algebra operations on vectors and matrices.
 Figure \ref{fig:graph_vertex_programming} shows an example where graph (a) is represented as a matrix (b), allowing us to execute vertex multiplication on the matrix transpose (c).

### 2017-11-16 paper/mechanism.tex
backup

@@
 \centering
-\includegraphics[width=0.4\textwidth]{figures/sample_graph}
+\includegraphics[width=0.5\textwidth]{figures/sample_graph}
 \caption{Sample Graph Representation}
@@
 
-\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=GraphMath Processing Model]
+\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=Graph Vertex Processing Model]
 GraphProgram(%\textit{G}%, %\textit{P}%, %\textit{N}%) :
@@
   	y := Process(%\textit{G}%,x)
-  	%\textit{P}% := Apply(y)
+  	%\textit{P}% := Apply(%\textit{P}%,y)
 
@@
   with Process(%\textit{G}%,x) := %\textit{G}% * x
-       Apply(y)     := y
+       Apply(%\textit{P}%,y)   := y
 \end{lstlisting}
@@
 A large variety of programming models have been proposed for describing Graph algorithms, expressing computation using vertex operations on matrices \cite{GraphMat} \cite{GraphLab}, \cite{Pregel}, \cite{MapGraph} \cite{GraphX} task-based models \cite{Galois} or domain-specific languages \cite{GreenMarl}.
-The vertex programming model has show great adoption for its ease of abstraction for describing algorithms using linear algebra and its efficient computation on commodity multi-core processors.
+The vertex programming model has show great adoption for its ease of abstraction for describing algorithms using Linear Algebra and its efficient computation on commodity multi-core processors.
+
 Figure \ref{fig:sample_graph} shows a simple 3 vertices graph example with its corresponding 3x3 matrix representation.
 For every edge between a source and destination vertex, the corresponding row and column entry in the matrix is activate.
+For instance, the matrix entry in row 0 and column 1 represents the edge between source \textit{v0} and destination \textit{v1}.
+using the matrix representation, we can apply an algorithm on the graph to calculate the total number of incoming edges at each vertex.
+This algorithm can be expressed using a simple matrix-vector operation as shown in equations (1) and (2) in Figure \ref{fig:sample_algorithm}.
+Several graph algorithms, including Breadth First Search (BFS) \cite{BFS}, PageRank \cite{PageRank}, Single Source Shortest-Path (SSSP) \cite{SSSP}, can be described similarly using the vertex programming model \cite{GraphMat}.
 
- deal  FlexGraph's architecture is based on GraphMat \cite{GraphMat}'s vertex computing paradigm where a graph algorithm is described as a set of linear algebra operations on vectors and matrices.
-Figure \ref{fig:graph_vertex_programming} shows an example where graph (a) is represented as a matrix (b), allowing us to execute vertex multiplication on the matrix transpose (c).
-The above linear algebra operation (c) generates the number of incoming edges for each vertex in the graph.
-All GraphMat algorithms, Breadth First Search (BFS) \cite{BFS}, PageRank \cite{PageRank}, Single Source Shortest-Path (SSSP) \cite{SSSP}, etc..
-are expressed in this fashion such that computation can be efficiently accelerated in hardware.
-In FlexGraph, we accelerate the matrix vertex computation on the FPGA, instead of executing it on the multi-core host processor.
+Listing 1 shows the three stages of a generalized graph vertex processing model - \textit{Assign}, \textit{Process} and \textit{Apply}, given a graph \textit{G}, some vertex properties \textit{P} and an iteration count \textit{N}.
+The \textit{Assign} stage generates an input vector \textit{x} using the vertex properties \textit{P}.
+The \textit{Process} stage applies the input vector \textit{x} to the graph \textit{G} and generates an output vector \textit{y}.
+The\textit{ Apply} gather the resulting output vector to update the vertex properties for the next iteration.
+The graph program executes the three stages for several iterations until it converges.
+A direct mapping of our incoming edges algorithm to this processing model is also provided in Listing 1, where the identity vector is used for the vector properties.
+The \textit{Assign} and \textit{Apply} stages of this program are simple identity operators, while the \textit{Process} stage perform a matrix-vector multiplication.
+The input graph can be represented as a sparse matrix and the computation done using a Sparse Matrix-Sparse Vector Multiplication (SpMSpV) kernel which can be accelerated on multi-core CPUs \cite{GraphMat} or on GPUs \cite{MapGraph}.
+FlexGraph's objective is to provide a more energy efficient graph processing backend using a FPGA for hardware acceleration \cite{Catapult}.
 
 \subsection{Intel Heterogeneous CPU-FPGA Platform}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/harp_arch}
+\caption{Intel HARP Architecture}
+\label{fig:harp_arch}
+\end{figure}
 

### 2017-11-16 cpu.cpp
backup

@@ -148,6 +148,8 @@ void cpu_device::check_output(const float* values, const uint32_t* masks, bool d
       if (mask != y_mask) {
         printf("*** dirty mask mismatch in partition%d, : actual=0x%x, expected=0x%x\n", p, mask, y_mask);
         ++errors;
+      } else {
+        printf("OK: dirty mask mismatch in partition%d, : actual=0x%x\n", p, mask);
       }
       for (int i = 0; i < 32; ++i) {
         if (y_mask & (1 << i)) {
@@ -156,6 +158,8 @@ void cpu_device::check_output(const float* values, const uint32_t* masks, bool d
           if (value != value_ref) {
             printf("*** value mismatch in partition%d, row=%d: actual=%f, expected=%f\n", p, i, value, value_ref);
             ++errors;
+          } else {
+            printf("Ok: partition%d, row=%d: actual=%f\n", p, i, value);
           }
         }
       }

### 2017-11-16 pe.cpp

@@ -27,14 +27,16 @@ void spmv_pe::describe() {
   ch_seq<ch_pe_state> state;
 
   //--
+  ch_bool do_enq = io.req.valid && io.req.ready;
+
+  //--
   auto mult_value = ch_fmult<ALTFP_SP_MULT>(io.req.data.a_value, io.req.data.x_value);
   mult_pipe_.io.enq.data.a_yindex = io.req.data.a_yindex;
   mult_pipe_.io.enq.data.is_end = io.req.data.is_end;
-  mult_pipe_.io.enq.valid = io.req.valid;
+  mult_pipe_.io.enq.valid = do_enq;
 
   //--
-  auto mult_yindex = mult_pipe_.io.deq.data.a_yindex;
-  auto y_raddr = (mult_yindex & 0x1f).slice<5>();
+  auto y_raddr = (mult_pipe_.io.deq.data.a_yindex & 0x1f).slice<5>();
   auto add_value = ch_fadd<ALTFP_SP_ADD_SUB>(mult_value, y_values_[y_raddr]);
   add_pipe_.io.enq.data.a_yindex = mult_pipe_.io.deq.data.a_yindex;
   add_pipe_.io.enq.data.is_end = mult_pipe_.io.deq.data.is_end;
@@ -45,10 +47,10 @@ void spmv_pe::describe() {
   auto y_waddr_mask = 1_b32 << y_waddr_;
 
   //--
-  __if (io.req.valid && !add_pipe_.io.deq.valid) (
+  __if (do_enq && !add_pipe_.io.deq.valid) (
     pending_reqs_.next = pending_reqs_ + 1;
   )
-  __elif (!io.req.valid && add_pipe_.io.deq.valid) (
+  __elif (!do_enq && add_pipe_.io.deq.valid) (
     pending_reqs_.next = pending_reqs_ - 1;
   );
 

### 2017-11-16 paper/mechanism.tex
backup

@@
 \end{figure}
+
+With the rising adoption of FPGAs in production data centers \cite{Catapult}, there is a need to increase its ecosystem by making them more energy efficient and accessible to both the users and designers.
+Modern CPU-FPGA platforms \cite{CPU-FPGA} provide tightly-coupled shared-memory CPU-FPGA integration \cite{CAPI} \cite{CCI}, making then easier to program and providing and efficient for collaborative execution.
 

### 2017-11-16 paper/mechanism.tex
backup

--- 2017-11-16 paper/mechanism.tex
@@
 
-\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, language=Python, caption=Graph Vertex Processing Model]
+\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, caption=Graph Vertex Processing Model]
 GraphProgram(%\textit{G}%, %\textit{P}%, %\textit{N}%) :
-  for %\textit{N}% iterations :
+  for_each %\textit{N}% :
     x := Assign(%\textit{P}%)
@@
            Assign(%\textit{P}%)    := %\textit{P}%
-  with Process(%\textit{G}%,x) := %\textit{G}% * x
+     where Process(%\textit{G}%,x) := %\textit{G}% * x
            Apply(%\textit{P}%,y)   := y
@@
 \end{figure}
+
+With the rising adoption of FPGAs in production data centers \cite{Catapult}, there is a need to increase its ecosystem by making them more energy efficient as well as accessible to both the users and designers.
+Modern CPU-FPGA platforms \cite{CPU-FPGA} provide tightly-coupled shared-memory CPU-FPGA integration \cite{CAPI} \cite{CCI}, making then easier to program and efficient for hardware acceleration.
+Figure \ref{fig:harp_arch} shows an overview architecture of the Intel Xeon-FPGA platform prototype that we used in our evaluation.
+It has a dual-socket system, one containing a 10-core Intel Xeon E5-2680 CPU and the other containing an Altera Stratix V 5SGXEA FPGA operating at 200 MHz.
+The two sockets are connected via a 6.4 GT/s QPI \cite{QPI} link for data transfer.
+The FPGA's area is partitioned into two main blocks - the Application Function Unit (AFU) or 'Green' bitstream allocated for the custom accelerator and the 'Blue' bitstream implementing the Cache Coherent Interface (CCI) for the accelerator.
+The CCI runtime implements address translation (1024 page table entries) with request re-ordering for a total of 2 GB of addressable memory.
+The AFU memory interface is 64-byte cache-line addressable and it is up to the accelerator to extend the memory interface if fine grained (e.g.
+single byte) access is desired.
 

### 2017-11-16 paper/mechanism.tex
backup

@@
 The \textit{Assign} and \textit{Apply} stages of this program are simple identity operators, while the \textit{Process} stage perform a matrix-vector multiplication.
-The input graph can be represented as a sparse matrix and the computation done using a Sparse Matrix-Sparse Vector Multiplication (SpMSpV) kernel which can be accelerated on multi-core CPUs \cite{GraphMat} or on GPUs \cite{MapGraph}.
+The input graph can be represented as a compressed sparse matrix to only store the non-zero entries of the graph.
+This data structure enables accelerated computation, often using a Sparse Matrix-Sparse Vector Multiplication (SpMSpV) kernel to target multi-core CPUs \cite{GraphMat} or GPUs \cite{MapGraph}.
+The computation that cannot be accelerated is simply executed on the general-purpose host CPU.
 FlexGraph's objective is to provide a more energy efficient graph processing backend using a FPGA for hardware acceleration \cite{Catapult}.
@@
 
+\begin{figure*}
+\centering
+\begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{figures/offload_model}
+    \caption{Offload model}
+    \label{fig:offload_model}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+    \includegraphics[width=\textwidth]{figures/collaborative_model}
+    \caption{Collaborative model}
+    \label{fig:collaborative_model}
+\end{subfigure}
+\caption{Shared Memory Compute Models}
+\label{fig:compute_model}
+\end{figure*}
+
 With the rising adoption of FPGAs in production data centers \cite{Catapult}, there is a need to increase its ecosystem by making them more energy efficient as well as accessible to both the users and designers.
@@
 The FPGA's area is partitioned into two main blocks - the Application Function Unit (AFU) or 'Green' bitstream allocated for the custom accelerator and the 'Blue' bitstream implementing the Cache Coherent Interface (CCI) for the accelerator.
-The CCI runtime implements address translation (1024 page table entries) with request re-ordering for a total of 2 GB of addressable memory.
-The AFU memory interface is 64-byte cache-line addressable and it is up to the accelerator to extend the memory interface if fine grained (e.g.
+The CCI runtime implements a 64 KB direct-mapped cache to support address translation (1024 page table entries) with request re-ordering for a total of 2 GB of addressable memory.
+The AFU memory interface is 64-byte cache-line addressable and it is up to the accelerator to extend the interface if fine grained (e.g.
 single byte) access is desired.
@@
 \subsection{Collaborative CPU-FPGA Computation}
+
+The prevalent execution model for accelerator design is the offload model \cite{Accelerators} where computation takes place on discrete copies of a shared resource that are optimized for efficient access by the target device.
+In this model of computation, the host processor applies some layout translation on the data before making it available to the client accelerator (see Figure \ref{fig:offload_model}).
+The additional latency of the translation process is generally mitigated using optimization techniques such as batching and double-buffering.
+The offload model is well suited for discrete memory systems where the necessary data transfer can be coupled with layout translation.
+One attractive application of shared-memory systems is the enabling of efficient collaborative computation (see Figure \ref{fig:collaborative_model}) in which all attached compute elements can access the same shared resource and modifying it during the course of the program execution.
+It is particularly ideal if the data layout provides efficient access by the compute devices.
+Contrarily to existing Graph Analytics Accelerators \cite{Graphicionado} \cite{GraphOps} that employ a custom optimized graph data structure for computation, Flexgraph uses a sparse matrix format for computation.
+This decision provides several advantages - first, it is storage efficient because the graph's physical memory is shared - second, it enables collaborative computation with the host CPU, leveraging the large ecosystem of matrix-based frameworks \cite{CombBlas} \cite{Pegasus} - third, it decouples the software and hardware design, enabling them to scale independently.
 

### 2017-11-16 paper/introduction.tex
backup

@@
 3.	We introduce hardware primitives for unstructured fine-grained accesses on coherent shared memory architectures.\\
-4.	A domain specific accelerator with single source programming and reconfiguration environment providing a design-efficient alternative to High-Level Synthesis.
+4. 	A Doubly-Compressed Sparse Matrix-Sparse Vector Multiplication Accelerator implementation on FPGA. 
+5. 	A domain specific accelerator with single source programming and reconfiguration environment providing a design-efficient alternative to High-Level Synthesis.
 
 The remainder of this paper is organized as follows:
-Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 describe FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 5 describes the experimental setup, section 6 describes our results analysis, section 7 describes the related work, section 8 summarizes our main contribution and results.
+Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 details the various architecture optimizations we used in FlexGraph, section 5 describes FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 6 describes the experimental setup, section 7 describes our results analysis, section 8 describes the related work, section 9 summarizes our main contribution and results.

### 2017-11-16 paper/mechanism.tex
backup

@@
 \centering
-\includegraphics[width=0.4\textwidth]{figures/DCSC_matrix_format}
+\includegraphics[width=0.5\textwidth]{figures/DCSC_matrix_format}
 \caption{DCSC hyperspace matrix format}
@@
 
-Most Data Analytics graphs are represented in a sparse matrix format to save memory space due to the large amount of empty edges that exist.
-FlexGraph's matrices are stored using a Doubly Compressed Sparse Column (DCSC) \cite{DCSC} format.
+FlexGraph uses the Doubly Compressed Sparse Column (DCSC) \cite{DCSC} format as underlying graph data structure.
 The format allows minimal traversal into the sparse matrix structure, saving necessary memory bandwidth when fetching empty columns.
@@
 DCSC saves on bandwidth by using an additional indirection  buffer inside its structure to only encode non-zero columns.
-It is important to note that the saving this format introduced additional space and compute overhead that only pays off when the matrix is very large and sparse, a prevalent characteristic of Graph Analytics datasets.
+It is important to note that this indirection can introduce additional storage overhead and access latency if the matrix is too small or not sparse enough.
+However, Graph Analytics datasets are generally very large and hyper-sparse which eliminate the DCSC format overhead.
 
-\subsection{The Sparse Matrix Multiplication Kernel}
+\subsection{Sparse Matrix-Sparse Vector Multiplication Kernel}
 
@@
 
-\section{FlexGraph Parallelization}
+\section{FlexGraph Optimizations}
 
@@
 
-\section{FlexGraph Optimizations}
-
 This section describes some optimizations we added to FlexGraph to improve its performance.

### 2017-11-17 paper/mechanism.tex
backup

@@
 However, Graph Analytics datasets are generally very large and hyper-sparse which eliminate the DCSC format overhead.
+In addition to a sparse matrix, FlexGraph also uses a sparse vector to provide property data as well as edge selection during computation.
+A bitmask is used to encode the non-zero entries in the input vector.
 
-\subsection{Sparse Matrix-Sparse Vector Multiplication Kernel}
+\subsection{The Sparse Matrix-Sparse Vector Multiplication Kernel}
 
-FlexGraph architecture implements a custom Sparse Matrix Vertex Multiplication (SPMV) kernel.
-Its architecture diverges from conventional implementations in two ways: First, it uses the dirty masks from  the input vertices to select the edges to process and return a new dirty masks capturing the intersection of the active input vertices and the non-zero edges actually visited.
-Second, it matrix data structure (DCSC) contains an additional indirection buffer for accessing the matrix columns.
+FlexGraph micro-architecture implements a Sparse Matrix-Sparse Vertex Multiplication (SpMSpV) kernel in FPGA.
+This kernel diverges from conventional Sparse Matrix-Vector Multiplication (SpMV) implementations in two ways - firstly, it uses a sparse data structure as input and output vector during computation - secondly, it uses a doubly-compressed sparse matrix (DCSC) format with additional memory indirection buffer when accessing the matrix columns.
 These two properties present unique performance challenges when designing the accelerator.
-Listing 1 shows the pseudo-code of the SPMV kernel.
-The program iterates through each column and fetch the corresponding input vertex active mask to check if the column should be process, then access all the non-zero rows of the matrix for that column to evaluate each edge.
-In the code's comments on the right are highlighted the different types performance hogs present; semi-random memory accesses (*), fully-random memory accesses (!) and control branches (?).
-The memory accesses for the columns and rows data (lines 4 and 9) are semi-random because only the access cannot be predicted and after that the address is simply incremented.
+Listing 1 shows the pseudo-code of the SpMSpV kernel.
+Given the start/end addresses (col\_start, col\_end) into the matrix columns buffer (coldata), for The program iterates through each column entry and fetches the corresponding column index (col) and rows buffer address (row\_start, row\_end).
+It then checks if the current column index (col) is active using the vertex bitmask, only then it proceeds with rows traversal loop where the matrix row\_data is accessed to retrieve corresponding row index and matrix value for computation.
+A Multiply-Accumulate (MAC) operation is then performed on the matrix and vertex data and output bitmask is updated.
+Lines 6 and 8 show the code region where external memory is randomly accessed to obtain the vertex data and active state.
+Lines 5 and 10 show the code region where external memory is accessed semi-random because only the first access cannot be predicted and after that the address is simply incremented.
 FlexGraph's architecture attempts to address some of those performance hogs using several optimizations detailed in section 4.
 
-\begin{lstlisting}[language=Python, caption=Pseudo-code for SPMV kernel]
-def SPMV_kernel(x_values, x_activemask) :
-  y_values = {0}, y_activemask = {0}
-    for col in (c_start, c_end) :            	
-      (a_x, r_start, r_end) = coldata[col]   # *
-      x_value  = x_values[a_x]               # !
-      x_active = x_activemask[a_x]           # !
-      if (x_active):                         # ?
-        for row in (r_start, r_end):         
-          (a_y, a_value) = rowdata[row]      # *
-          y_values[a_y] += a_value * x_value # !
-          x_activemask[a_y] = true           # !
-  return (y_values, y_activemask)
+\begin{lstlisting}[
+	linebackgroundcolor={
+	\ifnum\value{lstnumber}=6\color{red!10}\fi
+	\ifnum\value{lstnumber}=8\color{red!10}\fi
+	\ifnum\value{lstnumber}=5\color{orange!10}\fi
+	\ifnum\value{lstnumber}=10\color{orange!10}\fi}, 
+        escapechar=\%, 
+        emph={col_start, col_end, coldata, rowdata, values, active}, emphstyle=\color{blue}, language=Python, caption=Pseudo-code for SpMSpV kernel]
+def SpMSpV_kernel(a, x):
+  y_values[] = {0}
+  y_bitmask[] = {false}
+  for i in (a.col_start, a.col_end):            	
+    (col, row_start, row_end) = a.coldata[i]
+    x_active = x.bitmask[col]          
+    if (x_active) :                            
+      x_value  = x.values[col]              
+      for j in (row_start, row_end): 
+        (row, a_value) = a.rowdata[j]
+        y_values[row] += a_value * x_value
+        y_bitmask[row] = true          
+  return (y_values, y_bitmask)
 \end{lstlisting}
@@
 Externally, FlexGraph input and output signals implement an Accelerator Functional Unit (AFU) interface defined by Intel's Accelerator Abstraction Layer (AAL) \cite{Intel-FPGA}.
-AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL software runnign on the host processor.
+AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL software running on the host processor.
 Listing 2 shows AAL device interface implementation using Cocoh's API.

### 2017-11-23 common.h
LKG version with pipelined MAC support

@@ -32,18 +32,20 @@ using namespace ch::sim;
 
 namespace spmv {
 
-using ch_block = ch_bit<aal::aal_qpi0::cdata_width>;
+using qpi = aal::aal_qpi0;
 
-using ch_blk_addr = ch_bit<aal::aal_qpi0::addr_wdith>;
+using ch_block = ch_bit<qpi::cdata_width>;
 
-using ch_ptr = ch_bit<aal::aal_qpi0::addr_wdith>;
+using ch_blk_addr = ch_bit<qpi::addr_wdith>;
+
+using ch_ptr = ch_bit<qpi::addr_wdith>;
 
 enum {
   PE_COUNT       = 2,
   LOG2_PE_COUNT  = log2ceil(PE_COUNT),
   RQ_OWNER_BITS  = log2ceil(PE_COUNT+1),
 
-  PTR_MAX_VALUE  = (1 << aal::aal_qpi0::addr_wdith)-1,
+  PTR_MAX_VALUE  = (1 << qpi::addr_wdith)-1,
 
   PARTITION_SIZE = 32,
   LOG2_PARTITION_SIZE = 5,
@@ -108,31 +110,31 @@ __struct (ch_ctrl_hwcntrs_t, (
   (ch_bit32) vcache_stalls
 ));
 
-__inout (ch_matrix_dcsc_t, (
-  __in(ch_bit32) numparts,
-  __in(ch_bit<aal::aal_qpi0::addr_wdith>) partitions,
-  __in(ch_bit<aal::aal_qpi0::addr_wdith>) xindices,
-  __in(ch_bit<aal::aal_qpi0::addr_wdith>) ystarts,
-  __in(ch_bit<aal::aal_qpi0::addr_wdith>) yindices,
-  __in(ch_bit<aal::aal_qpi0::addr_wdith>) values
+__struct (ch_matrix_dcsc_t, (
+  (ch_bit32) numparts,
+  (ch_bit<qpi::addr_wdith>) partitions,
+  (ch_bit<qpi::addr_wdith>) xindices,
+  (ch_bit<qpi::addr_wdith>) ystarts,
+  (ch_bit<qpi::addr_wdith>) yindices,
+  (ch_bit<qpi::addr_wdith>) values
 ));
 
-__inout (ch_vertex_t, (
-  __in(ch_bit<aal::aal_qpi0::addr_wdith>) values,
[…]

### 2017-11-23 cpu.cpp

@@ -4,9 +4,9 @@
 using namespace spmv;
 using namespace spmv::host;
 
-static int random_select(int vmin, int vmax, float skew) {
-  float alpha((rand() * 1.0f) / RAND_MAX);
-  return (int)std::ceil(vmin + (vmax - vmin) * alpha * skew);  
+static int random_select(int vmin, int vmax, float weight) {
+  auto alpha = rand() * weight / RAND_MAX;
+  return (int)std::round(vmin + (vmax - vmin) * alpha);
 }
 
 // poluate random values into vertex
@@ -26,36 +26,36 @@ static void init_vertex(vertex_t* vertex) {
 }
 
 cpu_device::cpu_device(const char* mtx_file) 
-  : m_shared_mem(nullptr) 
-  , m_matrix(nullptr)
-  , m_vertex(nullptr) {
+  : shared_mem_(nullptr)
+  , matrix_(nullptr)
+  , vertex_(nullptr) {
   // initialiaze matrix file
   this->init(mtx_file);
 
   // dump stats
-  ch_dumpStats(std::cout, m_accelerator);
+  ch_dumpStats(std::cout, accelerator_);
 
   // dump verilog
-  ch_toVerilog("spmv.v", m_accelerator);
+  ch_toVerilog("spmv.v", accelerator_);
 }
 
 cpu_device::~cpu_device() {
   // release allocated resources
-  if (m_matrix)
-    delete m_matrix;
-  if (m_vertex)
-    delete m_vertex;
-  if (m_shared_mem)
-    delete[] m_shared_mem;
-  if (!m_qpi_read_requests.empty()) {
-    std::cout << "CPU had " << m_qpi_read_requests.size() << " pending read requests" << std::endl;
-    for (auto rq : m_qpi_read_requests) {
+  if (matrix_)
+    delete matrix_;
+  if (vertex_)
+    delete vertex_;
+  if (shared_mem_)
+    delete[] shared_mem_;
+  if (!qpi_read_requests_.empty()) {
+    std::cout << "CPU had " << qpi_read_requests_.size() << " pending read requests" << std::endl;
+    for (auto rq : qpi_read_requests_) {
       delete rq;
     }
   }
-  if (!m_qpi_read_requests.empty()) {
-    std::cout << "CPU had " << m_qpi_write_requests.size() << " pending write requests" << std::endl;
-    for (auto rq : m_qpi_write_requests) {
[…]

### 2017-11-23 cpu.h

@@ -7,13 +7,60 @@
 namespace spmv {
 namespace host {
 
+template <typename T>
+class accumulator {
+public:
+
+  accumulator(): count_(0) {}
+
+  void update(const T& value) {
+    if (0 == count_) {
+      min_ = value;
+      max_ = value;
+      sum_ = value;
+    } else {
+      if (value < min_) {
+        min_ = value;
+      } else
+      if (value > max_) {
+        max_ = value;
+      }
+      sum_ += value;
+    }
+    ++count_;
+  }
+
+  const T& get_min() const {
+    return min_;
+  }
+
+  const T& get_max() const {
+    return max_;
+  }
+
+  const T& get_sum() const {
+    return sum_;
+  }
+
+  uint64_t get_count() const {
+    return count_;
+  }
+
+private:
+
+  T min_;
+  T max_;
+  T sum_;
+  uint64_t count_;
+};
+
 struct rq_data_t {
   rq_data_t() {
     for (int i = 0; i < sizeof(__m); ++i) {
       __m[i] = 0;
     }
   }
-  byte_t __m[__div_ceil(aal::aal_qpi0::cdata_width, 8)];
+  byte_t __m[__div_ceil(qpi::cdata_width, 8)];
 };
 
 struct rq_mdata_t {
@@ -22,7 +69,7 @@ struct rq_mdata_t {
       __m[i] = 0;
     }
   }
-  byte_t __m[__div_ceil(aal::aal_qpi0::mdata_width, 8)];
+  byte_t __m[__div_ceil(qpi::mdata_width, 8)];
 };
 
 struct request_t {
@@ -44,20 +91,25 @@ struct request_t {
 class cpu_device {
 public:  
   enum {
+    CLOCK_RATE          = 200000000, // 200MHz
+
     SHARE_MEM_SIZE      = 0x80000000, // 2GB
     
     PARTITION_SIZE      = 32,
     
     CACHE_HIT_RATE      = 80, // 80%
     
-    MIN_RX_RD_CYCLES    = 2,  // cache hit
-    MIN_TX_RD_CYCLES    = 2,  // cache hit
-    MAX_RX_RD_CYCLES    = 2,
-    MAX_TX_RD_CYCLES    = 32,    
-    MIN_RX_RW_CYCLES    = 2,  // cache hit
-    MIN_TX_RW_CYCLES    = 2,  // cache hit
[…]

### 2017-11-23 dcsc_walk.cpp

@@ -22,11 +22,14 @@ __enum (ch_walk_state, 4, (
   (end_partition, 0xF)
 ));
 
-spmv_dcsc_walk::spmv_dcsc_walk(uint32_t id)
-  : id_(id)
-  , xmblock_addr_(PTR_MAX_VALUE) // default initialize to max
+spmv_dcsc_walk::spmv_dcsc_walk()
+  : xmblock_addr_(PTR_MAX_VALUE) // default initialize to max
   , xvblock_addr_(PTR_MAX_VALUE) // default initialize to max{
 {
+  //--
+  static uint32_t s_ids = 0;
+  id_ = s_ids++;
+
   //--
   io.ctrl.hwcntrs = hwcntrs_;
 
@@ -70,11 +73,13 @@ void spmv_dcsc_walk::describe() {
   io.ctrl.start.ready = (state == ch_walk_state::get_partition);
 
   //--
-  if (id_ == 0) {
+  /*if (id_ == 0) {
     ch_print(fstring("{0}: *** Walker%d: state={1}, axbuf.enq.val={2}, axbuf.deq.val={3}, asbuf.enq.val={4}, asbuf.deq.val={5}, asbuf_psz={6}, #p={7}", id_),
              ch_getTick(), state, axbuf_.io.enq.valid, axbuf_.io.deq.valid,
              asbuf_.io.enq.valid, asbuf_.io.deq.valid, asbuf_pending_size_, hwcntrs_.num_partitions);
   }
+  */
+
   // ch_walk_state FSM
   __switch (state) (
   __case (ch_walk_state::get_partition) (
@@ -470,7 +475,7 @@ void spmv_dcsc_walk::describe() {
     io.pe.valid         = false;
   ));
 
-  //--
+  /*//--
   if (id_ == 0) {
     ch_print(fstring("{0}: Walker%d: state={1}, rq_val={2}, rq_typ={3}, rq_adr={4}, "
                      "rr_val={5}, rr_typ={6}, "
@@ -481,5 +486,5 @@ void spmv_dcsc_walk::describe() {
       col_curr_, col_end_, a_xindex_, x_value_, row_curr_, row_end_,
       io.pe.data.a_yindex, io.pe.data.a_value, io.pe.data.x_value, io.pe.data.is_end, io.pe.valid
     );
-  }
+  }*/
 }

### 2017-11-23 dcsc_walk.h

@@ -43,7 +43,7 @@ public:
     (ch_enq_io<ch_pe_req_t>) pe
   );
 
-  spmv_dcsc_walk(uint32_t id);
+  spmv_dcsc_walk();
   ~spmv_dcsc_walk();
 
   void describe();

### 2017-11-23 lsu.cpp

@@ -4,9 +4,9 @@ using namespace spmv;
 using namespace spmv::accelerator;
 
 __enum (ch_wr_req_state, 3, (
-  get_request,
-  submit,
-  writemask,
+  ready,
+  write_value,
+  write_mask,
   flush,
   wait_for_flush
 ));
@@ -46,15 +46,13 @@ void spmv_lsu::read_req_thread() {
   //--
   auto qpi_rd_req_valid = io.qpi.rd_req.valid.asSeq();
   auto qpi_rd_req_addr  = io.qpi.rd_req.addr.asSeq();
+  auto qpi_rd_req_mdata = io.qpi.rd_req.mdata.asSeq();
 
   //--
   qpi_rd_req_valid.next = rd_req_arb_.io.out.valid;
   qpi_rd_req_addr.next  = get_baseaddr(rd_req_arb_.io.out.data.type) + rd_req_arb_.io.out.data.addr;
-
-  //--
-  ch_seq<ch_rd_mdata_t> qpi_rd_req_mdata;
-  qpi_rd_req_mdata.next = ch_rd_mdata_t(rd_req_arb_.io.out.grant, rd_req_arb_.io.out.data.type);
-  io.qpi.rd_req.mdata = ch_zext<aal::aal_qpi0::mdata_width>(qpi_rd_req_mdata.asBits());
+  qpi_rd_req_mdata.next = ch_zext<qpi::mdata_width>(
+        ch_rd_mdata_t(rd_req_arb_.io.out.grant, rd_req_arb_.io.out.data.type).asBits());
     
   //--
   ch_print("{0}: lsu_rd_req: valid={1}, grant={2}, type={3}, addr={4}",
@@ -75,7 +73,7 @@ void spmv_lsu::write_req_thread() {
   auto qpi_write_ready = (qw_state == ch_qpi_write_state::ready);
 
   //--
-  wr_req_arb_.io.out.ready = (state == ch_wr_req_state::get_request);
+  wr_req_arb_.io.out.ready = (state == ch_wr_req_state::ready);
   wr_cache_.io.evict.ready = (qpi_write_ready & !lsu_write_valid);
 
   ch_seq<ch_lsu_wr_req_t> req_data;
@@ -100,18 +98,21 @@ void spmv_lsu::write_req_thread() {
         ch_wr_mdata_t mdata(req_owner, req_data.type);
         qpi_wr_req_addr.next  = get_baseaddr(req_data.type) + req_data.addr;
[…]

### 2017-11-23 lsu.h

@@ -90,7 +90,7 @@ public:
 
   __io (
     (spmv_aal_device::qpi_io) qpi,
-    (ch_ctx_io) ctx,
+    __in(spmv_ctx_t) ctx,
     (ch_ctrl_lsu_io) ctrl,
     (ch_vec<ch_walker_lsu_io, PE_COUNT>) walkers,
     (ch_vec<ch_pe_lsu_io, PE_COUNT>) PEs

### 2017-11-23 opae.h

@@ -2,7 +2,7 @@
 
 #include <cash.h>
 
-namespace aal {
+namespace opae {
 
 //
 // Description:
@@ -14,7 +14,7 @@ namespace aal {
 using namespace ch::core;
 
 template <unsigned ADDR, unsigned MDATA, unsigned CDATA>
-struct aal_qpi {
+struct opae_qpi {
   static constexpr unsigned addr_wdith  = ADDR;
   static constexpr unsigned mdata_width = MDATA;
   static constexpr unsigned cdata_width = CDATA;
@@ -54,27 +54,29 @@ struct aal_qpi {
   ));
 };
 
-using aal_qpi0 = aal_qpi<20, 14, 512>;
+using opae_qpi0 = opae_qpi<20, 14, 512>;
 
-using aal_qpi1 = aal_qpi<42, 16, 512>;
-
-template <typename Context, typename Qpi = aal_qpi0>
-class aal_device {
+template <typename Context, typename Qpi = opae_qpi0>
+class opae_device {
 public:
   using qpi_io = typename Qpi::io;
   static_assert(ch_bitwidth_v<Context> == Qpi::cdata_width, "invalid context size");
+  static_assert(ch_direction_v<Context> == ch_direction::in, "invalid context direction");
 
   __io (
     (qpi_io)          qpi,
-    __in(Context)     ctx,
+    (Context)         ctx,
     (ch_in<ch_bit1>)  start,
     (ch_out<ch_bit1>) done
   );
 
-  aal_device() {}
-  virtual ~aal_device() {}
-  
+  opae_device() {}
+  virtual ~opae_device() {}
+
   virtual void describe() = 0;
 };
+
+template <typename Context>
+using opae_device0 = opae_device<Context>;
 
 }

### 2017-11-23 pe.cpp

@@ -4,19 +4,21 @@
 using namespace spmv;
 using namespace spmv::accelerator;
 
-__enum (ch_pe_state, 3, (
+__enum (ch_pe_state, 2, (
   ready,
   write_value0,
   write_value1,
-  write_mask,
-  clear
+  write_mask
 ));
 
-spmv_pe::spmv_pe(uint32_t id) : id_(id) {
+spmv_pe::spmv_pe() {
+  //--
+  static uint32_t s_ids = 0;
+  id_ = s_ids++;
+
   //--
   __if (y_wenable_) (
-    y_values_[y_waddr_] = y_value_;
-    y_mask_.next = mask_value_;
+    y_values_[y_waddr_] = y_value_;    
   );
 }
 
@@ -27,56 +29,75 @@ void spmv_pe::describe() {
   ch_seq<ch_pe_state> state;
 
   //--
-  ch_bool do_enq = io.req.valid && io.req.ready;
+  auto y_raddr_mask = ch_pvBits(1) << y_raddr_;
+  auto y_waddr_mask = ch_pvBits(1) << y_waddr_;
 
   //--
-  auto mult_value = ch_fmult<ALTFP_SP_MULT>(io.req.data.a_value, io.req.data.x_value);
-  mult_pipe_.io.enq.data.a_yindex = io.req.data.a_yindex;
-  mult_pipe_.io.enq.data.is_end = io.req.data.is_end;
-  mult_pipe_.io.enq.valid = do_enq;
+  ch_bool add_enable = (ch_pe_state::ready == state)
+                     || !add_pipe_.io.deq.valid; // discard non-valid heads
 
   //--
-  auto y_raddr = (mult_pipe_.io.deq.data.a_yindex & 0x1f).slice<5>();
-  auto add_value = ch_fadd<ALTFP_SP_ADD_SUB>(mult_value, y_values_[y_raddr]);
-  add_pipe_.io.enq.data.a_yindex = mult_pipe_.io.deq.data.a_yindex;
-  add_pipe_.io.enq.data.is_end = mult_pipe_.io.deq.data.is_end;
-  add_pipe_.io.enq.valid = mult_pipe_.io.deq.valid;
+  ch_bool mult_enable = (add_enable && 0 == (inflight_mask_ & y_raddr_mask))
+                     || !mult_pipe_.io.deq.valid; // discard non-valid heads
 
   //--
[…]

### 2017-11-23 pe.h

@@ -27,14 +27,15 @@ public:
     __out(ch_bool) is_idle
   );
   
-  spmv_pe(uint32_t id);
+  spmv_pe();
   ~spmv_pe();
   
   void describe();
 
 private:
 
-  using ch_bit480 = ch_bit<480>;
+  using ch_pvBits = ch_bit<PARTITION_SIZE>;
+  using ch_paBits = ch_bit<LOG2_PARTITION_SIZE>;
   
   uint32_t  id_;
 
@@ -42,21 +43,19 @@ private:
   ch_module<ch_pipe<ch_pipe_data_t, ALTFP_SP_ADD_SUB>> add_pipe_;
 
   ch_ram<ch_float32, PARTITION_SIZE> y_values_;
-  ch_seq<ch_bit32> y_mask_;
+  ch_seq<ch_pvBits> y_mask_;
+  ch_seq<ch_pvBits> y_mask_cpy_;
 
-  ch_seq<ch_bit32> inflight_mask_;
+  ch_seq<ch_pvBits> inflight_mask_;
 
   ch_seq<ch_bit32> pending_reqs_;
 
-  ch_float32       y_value_;
-  ch_bit<5>        y_waddr_;
-  ch_bit1          y_wenable_;
+  ch_float32  y_value_;
+  ch_paBits   y_raddr_;
+  ch_paBits   y_waddr_;
+  ch_bit1     y_wenable_;
 
-  ch_bit32         mask_value_;
-
-  ch_seq<ch_ptr>   y0_;
-
-  ch_seq<ch_bit<LOG2_PARTITION_SIZE>> y_clr_cntr_;
+  ch_seq<ch_ptr>   y0_;  
 };
 
 }

### 2017-11-23 simulator.cpp

@@ -15,24 +15,24 @@ static double elapsed_time(uint64_t start, uint64_t end) {
 }
 
 simulator::simulator(const char* mtx_file)
-  : m_vcd_file("spmv.vcd") {
+  : vcd_file_("spmv.vcd") {
   // initialize random seed
   srand(0);
     
   //--
-  m_cpu = new host::cpu_device(mtx_file);  
-  m_tracer = new ch_vcdtracer(m_vcd_file, m_cpu->get_accelerator());
+  cpu_ = new host::cpu_device(mtx_file);
+  tracer_ = new ch_vcdtracer(vcd_file_, cpu_->get_accelerator());
 }
 
 simulator::~simulator() {   
   // release resources
-  delete m_tracer;
-  delete m_cpu;
+  delete tracer_;
+  delete cpu_;
 }
 
 void simulator::run(ch_tick run_ticks) {
   // reset native simulator
-  ch_tick t = m_tracer->reset(0);
+  ch_tick t = tracer_->reset(0);
 
   uint64_t start = __rdtsc();
 
@@ -40,11 +40,11 @@ void simulator::run(ch_tick run_ticks) {
   for (;t < run_ticks;) {
     
     // tick the cpu device
-    if (!m_cpu->tick(t))
+    if (!cpu_->tick(t))
       break;
     
     // advance native simulation
-    t = m_tracer->step(t);
+    t = tracer_->step(t);
   }
 
   uint64_t end = __rdtsc();
@@ -61,5 +61,5 @@ void simulator::trace(ch_tick t) {
 
 void simulator::dump_stats(ch_tick t) {
   //--
-  m_cpu->dump_stats(t);
+  cpu_->dump_stats(t);
 }

### 2017-11-23 simulator.h

@@ -18,9 +18,9 @@ protected:
   
   void dump_stats(ch_tick t);
   
-  host::cpu_device* m_cpu;
-  ch_vcdtracer* m_tracer;
-  std::ofstream m_vcd_file;
+  host::cpu_device* cpu_;
+  ch_vcdtracer* tracer_;
+  std::ofstream vcd_file_;
 };
 
 }

### 2017-11-23 spmv.cpp

@@ -13,14 +13,6 @@ __enum (ch_ctrl_state, 3, (
 ));
 
 spmv_device::spmv_device() {
-  // create the PEs
-  PEs_.reserve(PE_COUNT);
-  walkers_.reserve(PE_COUNT);
-  for (int i = 0; i < PE_COUNT; ++i) {
-    PEs_.emplace_back(i);
-    walkers_.emplace_back(i);
-  }
-
   // bind modules
   for (int i = 0; i < PE_COUNT; ++i) {
     PEs_[i].io.req(walkers_[i].io.pe);
@@ -209,6 +201,9 @@ void spmv_device::dispatch_thread() {
         part_buf_.next = part_buf_ >> PARTITION_VALUE_BITS; // pop one entry
         part_curr_.next = part_curr_ + 1; // advance partition
         part_buf_size_.next = part_buf_size_ - 1;
+
+        //ch_print(fstring("{0}: *** assigned partition{1} to PE%d", i), ch_getTick(), part_curr_);
+
         // check if we can pop another partition from current block
         // we need at least two entries to proceed
         __if (part_curr_.next != numparts_

### 2017-11-23 spmv.h

@@ -35,8 +35,8 @@ protected:
 
   ch_block get_hwnctrs(const ch_hwcntr_addr& addr);
 
-  std::vector<ch_module<spmv_dcsc_walk>> walkers_;
-  std::vector<ch_module<spmv_pe>> PEs_;
+  std::array<ch_module<spmv_dcsc_walk>, PE_COUNT> walkers_;
+  std::array<ch_module<spmv_pe>, PE_COUNT> PEs_;
   ch_module<spmv_lsu> lsu_;
   ch_module<ch_pbuf_t> pbuf_;
 

### 2017-11-23 wrcache.h

@@ -123,9 +123,11 @@ public:
             enq_data.next = io.enq.data;
             state.next = ch_state::lookup;
           );
-        )__elif (io.flush) (
-          counter_.next = 0;
-          state.next = ch_state::flush;
+        )__else (
+          __if (io.flush) (
+            counter_.next = 0;
+            state.next = ch_state::flush;
+          );
         );
       )
       __case (ch_state::lookup) (

### 2017-11-23 common.h
refactoring

@@ -90,32 +90,42 @@ inline bool test_bitvector(const uint32_t* bitvec, uint32_t idx) {
   return value & (1 << bit);
 }
 
-__struct (ch_pe_hwcntrs_t, (
-  (ch_bit32) num_partitions,
+__struct (ch_walker_stats_t, (
+  (ch_bit32) num_parts,
   (ch_bit32) min_latency,
   (ch_bit32) max_latency,
   (ch_bit32) total_latency,
-  (ch_bit32) execute_stalls,
-  (ch_bit32) a_xindices_stalls,
-  (ch_bit32) x_masks_stalls,
-  (ch_bit32) a_startys_stalls,
+  (ch_bit32) a_colind_stalls,
+  (ch_bit32) a_rowptr_stalls,
+  (ch_bit32) a_rowind_stalls,
   (ch_bit32) a_values_stalls,
+  (ch_bit32) x_values_stalls,
+  (ch_bit32) x_masks_stalls,
+  (ch_bit32) execute_stalls
+));
+
+__struct (ch_pe_stats_t, (
   (ch_bit32) write_value_stalls,
   (ch_bit32) write_mask_stalls
 ));
 
-__struct (ch_ctrl_hwcntrs_t, (
-  (ch_bit32) a_partitions_stalls,
+__struct (ch_stats_t, (
+  (ch_walker_stats_t) walker,
+  (ch_pe_stats_t) pe
+));
+
+__struct (ch_ctrl_stats_t, (
+  (ch_bit32) a_colptr_stalls,
   (ch_bit32) vcache_hits,
   (ch_bit32) vcache_stalls
 ));
 
 __struct (ch_matrix_dcsc_t, (
-  (ch_bit32) numparts,
-  (ch_bit<qpi::addr_wdith>) partitions,
-  (ch_bit<qpi::addr_wdith>) xindices,
-  (ch_bit<qpi::addr_wdith>) ystarts,
-  (ch_bit<qpi::addr_wdith>) yindices,
+  (ch_bit32) num_parts,
+  (ch_bit<qpi::addr_wdith>) col_ptr,
+  (ch_bit<qpi::addr_wdith>) col_ind,
+  (ch_bit<qpi::addr_wdith>) row_ptr,
+  (ch_bit<qpi::addr_wdith>) row_ind,
   (ch_bit<qpi::addr_wdith>) values
 ));
 
[…]

### 2017-11-23 cpu.cpp

@@ -77,15 +77,15 @@ void cpu_device::init(const char* mtx_file) {
   offset = matrix_->copy(shared_mem_, offset, SHARE_MEM_SIZE, accelerator_.io.ctx.a);
   offset = vertex_->copy(shared_mem_, offset, SHARE_MEM_SIZE, accelerator_.io.ctx.x);
   offset = vertex_->copy(nullptr, offset, SHARE_MEM_SIZE, accelerator_.io.ctx.y);
-  accelerator_.io.ctx.hwcntrs = (offset / BLOCK_SIZE);
+  accelerator_.io.ctx.stats = (offset / BLOCK_SIZE);
   offset += (1 + PE_COUNT) * BLOCK_SIZE;
   assert(offset <= SHARE_MEM_SIZE);
   
-  /*printf("*** partitions=0x%x, a_xindices=0x%x, a_startys=0x%x, a_yindices=0x%x, a_values=0x%x, x_values=x0%x, x_masks=0x%x\n",
-         (uint32_t)m_afu_ctx.a.partitions,
-         (uint32_t)m_afu_ctx.a.xindices,
-         (uint32_t)m_afu_ctx.a.ystarts,
-         (uint32_t)m_afu_ctx.a.yindices,
+  /*printf("*** a_colptr=0x%x, a_colind=0x%x, a_rowptr=0x%x, a_rowind=0x%x, a_values=0x%x, x_values=x0%x, x_masks=0x%x\n",
+         (uint32_t)m_afu_ctx.a.col_ptr,
+         (uint32_t)m_afu_ctx.a.col_ind,
+         (uint32_t)m_afu_ctx.a.row_ptr,
+         (uint32_t)m_afu_ctx.a.row_ind,
          (uint32_t)m_afu_ctx.a.values,
          (uint32_t)m_afu_ctx.x.values,
          (uint32_t)m_afu_ctx.x.masks);*/
@@ -95,15 +95,17 @@ void cpu_device::init(const char* mtx_file) {
   //this->check_output(nullptr, nullptr, true);
 }
 
-void cpu_device::check_output(const float* values, const uint32_t* masks, bool debug) {
+void cpu_device::check_output(const float* values,
+                              const uint32_t* masks,
+                              bool debug) {
   float y_values[32];
-  int   y_rows[32];
+  int   y_rowind[32];
   uint32_t y_mask;
   uint32_t y0;
   
   uint64_t errors = 0;
   
-  for (int p = 0, numparts = matrix_->numparts; p < numparts; ++p) {
[…]

### 2017-11-23 cpu.h

@@ -140,7 +140,9 @@ private:
   
   void check_result(ch_tick t);
   
-  void check_output(const float* values, const uint32_t* masks, bool debug);
+  void check_output(const float* values,
+                    const uint32_t* masks,
+                    bool debug);
   
   ch_device<accelerator::spmv_device> accelerator_;
 

### 2017-11-23 dcsc_walk.cpp

@@ -4,22 +4,22 @@ using namespace spmv;
 using namespace spmv::accelerator;
 
 __enum (ch_walk_state, 4, (
-  (get_partition, 0),
-  (get_a_xindex, 1),
-  (get_a_starty, 2),
-  (wait_for_a_starty, 3),
-  (get_a_endy, 4),
-  (wait_for_a_endy, 5),
-  (check_x_mask, 6),
-  (get_x_mask, 7),
-  (wait_for_x_mask, 8),
-  (get_x_value, 9),
-  (get_a_yindex, 0xA),
-  (get_a_value, 0xB),
-  (wait_for_data, 0xC),
-  (execute, 0xD),
-  (next_column, 0xE),
-  (end_partition, 0xF)
+  ready,
+  get_a_colidx,
+  get_a_rstart,
+  wait_for_a_rstart,
+  get_a_rend,
+  wait_for_a_rend,
+  check_x_mask,
+  get_x_mask,
+  wait_for_x_mask,
+  get_x_value,
+  get_a_rowidx,
+  get_a_value,
+  wait_for_data,
+  execute,
+  next_column,
+  end_partition
 ));
 
 spmv_dcsc_walk::spmv_dcsc_walk()
@@ -31,7 +31,7 @@ spmv_dcsc_walk::spmv_dcsc_walk()
   id_ = s_ids++;
 
   //--
-  io.ctrl.hwcntrs = hwcntrs_;
+  io.ctrl.stats = stats_;
 
   //--
   axbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
@@ -42,9 +42,9 @@ spmv_dcsc_walk::spmv_dcsc_walk()
   xmbuf_.io.enq.data = io.lsu.rd_rsp.data.data;
 
   //--
-  axbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_xindices);
-  asbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_startys);
-  aybuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_yindices);
+  axbuf_.io.enq.valid = io.lsu.rd_rsp.valid && (io.lsu.rd_rsp.data.type == ch_rd_request::a_colind);
[…]

### 2017-11-23 dcsc_walk.h

@@ -31,7 +31,7 @@ __struct (ch_pe_start_req_t, (
 __inout (ch_ctrl_pe_io, (
   (ch_deq_io<ch_pe_start_req_t>) start,
   __in(ch_bit64) timer,
-  __out(ch_pe_hwcntrs_t) hwcntrs
+  __out(ch_walker_stats_t) stats
 ));
 
 class spmv_dcsc_walk {
@@ -55,7 +55,7 @@ private:
   ch_seq<ch_ptr> col_curr_;
   ch_seq<ch_ptr> col_end_;
 
-  ch_seq<ch_ptr> a_xindex_;
+  ch_seq<ch_ptr> a_colidx_;
 
   ch_seq<ch_ptr> row_curr_;
   ch_seq<ch_ptr> row_end_;
@@ -91,7 +91,7 @@ private:
   ch_seq<ch_block> xvblock_;
 
   ch_seq<ch_bit64> prof_start_;
-  ch_seq<ch_pe_hwcntrs_t> hwcntrs_;
+  ch_seq<ch_walker_stats_t> stats_;
 };
 
 }

### 2017-11-23 graph.cpp

@@ -10,47 +10,50 @@ struct __attribute__((aligned(16))) edge_t {
 };
 
 // sort by partition (pid), then cols (dst), then rows (src)
-static bool compare_edges_dcsc(const edge_t& a, const edge_t& b) {
-  if (a.pid < b.pid)
+static bool compare_edges_dcsc(const edge_t& lhs, const edge_t& rhs) {
+  if (lhs.pid < rhs.pid)
     return true;
-  else if (a.pid > b.pid)
+  else if (lhs.pid > rhs.pid)
     return false;
-  if (a.dst < b.dst)
+  if (lhs.dst < rhs.dst)
     return true;
-  else if (a.dst > b.dst)
+  else if (lhs.dst > rhs.dst)
     return false;
-  if (a.src < b.src)
+  if (lhs.src < rhs.src)
     return true;
-  else if (a.src > b.src)
+  else if (lhs.src > rhs.src)
     return false;
   return false;
 }
 
-static uint32_t static_partition(uint32_t *partitions, uint32_t num_elements, uint32_t num_partitions, uint32_t alignment) {
+static uint32_t generate_partition(uint32_t *partitions,
+                                   uint32_t num_elements,
+                                   uint32_t num_parts,
+                                   uint32_t alignment) {
   partitions[0] = 0;
   uint32_t p = 1;
   uint32_t nnz = 1;
   uint32_t aligned_size = __div_ceil(num_elements, alignment) * alignment;
-  uint32_t partition_size = __div_ceil(aligned_size, num_partitions);
-  partition_size = __div_ceil(partition_size, alignment) * alignment;
-  for (uint32_t n = partition_size; n < aligned_size; n += partition_size) {
+  uint32_t part_size = __div_ceil(aligned_size, num_parts);
+  part_size = __div_ceil(part_size, alignment) * alignment;
+  for (uint32_t n = part_size; n < aligned_size; n += part_size) {
     partitions[p++] = n;
     ++nnz;
   }
   do {
     partitions[p++] = aligned_size;
-  } while (p <= num_partitions);
+  } while (p <= num_parts);
   return nnz;
 }
 
-static void set_edge_pointers(uint32_t *edge_pointers,
-                              const uint32_t *row_pointers,
-                              uint32_t num_partitions,
[…]

### 2017-11-23 graph.h

@@ -11,12 +11,12 @@ public:
   uint32_t  rows;
   uint32_t  nnz;
   uint32_t  nzx;
-  uint32_t  numparts;
-  uint32_t  element_size;
-  uint32_t* partitions;
-  uint32_t* xindices;
-  uint32_t* ystarts;
-  uint32_t* yindices;  
+  uint32_t  num_parts;
+  uint32_t  data_size;
+  uint32_t* col_ptr;
+  uint32_t* col_ind;
+  uint32_t* row_ptr;
+  uint32_t* row_ind;
   byte_t*   values;    
   
   mdcsc_t();
@@ -25,12 +25,12 @@ public:
     uint32_t rows,
     uint32_t nnz,
     uint32_t nzx,
-    uint32_t numparts,
-    uint32_t element_size) {
-    this->init(cols, rows, nnz, nnz, numparts, element_size);  
+    uint32_t num_parts,
+    uint32_t data_size) {
+    this->init(cols, rows, nnz, nnz, num_parts, data_size);
   }
   
-  mdcsc_t(const char* mtx_file, mtx_format edge_type, uint32_t partition_size);
+  mdcsc_t(const char* mtx_file, mtx_format edge_type, uint32_t part_size);
   ~mdcsc_t();
   
   
@@ -38,25 +38,31 @@ public:
             uint32_t rows,
             uint32_t nnz,
             uint32_t nzx,
-            uint32_t numparts,
-            uint32_t element_size);
+            uint32_t num_parts,
+            uint32_t data_size);
   
   void allocate();
   
-  size_t copy(byte_t* dest, size_t offset, size_t size, ch_scalar_t<ch_matrix_dcsc_t>& desc);
+  size_t copy(byte_t* dest,
+              size_t offset,
+              size_t size,
+              ch_scalar_t<ch_matrix_dcsc_t>& desc);
 };
 
 struct vertex_t {
   byte_t*   values;
   uint32_t* masks;
   uint32_t  size;
-  uint32_t  element_size;
+  uint32_t  data_size;
   
   vertex_t();
-  vertex_t(uint32_t size, uint32_t element_size);
+  vertex_t(uint32_t size, uint32_t data_size);
   ~vertex_t();
   
-  size_t copy(byte_t* dest, size_t offset, size_t size, ch_scalar_t<ch_vertex_t>& desc);
+  size_t copy(byte_t* dest,
+              size_t offset,
+              size_t size,
+              ch_scalar_t<ch_vertex_t>& desc);
   
   void dump();
 };

### 2017-11-23 lsu.cpp

@@ -101,8 +101,7 @@ void spmv_lsu::write_req_thread() {
         qpi_wr_req_mdata.next = ch_zext<qpi::mdata_width>(mdata.asBits());
         // go to write
         qw_state.next = ch_qpi_write_state::write;
-      )
-      __else (
+      )__else (
         __if (wr_cache_.io.evict.valid) (
           // get the data
           auto owner = ch_zext<PE_COUNT+1>(wr_cache_.io.evict.data.owner); // add ctrl bit
@@ -198,8 +197,8 @@ void spmv_lsu::write_req_thread() {
       wr_cache_.io.enq.data.owner = 0;
       wr_cache_.io.enq.data.tag   = 0;
       wr_cache_.io.enq.data.data  = 0;
-      wr_cache_.io.enq.valid = false;
-      wr_cache_.io.flush = false;
+      wr_cache_.io.enq.valid      = false;
+      wr_cache_.io.flush          = false;
       //--
       lsu_write_valid = false;
     ));
@@ -248,13 +247,13 @@ void spmv_lsu::write_rsp1_thread() {
 ch_blk_addr spmv_lsu::get_baseaddr(const ch_rd_request& rq_type) {
   ch_blk_addr addr;
   __switch (rq_type) (
-  __case (ch_rd_request::a_partition) (addr = io.ctx.a.partitions;)
-  __case (ch_rd_request::a_xindices)  (addr = io.ctx.a.xindices;)
-  __case (ch_rd_request::a_startys)   (addr = io.ctx.a.ystarts;)
-  __case (ch_rd_request::a_yindices)  (addr = io.ctx.a.yindices;)
-  __case (ch_rd_request::a_values)    (addr = io.ctx.a.values;)
-  __case (ch_rd_request::x_values)    (addr = io.ctx.x.values;)
-  __default                           (addr = io.ctx.x.masks;)
+  __case (ch_rd_request::a_colptr) (addr = io.ctx.a.col_ptr;)
+  __case (ch_rd_request::a_colind) (addr = io.ctx.a.col_ind;)
+  __case (ch_rd_request::a_rowptr) (addr = io.ctx.a.row_ptr;)
+  __case (ch_rd_request::a_rowind) (addr = io.ctx.a.row_ind;)
+  __case (ch_rd_request::a_values) (addr = io.ctx.a.values;)
+  __case (ch_rd_request::x_values) (addr = io.ctx.x.values;)
[…]

### 2017-11-23 lsu.h

@@ -21,10 +21,10 @@ enum {
 };
 
 __enum (ch_rd_request, 3, (
-  a_partition,
-  a_xindices,
-  a_startys,
-  a_yindices,
+  a_colptr,
+  a_colind,
+  a_rowptr,
+  a_rowind,
   a_values,
   x_values,
   x_masks
@@ -33,7 +33,7 @@ __enum (ch_rd_request, 3, (
 __enum (ch_wr_request, 2, (
   y_values,
   y_masks,
-  hwcntrs
+  stats
 ));
 
 __struct(ch_rd_mdata_t, (
@@ -67,14 +67,14 @@ __struct (ch_lsu_wr_rsp_t, (
 ));
 
 __inout (ch_ctrl_lsu_io, (
-  (ch_deq_io<ch_lsu_rd_req_t>) rd_req,
-  (ch_deq_io<ch_lsu_wr_req_t>) wr_req,
+  (ch_deq_io<ch_lsu_rd_req_t>)   rd_req,
+  (ch_deq_io<ch_lsu_wr_req_t>)   wr_req,
   (ch_valid_io<ch_lsu_rd_rsp_t>) rd_rsp,
-  __out(ch_bit32) outstanding_writes
+  __out(ch_bit32)                outstanding_writes
 ));
 
 __inout (ch_walker_lsu_io, (
-  (ch_deq_io<ch_lsu_rd_req_t>) rd_req,
+  (ch_deq_io<ch_lsu_rd_req_t>)   rd_req,
   (ch_valid_io<ch_lsu_rd_rsp_t>) rd_rsp
 ));
 

### 2017-11-23 pe.cpp

@@ -17,6 +17,9 @@ spmv_pe::spmv_pe() {
   id_ = s_ids++;
 
   //--
+  io.stats = stats_;
+
+  //--
   __if (y_wenable_) (
     y_values_[y_waddr_] = y_value_;    
   );
@@ -52,7 +55,7 @@ void spmv_pe::describe() {
       ch_select(0 != (y_mask_ & y_raddr_mask), y_values_[y_raddr_], 0);
 
   // Multiply pipeline
-  mult_pipe_.io.enq.data.a_yindex = io.req.data.a_yindex;
+  mult_pipe_.io.enq.data.a_rowind = io.req.data.a_rowind;
   mult_pipe_.io.enq.data.is_end = io.req.data.is_end;
   mult_pipe_.io.enq.valid = io.req.valid && io.req.ready;
   mult_pipe_.io.deq.ready = mult_enable;
@@ -60,11 +63,11 @@ void spmv_pe::describe() {
         io.req.data.a_value, io.req.data.x_value, mult_enable);
 
   // Adder pipeline
-  add_pipe_.io.enq.data.a_yindex = mult_pipe_.io.deq.data.a_yindex;
+  add_pipe_.io.enq.data.a_rowind = mult_pipe_.io.deq.data.a_rowind;
   add_pipe_.io.enq.data.is_end = mult_pipe_.io.deq.data.is_end;
   add_pipe_.io.enq.valid = mult_pipe_.io.deq.valid && mult_enable;
   add_pipe_.io.deq.ready = add_enable;
-  y_raddr_ = (mult_pipe_.io.deq.data.a_yindex & 0x1f).slice<5>();  
+  y_raddr_ = (mult_pipe_.io.deq.data.a_rowind & 0x1f).slice<5>();  
   ch_float32 add_value = ch_fadd<ALTFP_SP_ADD_SUB>(
         mult_value, prev_y_value, add_enable);
 
@@ -73,8 +76,7 @@ void spmv_pe::describe() {
   auto pe_commit = (ch_pe_state::ready == state) && add_pipe_.io.deq.valid;
   __if (pe_issue && !pe_commit) (
     pending_reqs_.next = pending_reqs_ + 1;
-  )
-  __else (
+  )__else (
     __if (!pe_issue && pe_commit) (
       pending_reqs_.next = pending_reqs_ - 1;
     );
@@ -98,7 +100,7 @@ void spmv_pe::describe() {
 
   //--
[…]

### 2017-11-23 pe.h

@@ -7,14 +7,14 @@ namespace spmv {
 namespace accelerator {
 
 __struct (ch_pe_req_t, (
-  (ch_ptr)     a_yindex,
+  (ch_ptr)     a_rowind,
   (ch_float32) a_value,
   (ch_float32) x_value,
   (ch_bool)    is_end
 ));
 
 __struct (ch_pipe_data_t, (
-  (ch_ptr)  a_yindex,
+  (ch_ptr)  a_rowind,
   (ch_bool) is_end
 ));
 
@@ -24,6 +24,7 @@ public:
   __io(
     (ch_deq_io<ch_pe_req_t>) req,
     (ch_flip_t<ch_pe_lsu_io>) lsu,
+    __out(ch_pe_stats_t) stats,
     __out(ch_bool) is_idle
   );
   
@@ -56,6 +57,8 @@ private:
   ch_bit1     y_wenable_;
 
   ch_seq<ch_ptr>   y0_;  
+
+  ch_seq<ch_pe_stats_t> stats_;
 };
 
 }

### 2017-11-23 spmv.cpp

@@ -8,7 +8,7 @@ __enum (ch_ctrl_state, 3, (
   get_partition,
   wait_for_execute,
   flush_buffers,
-  write_hwcntrs,
+  write_stats,
   wait_for_writes
 ));
 
@@ -29,7 +29,7 @@ spmv_device::~spmv_device() {
 
 void spmv_device::describe() {  
   //--
-  numparts_ = io.ctx.a.numparts.slice<ch_bitwidth_v<ch_ptr>>();
+  num_parts_ = io.ctx.a.num_parts.slice<ch_bitwidth_v<ch_ptr>>();
 
   // timer
   ch_seq<ch_bit64> timer;
@@ -52,39 +52,43 @@ void spmv_device::main_thread() {
   auto done = io.done.asSeq();
 
   //--
-  pbuf_.io.enq.data = lsu_.io.ctrl.rd_rsp.data.data;
-  pbuf_.io.enq.valid = lsu_.io.ctrl.rd_rsp.valid && (lsu_.io.ctrl.rd_rsp.data.type == ch_rd_request::a_partition);
+  pbuf_.io.enq.data  = lsu_.io.ctrl.rd_rsp.data.data;
+  pbuf_.io.enq.valid = lsu_.io.ctrl.rd_rsp.valid
+                   && (lsu_.io.ctrl.rd_rsp.data.type == ch_rd_request::a_colptr);
 
   //--
   pbuf_pending_size_.next = pbuf_pending_size_ +
-      ch_zext<32>(lsu_.io.ctrl.rd_req.valid && lsu_.io.ctrl.rd_req.ready && (lsu_.io.ctrl.rd_req.data.type == ch_rd_request::a_partition)) -
+      ch_zext<32>(lsu_.io.ctrl.rd_req.valid
+               && lsu_.io.ctrl.rd_req.ready
+               && (lsu_.io.ctrl.rd_req.data.type == ch_rd_request::a_colptr)) -
       ch_zext<32>(pbuf_.io.deq.ready);
 
   //--
   ch_bool all_PEs_ready = walkers_[0].io.ctrl.start.ready && PEs_[0].io.is_idle;
   for (int i = 1; i < PE_COUNT; ++i) {
-    all_PEs_ready = ch_clone(all_PEs_ready) && walkers_[i].io.ctrl.start.ready && PEs_[i].io.is_idle;
+    all_PEs_ready = ch_clone(all_PEs_ready)
+                 && walkers_[i].io.ctrl.start.ready
[…]

### 2017-11-23 spmv.h

@@ -19,7 +19,7 @@ using ch_part_buf = ch_bit<ch_bitwidth_v<ch_block> + PARTITION_VALUE_BITS>;
 
 using ch_pbuf_t = ch_queue<ch_block, PBUF_SIZE>;
 
-using ch_hwcntr_addr = ch_bit<LOG2_HWCNTR_ADDR_SIZE>;
+using ch_stats_addr = ch_bit<LOG2_HWCNTR_ADDR_SIZE>;
 
 class spmv_device : public spmv_aal_device {
 public:
@@ -33,7 +33,7 @@ protected:
   void main_thread();
   void dispatch_thread();
 
-  ch_block get_hwnctrs(const ch_hwcntr_addr& addr);
+  ch_block get_stats(const ch_stats_addr& addr);
 
   std::array<ch_module<spmv_dcsc_walk>, PE_COUNT> walkers_;
   std::array<ch_module<spmv_pe>, PE_COUNT> PEs_;
@@ -45,15 +45,15 @@ protected:
 
   ch_seq<ch_bit32> pbuf_pending_size_;
 
-  ch_seq<ch_bit32> a_partitions_stalls_;
-
   ch_seq<ch_ptr> part_blk_curr_;
   ch_seq<ch_ptr> part_blk_end_;
   ch_seq<ch_ptr> part_curr_;
 
-  ch_seq<ch_hwcntr_addr> hwcntrs_addr_;
+  ch_seq<ch_stats_addr> stats_addr_;
 
-  ch_ptr numparts_;
+  ch_seq<ch_ctrl_stats_t> stats_;
+
+  ch_ptr num_parts_;
 };
 
 }

### 2017-11-29 pe.cpp
backup

@@ -91,10 +91,9 @@ void spmv_pe::describe() {
     )__else (
       inflight_mask_.next = inflight_mask_ & ~y_waddr_mask;
     );
-  )__else (
-    __if (adder_issue) (
-      inflight_mask_.next = inflight_mask_ | y_raddr_mask;
-    );
+  )
+  __elif (adder_issue) (
+    inflight_mask_.next = inflight_mask_ | y_raddr_mask;
   );
 
   //--

### 2017-11-30 dcsc_walk.cpp
backup

@@ -373,12 +373,7 @@ void spmv_dcsc_walk::describe() {
         // get x_value from local storage
         x_value_.next = (xvblock_ >> INT32_TO_BLOCK_BITSHIFT(a_colidx_)).slice<32>().as<ch_float32>();
       );
-      // fetch first (a_rowind, a_value) blocks
-      arblock_.next = arbuf_.io.deq.data;
-      avblock_.next = avbuf_.io.deq.data;
-      arbuf_.io.deq.ready = true;
-      avbuf_.io.deq.ready = true;
-      // proceed to execution
+      // go to execute
       state.next = ch_walk_state::execute;
     )__else (
       // profiling
@@ -387,8 +382,8 @@ void spmv_dcsc_walk::describe() {
   )
   __case (ch_walk_state::execute) (
     // push data to PE
-    io.pe.data.a_rowind = (arblock_ >> INT32_TO_BLOCK_BITSHIFT(row_curr_)).slice<ch_bitwidth_v<ch_ptr>>();
-    io.pe.data.a_value  = (avblock_ >> INT32_TO_BLOCK_BITSHIFT(row_curr_)).slice<32>().as<ch_float32>();
+    io.pe.data.a_rowind = (arbuf_.io.deq.data >> INT32_TO_BLOCK_BITSHIFT(row_curr_)).slice<ch_bitwidth_v<ch_ptr>>();
+    io.pe.data.a_value  = (avbuf_.io.deq.data >> INT32_TO_BLOCK_BITSHIFT(row_curr_)).slice<32>().as<ch_float32>();
     io.pe.data.x_value  = x_value_.as<ch_float32>();
     io.pe.data.is_end   = false;
     io.pe.valid         = true;
@@ -401,13 +396,14 @@ void spmv_dcsc_walk::describe() {
       __if (row_curr_.next != row_end_) (
         // check if last entry in block
         __if ((row_curr_ & 0xf) == 0xf) (
-          // fetch the next (a_rowind, a_value) blocks
-          arblock_.next = arbuf_.io.deq.data;
-          avblock_.next = avbuf_.io.deq.data;
+          // pop fifo
           arbuf_.io.deq.ready = true;
           avbuf_.io.deq.ready = true;
         );
       )__else (
+        // pop fifo
+        arbuf_.io.deq.ready = true;
+        avbuf_.io.deq.ready = true;
         // go to next column
         state.next = ch_walk_state::next_column;
       );

### 2017-11-30 dcsc_walk.h

@@ -85,8 +85,6 @@ private:
 
   ch_seq<ch_block> acblock_;
   ch_seq<ch_block> asblock_;
-  ch_seq<ch_block> arblock_;
-  ch_seq<ch_block> avblock_;
   ch_seq<ch_block> xmblock_;
   ch_seq<ch_block> xvblock_;
 

### 2017-12-18 paper/introduction.tex
backup

@@
 The remainder of this paper is organized as follows:
-Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 details the various architecture optimizations we used in FlexGraph, section 5 describes FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 6 describes the experimental setup, section 7 describes our results analysis, section 8 describes the related work, section 9 summarizes our main contribution and results.
+Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 describes FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 5 describes the experimental setup, section 6 describes our results analysis, section 7 describes the related work, section 8 summarizes our main contribution and results.

### 2017-12-18 paper/mechanism.tex
backup

@@
 
-\begin{figure}[ htbp ] 
+\begin{figure}[htbp] 
 \begin{equation}
@@
 
-\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, caption=Graph Vertex Processing Model]
+\begin{lstlisting}[
+	emph={Assign, Process, Apply}, emphstyle=\color{blue}, 
+	numbers=none, 
+	escapechar=\%, 
+	caption=Graph Vertex Processing Model,
+	label={lst:vertex_model}]
 GraphProgram(%\textit{G}%, %\textit{P}%, %\textit{N}%) :
@@
 
-Listing 1 shows the three stages of a generalized graph vertex processing model - \textit{Assign}, \textit{Process} and \textit{Apply}, given a graph \textit{G}, some vertex properties \textit{P} and an iteration count \textit{N}.
+Listing \ref{lst:vertex_model} shows the three stages of a generalized graph vertex processing model - \textit{Assign}, \textit{Process} and \textit{Apply}, given a graph \textit{G}, some vertex properties \textit{P} and an iteration count \textit{N}.
 The \textit{Assign} stage generates an input vector \textit{x} using the vertex properties \textit{P}.
@@
 
-\begin{figure*}
-\centering
-\begin{subfigure}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{figures/offload_model}
-    \caption{Offload model}
-    \label{fig:offload_model}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-    \includegraphics[width=\textwidth]{figures/collaborative_model}
-    \caption{Collaborative model}
-    \label{fig:collaborative_model}
-\end{subfigure}
-\caption{Shared Memory Compute Models}
-\label{fig:compute_model}
-\end{figure*}
-
 With the rising adoption of FPGAs in production data centers \cite{Catapult}, there is a need to increase its ecosystem by making them more energy efficient as well as accessible to both the users and designers.
@@
 \subsection{Collaborative CPU-FPGA Computation}
+
+\begin{figure}[htbp]
+\centering
+\begin{subfigure}{0.25\textwidth}
+    \includegraphics[width=\textwidth]{figures/offload_model}
+    \caption{Offload model}
+    \label{fig:offload_model}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.21\textwidth}
+    \includegraphics[width=\textwidth]{figures/collaborative_model}
+    \caption{Collaborative model}
+    \label{fig:collaborative_model}
+\end{subfigure}
+\caption{Shared Memory Compute Models}
+\label{fig:compute_model}
+\end{figure}
 
@@
 \begin{figure}[htbp]
-\centering
-\includegraphics[width=0.5\textwidth]{figures/DCSC_matrix_format}
-\caption{DCSC hyperspace matrix format}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=\textwidth]{figures/csc_format}
+    \vspace{0pt}
+\end{subfigure}
+\begin{subfigure}{0.33\textwidth}
+    \includegraphics[width=\textwidth]{figures/dcsc_format}
+\end{subfigure}
+\caption{Matrix formats comparison}
 \label{fig:DCSC_matrix_format}
@@
 The format allows minimal traversal into the sparse matrix structure, saving necessary memory bandwidth when fetching empty columns.
-Figure \ref{fig:DCSC_matrix_format} shows a sample matrix A = \{(0,5,a), (0,7,b), (6,3,c), (7,1,d)\} with four non-zero edges (src, dst, weight) encoded using DCSC versus the conventional CSC \cite{CSC} format.
+Figure \ref{fig:DCSC_matrix_format} shows the memory layout for a sample matrix A = \{(0,5,a), (0,7,b), (6,3,c), (7,1,d)\} with four non-zero edges (src, dst, weight) encoded using DCSC versus the conventional CSC \cite{CSC} format.
 The traversal over CSC requires accessing all consecutive column ranges \textit{(cols[i+2] - cols[i])} even though most of the distances are empty.
@@
 
+\begin{lstlisting}[
+	linebackgroundcolor={
+	\ifnum\value{lstnumber}=6\color{red!10}\fi
+	\ifnum\value{lstnumber}=8\color{red!10}\fi
+	\ifnum\value{lstnumber}=5\color{orange!10}\fi
+	\ifnum\value{lstnumber}=10\color{orange!10}\fi}, 
+    escapechar=\%, 
+    emph={col_ind, col_ptr, row_ind, row_ptr, a_val}, emphstyle=\color{blue}, 
+    caption=Pseudo-code for SpMSpV kernel,
+    label={lst:spmv_code}]
+def SpMSpV_kernel(a, x):
+  y_values[] = {0}
+  y_bitmask[] = {false}
+  for_each i in a.col_ptr
+    (col_ind, row_ptr) := a.col_data[i]
+    x_active = x.bitmask[col_ind]          
+    if (x_active):                            
+      x_val = x.values[col_ind]              
+      for_each j in row_ptr
+        (row_ind, a_val) := a.row_data[j]
+        y_values[row_ind] += a_val * x_val
+        y_bitmask[row_ind] = true          
+  return (y_values, y_bitmask)
+\end{lstlisting}
+
 FlexGraph micro-architecture implements a Sparse Matrix-Sparse Vertex Multiplication (SpMSpV) kernel in FPGA.
@@
 These two properties present unique performance challenges when designing the accelerator.
-Listing 1 shows the pseudo-code of the SpMSpV kernel.
+Listing \ref{lst:spmv_code} shows the pseudo-code of the SpMSpV kernel.
 Given the start/end addresses (col\_start, col\_end) into the matrix columns buffer (coldata), for The program iterates through each column entry and fetches the corresponding column index (col) and rows buffer address (row\_start, row\_end).
@@
 
-\begin{lstlisting}[
-	linebackgroundcolor={
-	\ifnum\value{lstnumber}=6\color{red!10}\fi
-	\ifnum\value{lstnumber}=8\color{red!10}\fi
-	\ifnum\value{lstnumber}=5\color{orange!10}\fi
-	\ifnum\value{lstnumber}=10\color{orange!10}\fi}, 
-        escapechar=\%, 
-        emph={col_start, col_end, coldata, rowdata, values, active}, emphstyle=\color{blue}, language=Python, caption=Pseudo-code for SpMSpV kernel]
-def SpMSpV_kernel(a, x):
-  y_values[] = {0}
-  y_bitmask[] = {false}
-  for i in (a.col_start, a.col_end):            	
-    (col, row_start, row_end) = a.coldata[i]
-    x_active = x.bitmask[col]          
-    if (x_active) :                            
-      x_value  = x.values[col]              
-      for j in (row_start, row_end): 
-        (row, a_value) = a.rowdata[j]
-        y_values[row] += a_value * x_value
-        y_bitmask[row] = true          
-  return (y_values, y_bitmask)
-\end{lstlisting}
-
 \subsection{FlexGraph Microarchitecture}
 
-Externally, FlexGraph input and output signals implement an Accelerator Functional Unit (AFU) interface defined by Intel's Accelerator Abstraction Layer (AAL) \cite{Intel-FPGA}.
-AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL software running on the host processor.
-Listing 2 shows AAL device interface implementation using Cocoh's API.
-Our FlexGraph Accelerator's class in Cocoh simply derives from this interface and override the \textit{initialize()} function to provide its implementation and Cocoh takes care of generating the corresponding verilog module.
-It implements three input signals \textit{(start, qpi\_in, ctx)} and two output signals \textit{(qpi\_out, done)}.
-The AFU starts execution when the \textit{start} signal is asserted and communicate completion by asserting the \textit{done} signal.
-The \textit{ctx} signal provides application's specific data like constants and buffers address in the case of FlexGraph.
-The \textit{qpi} in/out signals implement Intel Quick Path Interconnect (QPI) interface \cite{QPI}, providing single channel read/write ports for accessing external shared memory.
-The FPGA socket hosts a 64-byte cache coherent interface (CCI) \cite{CCI} that connects to the host processor's last level cache (LLC) via QPI.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/microarchitecture}
+\caption{FlexGraph Microarchitecture}
+\label{fig:microarchitecture}
+\end{figure}
+
+Figure \ref{fig:microarchitecture} illustrates the microarchitecture of FlexGraph accelerator.
+Externally, FlexGraph input and output ports implement an Accelerator Functional Unit (AFU) interface defined by Intel's Accelerator Abstraction Layer (AAL) \cite{Intel-FPGA}.
+AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL runtime on the host processor.
+The interface exposes four signals \textit{qpi, ctx, start, done}, where the \textit{qpi} bus handles shared-memory communications, \textit{ctx} is used for passing custom input context to the accelerator, in our case we pass down all data structure layout information needed to access memory, \textit{start} and \textit{done} are the control knobs for starting and ending execution.
+At a high-level, Flexgraph's
+microarchitecture consists for a six building blocks: The main controller, the DCSC matrix walkers, the processing elements, the stream writers, the vertex caches and the load/store unit (LSU).
+
+\subsection{Main Controller}
+
+Upon reception of the \textit{start} signal from the host processor, the main controller is responsible for scheduling matrix partitions for execution on each processing element.
+It is comprised of two execution threads: A request thread that fetches partition data from the load/store unit into a local buffer, a dispatch thread that submit received partition data to individual matrix walker engines for processing.
+After all matrix partitions have been processed, the controller sends performance counters back to the host processor and activates the \textit{done} signal.
+
+\subsection{DCSC Matrix Walkers}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/micro_walker}
+\caption{DCSC Walker}
+\label{fig:micro_walker}
+\end{figure}
+
+The DCSC Matrix Walker is a custom data structure traversal engine for the DCSC Matrix.
+Figure \ref{fig:micro_walker} shows an overview of the engine microarchitecture.
+It is implemented as a four-stage state machine where each stage accesses a portion of the matrix layout as reflected in the pseudo-code in Listing \ref{lst:spmv_code}.
+In the first stage, the engine fetches column data (\textit{col\_ind, row\_ptr}) from memory given the partition address.
+In the second stage, the engine fetches the vertex active mask from memory given the \textit{col\_ind} value obtained in the previous stage.
+In the third stage, the engine checks if the current property is active using the mask obtained in the previous stage, if so passes the associated column data to the next stage.
+In the fourth stage, the engine fetches the vertex data and the matrix row data (\textit{row\_ind, a\_val}) from memory for each active column data and send those values to the connected processing element where the actual computation takes place.
+Each fetch unit is implemented as a Stream Reader engine to handle coarse-grain memory access for the associated data.
+It consists of a request module for sending memory read access command, a fifo buffer for receiving requested blocks and a temporary block buffer for accessing sub-block elements.
+The vertex data (values, masks) are fetched from dedicated vertex caches to enable reuse by the walker engines.
+We use a crossbar to multiplex outgoing requests from the walker engines to the LSU and vertex caches.
+
+\subsection{Processing Elements}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/micro_pe}
+\caption{Processing Element}
+\label{fig:micro_pe}
+\end{figure}
+
+Figure \ref{fig:micro_pe} shows an overview of a processing element microarchitecture.
+It main objective is to perform a floating-point multiplication on the incoming vertex and matrix values from the connected walker engine and accumulate the result until the whole partition has been processed.
+It is comprised of eight parallel pipelined Multiply-Accumulate (MAC) units attached to a coalescing output module.
+Eight is the maximum number of matrix row data that can be stored in a single cache block.
+Each MAC lane consists of pipelined floating-point multiply module, a pipelined floating-point Add unit, a Read-After-Write (RAW) control unit that stalls the pipeline if the current \textit{row\_ind} is pending addition.
+We build a bitmask given the \textit{row\_ind} to track outstanding operations in addition units.
+The accumulated result is stored into a local buffer indexed by the \textit{row\_ind} value.
+Each processing element also computes the aggregate active rows mask for the given partition.
+At the end of the partition processing, the coalescing unit merges output values and masks according to their associated block address and sends the result to the Stream Writers.
+
+\subsection{Stream Writers}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/micro_writer}
+\caption{Stream Writer}
+\label{fig:micro_writer}
+\end{figure}
+
+The Stream Writers are coalescing write buffers that batches write requests to memory until entire blocks are updated before they are submitted.
+We used two stream writers, one for the storing vertex output and one for storing the output mask.
+Figure \ref{fig:micro_writer} shows an overview of a stream writer microarchitecture.
+Internally, a stream writer has an architecture similar to a cache but has been augmented with block ownership information and the stored value is merged, not overwritten.
+It consists of four main modules: A data store, A tag store, and lookup unit and an eviction unit.
+The data store holds all the blocks sent to the device.
+The tag store keeps the meta-data associated with each block in the data store.
+A tag consists of the block address and an owners list.
+The owners list is implemented as a bitmask that tracks the processing elements currently using a given block.
+In a typical operation where the writer receives a block write request, the lookup unit is first invoked to find an existing entry for the corresponding block in the tag store.
+If there is a match and the block is already owned, the incoming value is simply merged with existing one.
+If a match was found, but the block is not owned, the previosuly owned block is released and this new one is acquired before data update.
+If no match is found, the device allocates a new block for the request and evicts existing data to memory if dirty.
+The writer also support an explicit flush command to the eviction of all dirty blocks to memory when all processing is completed.
+
+\subsection{Vertex Caches}
+
+FlexGraph uses two direct-mapped vertex caches to keep the vertex values and mask in the accelerator for reuse by the matrix walker engines.
+Their source port is connected to the load/store unit via an arbiter that multiplexes outgoing read requests with the controller and the matrix walkers.
+It has a typical cache architecture with an input buffer for tracking outstanding requests.
+
+\section{FlexGraph Software-Hardware Codesign}
 
@@
 \end{lstlisting}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/microarchitecture}
-\caption{FlexGraph Microarchitecture}
-\label{fig:microarchitecture}
-\end{figure}
-
-Figure \ref{fig:microarchitecture} illustrates FlexGraph accelerator microarchitecture.
-It is comprised of four main module types, a controller, processing elements (PEs), a load store unit (LSU) and vertex caches.
-The main controller is responsible for starting the accelerator, scheduling tasks for execution, reporting hardware counters and terminating the execution.
-The processing elements execute partition tasks assigned to them by the controller and communicate with the LSU to access the matrix and vertex data for their partition.
-They are also responsible for sending their final output result back to memory.
-The LSU is the module responsible for managing external communication between the accelerator and memory via the QPI interface.
-It directly binds to the QPI ports defined in listing 2.
-The vertex caches store intermediate vertex data and active masks for sharing between the processing elements.
-FlexGraph processing pipeline sightly resembles the SPMV execution steps illustrated in listing 1.
-We made several important modifications to it to improve performance.
-
-\section{FlexGraph Optimizations}
-
-The first optimization we performed in FlexGraph early during the design phase was the support of multiple processing units.
-The reasoning behind it was to extract the maximum bandwidth out of the LSU such that the QPi is always busy processing a request when some processing element are stalled waiting for their data to return.
-the other advantage of parallelization for FlexGraph is to increase the overall accelerator throughput by processing multiple a larger chunk of the workload per unit of time.
-The scheduling of the partitions for execution on the processing elements is controlled by a dispatch unit inside the main controller module.
-The dispatch unit fetches partitions data from the LSU and pass down each partition in a first come first serve fashion to the processing elements.
-
-\subsection{DCSC Matrix partitioning}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/matrix_partitioning}
-\caption{DCSC Matrix Partitioning}
-\label{fig:matrix_partitioning}
-\end{figure}
-
-To enable the parallelization of FlexGraph tasks, the DCSC sparse matrix is first partitioned into aligned partitions of 32 consecutive rows containing non-zero edges.
-we choose a partition size of 32 mainly to match the 32-bit size of the bitmasks encoding the vertices that are active.
-These masks are used by the software on the host processor to determine the regions of the acceleration's output buffer that have been updated.
-Figure \ref{fig:matrix_partitioning} illustrates sample matrix partitioning in which the non-zero horizontal regions have been broken into two partitions P0 and P1.
-The non-zero column ranges \textit{(c0, c1, c2, c3)} represent the selected chunks covered by each partition.
-Partition P0 will only contain ranges \textit{c0} and \textit{c2} while P2 has non-zeros in all four ranges.
-The partitioning scheme is not ideal because of the workload in-balance that might exist when the number of non-zero varies disproportionally between partitions.
-
-\subsection{Synchronising Memory Accesses}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.35\textwidth]{figures/write_masks_synchronization}
-\caption{Active Masks Write Synchronization}
-\label{fig:write_masks_synchronization}
-\end{figure}
-
-An implementation challenge we faced when supporting multiple processing elements in FlexGraph was the synchronization of write accesses for the output active masks.
-Because the LSU data transfer granularity is 64 byte (matching CCI \cite{CCI}), FlexGraph need to manage shared write accesses to the same block to avoid having a processing element override the content of another one.
-Luckily this doesn't pose any problem for writing out partition output values because the 32 rows within a partition occupy 32 * 4 bytes = 2 x 64 bytes blocks.
-Because the partitioning reinforces 32 rows alignment, each processing element can safely write into their assigned blocks.
-However, writing out the active masks poses a challenge because a single 4-byte mask encodes the active states of all 32 rows in a partition.
-This causes the processing elements to potentially share a single 64-byte block when writing the masks out to memory.
-To alleviate the contention, FlexGraph keep N active 64-byte blocks locally assigned to either one of the N processing elements.
-It assigns an ownership bit mask to each block to track their reference count and the processing element assigned to them.
-It also tracks the current address value assigned to each block.
-When a processing element is ready to write its 4-byte active mask, it passes the mask address and value to the LSU.
-The LSU goes to last active block the processing element previous wrote to and check if the block address matches.
-If so, it simply adds the new content to the existing block.
-If there is no match, it clears its ownership bit and flushes the block to memory if the ownership mask goes to zero.
-Then it looks up the other active block if anyone already has the address to use it, otherwise if it acquires the last unused block.
-Figure \ref{fig:write_masks_synchronization} shows a simplified illustration of the scheme with two processing elements.
-We don't need to keep more than N blocks in local storage for this scheme to work because the block address references are always incremental and never regress, meaning that there is always going to be free block available to use.
-It is important to also point out that for this scheme to work, the operation has to be atomic.
-FlexGraph has a single communication channel between all processing elements and the LSU which does some round robin arbitration and blocks the write mask request until it is committed.
-We also investigated an alternative solution to avoid this synchronization, which is simply increasing the partition size to 16 * 32 rows, allowing the aggregated active masks to occupy a full 64-byte block.
-They were two major issues with this approach.
-Firstly, the total size of all resident partitions will be too large if we support multiple processing elements.
-Secondly, it will worsen the workload imbalance that is already present with 32 rows.
-
-This section describes some optimizations we added to FlexGraph to improve its performance.
-
-\subsection{Optimizing Memory Accesses via Stream Buffers}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=0.4\textwidth]{figures/stream_buffers}
-\caption{Stream Buffers}
-\label{fig:stream_buffers}
-\end{figure}
-
-As stated before, FlexGraph's LSU transfer granularity is 64 bytes, which is much larger than the 4 bytes of elements accessed by the processing elements.
-To save on bandwidth, the LSU uses stream buffers that fetch a 64-byte blocks of data but extracts 4-byte elements at the time.
-It is implemented using a fifo structure in the back-end where 64-byte responses from QPI go
- into.
-In the front-end, there is a temporary 64-byte block that is extracted from the fifo on demand to deliver 4-byte element at the time.
-We employ a shift register to extract the 4-byte element to pass it to the processing element.
-Figure \ref{fig:stream_buffers} shows an illustration of the stream buffer concept.
-
-\subsection{Caching Vertex Values and Masks}
-
-Another optimization we implemented in FlexGraph is the caching of vertex values and masks.
-Looking at the SPMV speudo-code in listing 1, we can observe in lines 5 and 6 that there is a random memory access to buffers \textit{x\_values} and \textit{x\_activemask}.
-FlexGraph employs a stream buffer like concept to consume 64-byte vertex data once they arrive inside the processing element to each iteration loop.
-However, the same block can be referenced by another processing element and sharing them inside the cache structure can save unnecessary memory traffic.
-We implemented two small fully associative caches with first-in-First-out replacement to hold active vertex data during execution.
-The caches are implemented inside the LSU which is responsible for managing them.
-When a vertex request arrives from a processing element, the LSU first looks up the cache if the block is already present and return it.
-If the block is not there, it sends the request to QPI.
-QPI output interface support a 14-bit metadata field that is used to identify the block once it is returned.
-We use that field to store the index of the block such that we can compute the cache tag when it arrives.
-Because the cache is small, we implemented a one cycle tag lookup for the LSU to known if the block is present and accessible.
-
-\subsection{Executing Non-blocking Memory writes}
-
-FlexGraph memory writes are all non-blocking, this applies to both the output values and active masks.
-This allows the processing elements to push their write request and resume execution while the LSU is processing them.
-Upon write responses from the QPI channel, the LSU internally keeps track of the count of outstanding requested to ensure that all writes have completed.
-The QPI interface exposes two write response ports by which the memory replies could be sent.
-This allows servicing two responses simultaneously.
-The LSU implements a mechanism to process both channels simultaneously when they are active.
-At completion time when all processing elements are done executing the current run, the controller waits for all outstanding write requests to complete before asserting the \textit{done} signal.
-
-\section{FlexGraph Software-Hardware Codesign}

### 2017-12-18 dcsc_walk.cpp

@@ -4,22 +4,22 @@ using namespace spmv;
 using namespace spmv::accelerator;
 
 __enum (ch_walk_state, (
-  (ready,             1<<0),
-  (get_a_colidx,      1<<1),
-  (get_a_rstart,      1<<2),
-  (wait_for_a_rstart, 1<<3),
-  (get_a_rend,        1<<4),
-  (wait_for_a_rend,   1<<5),
-  (check_x_mask,      1<<6),
-  (get_x_mask,        1<<7),
-  (wait_for_x_mask,   1<<8),
-  (get_x_value,       1<<9),
-  (get_a_rowidx,      1<<10),
-  (get_a_value,       1<<11),
-  (wait_for_data,     1<<12),
-  (execute,           1<<13),
-  (next_column,       1<<14),
-  (end_partition,     1<<15)
+  ready,
+  get_a_colidx,
+  get_a_rstart,
+  wait_for_a_rstart,
+  get_a_rend,
+  wait_for_a_rend,
+  check_x_mask,
+  get_x_mask,
+  wait_for_x_mask,
+  get_x_value,
+  get_a_rowidx,
+  get_a_value,
+  wait_for_data,
+  execute,
+  next_column,
+  end_partition
 ));
 
 spmv_dcsc_walk::spmv_dcsc_walk()
@@ -222,19 +222,19 @@ void spmv_dcsc_walk::describe() {
   }
   __case (ch_walk_state::check_x_mask) {
     // check if the current mask block is valid
-    ch_ptr x_mask_index = a_colidx_ >> 5; // divide by 32 bitmask
+    /*ch_ptr x_mask_index = a_colidx_ >> 5; // divide by 32 bitmask
     ch_ptr x_mask_addr = INT32_TO_BLOCK_ADDR(x_mask_index);
     __if (x_mask_addr == xmblock_addr_) {
       // check if the index is valid
       ch_bit32 mask = (xmblock_ >> INT32_TO_BLOCK_BITSHIFT(x_mask_index)).slice<32>();
-      __if ((mask & (1_b32 << (a_colidx_ & 0x1f))) != 0) {
+      __if ((mask & (1_b32 << (a_colidx_ & 0x1f))) != 0) {*/
         // check if the current value block is valid
         ch_ptr x_value_addr = INT32_TO_BLOCK_ADDR(a_colidx_);
         __if (x_value_addr == xvblock_addr_) {
[…]

### 2017-12-18 lsu.cpp

@@ -4,16 +4,16 @@ using namespace spmv;
 using namespace spmv::accelerator;
 
 __enum (ch_wr_req_state, (
-  (ready,           1<<0),
-  (write_value,     1<<1),
-  (write_mask,      1<<2),
-  (flush,           1<<3),
-  (wait_for_flush,  1<<4)
+  ready,
+  write_value,
+  write_mask,
+  flush,
+  wait_for_flush
 ));
 
 __enum (ch_qpi_write_state, (
-  (ready, 1<<0),
-  (write, 1<<1)
+  ready,
+  write
 ));
 
 spmv_lsu::spmv_lsu() {

### 2017-12-18 pe.cpp

@@ -5,10 +5,10 @@ using namespace spmv;
 using namespace spmv::accelerator;
 
 __enum (ch_pe_state, (
-  (ready,         1<<0),
-  (write_value0,  1<<1),
-  (write_value1,  1<<2),
-  (write_mask,    1<<3)
+  ready,
+  write_value0,
+  write_value1,
+  write_mask
 ));
 
 spmv_pe::spmv_pe() {

### 2017-12-18 spmv.cpp

@@ -4,12 +4,12 @@ using namespace spmv;
 using namespace spmv::accelerator;
 
 __enum (ch_ctrl_state, (
-  (ready,             1<<0),
-  (get_partition,     1<<1),
-  (wait_for_execute,  1<<2),
-  (flush_buffers,     1<<3),
-  (write_stats,       1<<4),
-  (wait_for_writes,   1<<5)
+  ready,
+  get_partition,
+  wait_for_execute,
+  flush_buffers,
+  write_stats,
+  wait_for_writes
 ));
 
 spmv_device::spmv_device() {
@@ -95,7 +95,7 @@ void spmv_device::main_thread() {
       // wait for LSU ack
       __if (lsu_.io.ctrl.rd_req.ready) {
         part_blk_curr_.next = part_blk_curr_ + 1;
-        __if (part_blk_curr_.next == part_blk_end_) {
+        __if (part_blk_curr_ == part_blk_end_) { //=part_blk_curr_.next
           // go wait for execution to complete
           state.next = ch_ctrl_state::wait_for_execute;
         };
@@ -216,8 +216,8 @@ void spmv_device::dispatch_thread() {
 
           // check if we can pop another partition from current block
           // we need at least two entries to proceed
-          __if (part_curr_.next != num_parts_
-             && part_buf_size_.next != 1) {
+          __if (part_curr_ != num_parts_    //=part_curr_.next
+             && part_buf_size_ != 1) {      //=part_buf_size_.next
             // goto next PE
             state.next = 1 + ((i+1 != PE_COUNT) ? (i+1) : 0);
           } __else {

### 2017-12-18 wrcache.h

@@ -13,11 +13,11 @@ private:
   using ch_tag = ch_bit<TagBits>;
 
   __enum (ch_state, (
-    (ready,       1<<0),
-    (lookup,      1<<1),
-    (check_evict, 1<<2),
-    (evict,       1<<3),
-    (flush,       1<<4)
+    ready,
+    lookup,
+    check_evict,
+    evict,
+    flush
   ));
 
   __struct (entry_t, (

### 2018-03-06 paper/conclusion.tex
backup

\section{Future Work}

As stated in the results discussion, FlexGraph performance can be greatly improved once we implement the proposed optimizations. Doing this first step will be our priority for future work.
The Cocoh C++ library we used in this work did not yet have support for verilog generation, we plan to add that support such that we could install the accelerator on the actual FPGA board for evaluation.
We also anticipate to do some power and energy efficiency analysis ion the final design, hopefully identifying all the components in the system consuming the most energy. 
We also plan to add some real world datasets to our evaluation, popular benchmarks include Flicker, Facebook, Wikipedia, Twitter, Netflix and LiveJournal.

\section{Conclusion}

In this paper we presented FlexGraph, a reconfigurable Graph Analytics Accelerator targeting commodity heterogeneous CPU-FPGA platforms. FlexGraph was implemented using Cocoh C++ libary, providing a productive and efficiency development support for modifying the accelerator when needed to accommodate all graph algorithms. FlexGraph architecture implements a customized SPMV kernel adapted for graphs vertex computation. It implements some memory optimizations and tasks parallelization to improve its performance. FlexGraph current performance is not ideal and we identified several strategies to fix that, including doing some memory prefetching and adding additional processing elements.
